### 2011-05-17 paper/paper.tex
Added paper

\documentclass[11pt]{article}
\usepackage{acl-hlt2011}
\usepackage{color}
\usepackage{graphicx}
\usepackage{times}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{algorithmic}
\usepackage{multirow}
\usepackage{url}
\DeclareMathOperator*{\argmax}{arg\,max}
\setlength\titlebox{6.5cm}    % Expanding the titlebox

\title{Towards an Automated Sportscasting System for Chess}
\author{
  Charles Chen \\
  {\tt charleschen@berkeley.edu} \\
\And
  Richard Shin \\
  {\tt ricshin@berkeley.edu} \\
\And
  Jinghao Yan \\
  {\tt jinghao@berkeley.edu} \\
}
\date{}

\begin{document}
\maketitle

\begin{abstract}
We investigate applications of statistical learning to automated sportscasting for chess. In particular, we 
\end{abstract}

\section{Introduction}
Chess has a long history of intriguing enthusiasts due to its complexity. Even after years of play, each game is new challenge, a new problem to solve. It also has a history of prominence as a sport where thousands are engulfed in the battles of grandmasters. Hence, we believed that due to its complexity and agelessness, chess should have a good sportscaster in order to inform and entertain.

In the field of natural language processing, many have investigated ways to derive meaning from text, while others have studied automated text generation. We believe that statistical learning is one approach for automatic sportscasting for chess.

Given the highly intricate nature of chess and other similar abstract strategy games, experts have devised a standard high-level structures like openings, strategies, tactics and time advantage. We intend to leverage statistical learning to identify these structures in chess games.

While chess is one instance where statistical machine learning can help in automated sportscasting, we intend for our approach to be general. Although chess is characterized by many complex rules and interactions, where chess-specific approaches may work better for chess, we shied away from creating too much dependency to chess. As a result, we chose not to use contemporary chess tools like chess engines--libraries that search the game tree to examine the breadth of options available and identify the outcomes of each--specifically because we wanted to be as general as possible.

\begin{itemize}
\item Different applications of similar techniques
\item Differences from sportscasting for, e.g., RoboCup
\item Text about approach
\end{itemize}

\section{Related Work}
\begin{itemize}
\item Ray Mooney
\item Other chess commentating
\item Autoencoding
\end{itemize}

\section{Data sources}
We obtained chess games, both annotated and unannotated, from various websites in the PGN format \cite{pgn}, the universal format for compactly representing chess games, because we anticipated that many open source PGN parsers would be available.

The following websites contain chess games with annotations: \url{http://wwwu.uni-klu.ac.at/gossimit/c/tactic.htm} \url{http://webplaza.pt.lu/public/ckaber/Chess.htm\#Chess\%20training}. We obtained example games from them to feed into our classifier for the supervised component of our system.

For the unsupervised component, we obtained as many chess games as possible from a variety of websites like chessgames.com \cite{other sites}. In all, we obtained over five million untagged games spread over 600,000 files and 5 GB of data. On top of that, we obtained over 5000 tagged games.

\subsection{Data Sanitation}	
In order to ensure consistency and integrity over the games obtained from disparate sources, we sanitized them in multiple ways. We parsed each game in the data set and removed information that would confused the parser. For example, some games had English or German comments by humans that were not structured and thus meaningful for machine learning. Others had alternate branches, sometimes up to multiple levels (alternate branches within alternate branches). Still others represented games in a non-standard way (For example, pawn promotions were represented without the Ò=Ó), which we had to fix. By fixing our input games, we ensured a consistent format for our machine learning system.

In addition, because the games were from multiple sources, there was significant overlap in games (approximately 50%). Therefore, we also went through all of the downloaded games and eliminated duplicate games: Multiple games that had the same initial position and move sequence were considered redundant and only one of each were kept.

\subsection{Parsing}
We tried various open source PGN parsers, each succumbing to flaws like mishandling uncommon cases (like promotions and special moves like the En Passant). Finally we chose chesspresso parser, which we fixed and modified for our needs. Chesspresso takes a PGN file as input and returns a data structure that represents the initial position of the board and the subsequent moves (and all metadata necessary to recreate the PGN file). We also extended it to be more robust to variations in input format, in addition to being able to export PGNs in a canonical form, which was crucial for deduplication.

\section{Features}
\subsection{Sparse Autoencoding}
\subsection{Manual baseline}

\section{Implementation} % Parallelization and Scaling

\section{Classification}

\section{Experiments}
\subsection{Results}
\subsection{Discussion}

\section{Conclusion}

\section{Acknowledgements}
We would like to thank Dan Klein for teaching us statistical machine learning techniques.
\section*{Appendix}

\end{document}

### 2011-05-18 src/edu/berkeley/nlp/classification/WritableFeatureVector.java
ClassifyLabeledData.

@@ -5,8 +5,8 @@ import java.io.Writer;
 
 import jnisvmlight.LabeledFeatureVector;
 
-public class JinghaoFeatureVector extends LabeledFeatureVector {
-	public JinghaoFeatureVector(double d, int[] range, double[] vector) {
+public class WritableFeatureVector extends LabeledFeatureVector {
+	public WritableFeatureVector(double d, int[] range, double[] vector) {
 		super(d, range, vector);
 	}
 

### 2011-05-18 src/edu/berkeley/nlp/classification/ClassifyLabeledData.java
Included F1 score

@@ -135,6 +135,7 @@ public class ClassifyLabeledData {
 
 				WritableFeatureVector lfv = new WritableFeatureVector(1.0, range(
 						1, vector.length + 1), vector);
+
 				lfv.write(writerToInput);
 			}
 			System.out.print('.');
@@ -148,10 +149,10 @@ public class ClassifyLabeledData {
 			}
 			System.out.print('.');
 
-			for (List<PositionWithMoves> example : validationSet.keySet()) {
+			for (List<PositionWithMoves> example : testSet.keySet()) {
 				double[] vector = featureArrays.get(example);
 
-				new WritableFeatureVector(validationSet.containsEntry(example, tag) ? 1.0 : -1.0,
+				new WritableFeatureVector(testSet.containsEntry(example, tag) ? 1.0 : -1.0,
 						range(1, vector.length + 1), 
 						vector)
 				.write(writerToTest);
@@ -173,7 +174,7 @@ public class ClassifyLabeledData {
 			System.out.println(" Done");
 		}
 		
-		printResults(predict(models, testFiles, validationSet));
+		printResults(predict(models, testFiles, testSet));
 	}
 	
 	public static int[] predict(Map<String, File> models, 
@@ -244,6 +245,7 @@ public class ClassifyLabeledData {
 		System.out.printf("False negative: %d (%f)\n", fn, (double) fn
 				/ total);
 		System.out.printf("Precision/Recall: %f, %f\n", precision, recall);
+		System.out.printf("F1 Score: %f\n", 2*precision*recall/(precision+recall));
 	}
 
 	private static int[] range(int from, int to) {

### 2011-05-18 src/edu/berkeley/nlp/chess/PositionWithMoves.java
Update

@@ -61,21 +61,22 @@ public class PositionWithMoves implements Serializable {
 	}
 	
 	static public List<PositionWithMoves> getPrefix(List<PositionWithMoves> input, int limit) {
+		return Lists.newArrayList(input.subList(0, limit));/*
 		List<PositionWithMoves> result = Lists.newArrayList();
 		int numUsed = 0;
 		for (PositionWithMoves pwm : input) {
-			if (numUsed == limit - 1) {
-				result.add(new PositionWithMoves(pwm.previousMove, pwm.position, null));
-				++numUsed;
-				break;
-			} else {
+			if (numUsed < limit) {
 				result.add(pwm);
 				if (pwm.nextMove == null) numUsed += 1;
 				else numUsed += 2;
+			} else {
+				result.add(new PositionWithMoves(pwm.previousMove, pwm.position, null));
+				++numUsed;
+				break;
 			}
 		}
 		Preconditions.checkState(numUsed == limit);
-		return result;
+		return result;*/
 	}
 	
 	public static class ConcatenatingFeaturizer implements Featurizer<PositionWithMoves> {

### 2011-05-18 src/edu/berkeley/nlp/chess/pgn/LabeledGamesProcessor.java

@@ -90,9 +90,11 @@ public class LabeledGamesProcessor {
 						continue;
 					}
 					counts[boards.size() - 1]++;
-					for (int i = 1; i <= sliceLength && i <= boards.size(); ++i) {
+					
+//					for (int i = 1; i <= sliceLength && i <= boards.size(); ++i) {
+					if (boards.size() >= sliceLength) {
 						String tag = getTag(filename);
-						tags.put(PositionWithMoves.getPrefix(boards, i), tag);
+						tags.put(PositionWithMoves.getPrefix(boards, sliceLength), tag);
 						tagsSet.add(tag);
 					}
 

### 2011-05-18 src/edu/berkeley/nlp/classification/JinghaoFeatureVector.java

@@ -6,6 +6,8 @@ import java.io.Writer;
 import jnisvmlight.LabeledFeatureVector;
 
 public class JinghaoFeatureVector extends LabeledFeatureVector {
+	private static final long serialVersionUID = -2361850878723731574L;
+
 	public JinghaoFeatureVector(double d, int[] range, double[] vector) {
 		super(d, range, vector);
 	}

### 2011-05-18 paper/paper.tex
Sad code that demonstrates random is better than SVM

@@
 Even after years of play, each game is new challenge, a new problem to solve.
-It also has a history of prominence as a sport where thousands are engulfed in the battles of grandmasters.
-Hence, we believed that due to its complexity and agelessness, chess should have a good sportscaster in order to inform and entertain.
+Historically, chess has been prominent as a sport where the battles of grandmasters engulf thousands; given the highly intricate nature of chess (and other similar abstract strategy games), experts have scrutinized these games to devise a standard high-level structures like openings, strategies, tactics, and time advantage, which they use to explain maneuvers and blunders that occur in the games.
+Devotees of chess also attempt to improve their own performance in terms of these high-level structures, reading chess theorists' comments about games by high-ranking players.
 
-In the field of natural language processing, many have investigated ways to derive meaning from text, while others have studied automated text generation.
-We believe that statistical learning is one approach for automatic sportscasting for chess.
+In this paper, with the end goal of automatically producing textual commentary similar to those produced by experts for grandmaster games, we investigate whether we might automatically identify such high-level patterns in real games.
+Specifically, we want to see how statistical machine learning techniques can help us automatically identify such structures using a large number of real chess games as training data, where some of the games may be already annotated with these high-level structures, while others may not be.
 
-Given the highly intricate nature of chess and other similar abstract strategy games, experts have devised a standard high-level structures like openings, strategies, tactics and time advantage.
-We intend to leverage statistical learning to identify these structures in chess games.
+While others have created highly rigid and rule-based approaches to annotating chess games~\cite{cambridge-chess-annotation}, we specifically sought to avoid such approaches which require a lot of manual engineering and chess-specific expertise in order to obtain good results.
+In particular, for chess, there exists a large number of computer chess tools like chess engines, which search the game tree to examine the breadth of options available and identify the outcomes of each.
+We would like our approaches to have general applicability to other domains, so we chose to specifically avoid using such tools as much as possible.
 
-While chess is one instance where statistical machine learning can help in automated sportscasting, we intend for our approach to be general.
-Although chess is characterized by many complex rules and interactions, where chess-specific approaches may work better for chess, we shied away from creating too much dependency to chess.
-As a result, we chose not to use contemporary chess tools like chess engines--libraries that search the game tree to examine the breadth of options available and identify the outcomes of each--specifically because we wanted to be as general as possible.
+To that end, we investigate the use of unsupervised feature learning and deep learning techniques on transcripts of chess games to see if they could automatically identify the high-level structures that make up a chess game.
+Deep learning techniques, like the sparse stacked auto-encoder, learns a more efficient way to represent some input given a large set of examples.
+In other fields such as computer vision, researchers have found that these techniques can give state-of-the-art results rivaling or exceeding features hand-engineered by experts.
+While learning a more compact way to represent parts of chess games, auto-encoders may be able to discover high-level structures similar to those used by human experts.
+
+As a comparison, we also implement a small number of chess-specific, hand-engineered features based on information from a chess engine, such as the number of moves available from a specific board position.
+Given feature vectors made from hand-labeled games, we train SVM classifiers to predict the label of an unlabeled feature vector.
+We compare its performance to automatically learned features to see which one can more successfully predict high-level structures.
+
+To conduct our experiments, we collect several thousand ``tactical training positions'' organized by theme, intended for use by human chess players to learn how to recognize such situations and improve their playing skills.
+We also collect a large number of unannotated chess games (about 2.5 million) from various sources on the web for use with semi-supervised and unsupervised learning techniques.
+After training our models, for each of the positions in the training data labeled with a theme (or high-level structure), we attempt to predict the correct theme(s) for a position using the models.
+
+Unfortunately, we found that our approaches did not deliver accurate
 
 \begin{itemize}
+
 \item Different applications of similar techniques
@@
 \section{Related Work}
-\begin{itemize}
-\item Ray Mooney
-\item Other chess commentating
-\item Autoencoding
-\end{itemize}
+\paragraph{Deep learning.}
+\paragraph{Automated sportscasting.}
+\paragraph{Automatic chess commentary.}
 
-\section{Data sources}
+\section{Data Sources}
+
 We obtained chess games, both annotated and unannotated, from various websites in the PGN format \cite{pgn}, the universal format for compactly representing chess games, because we anticipated that many open source PGN parsers would be available.
 
-The following websites contain chess games with annotations: \url{http://wwwu.uni-klu.ac.at/gossimit/c/tactic.htm} \url{http://webplaza.pt.lu/public/ckaber/Chess.htm\#Chess\%20training}.
+The following websites contain chess games with annotations: \url{http://wwwu.uni-klu.ac.at/gossimit/c/tactic.htm} \url{http://webplaza.pt.lu/public/ckaber/Chess.htm#Chess training}.
 We obtained example games from them to feed into our classifier for the supervised component of our system.
@@
 Others had alternate branches, sometimes up to multiple levels (alternate branches within alternate branches).
-Still others represented games in a non-standard way (For example, pawn promotions were represented without the Ò=Ó), which we had to fix.
+Still others represented games in a non-standard way (For example, pawn promotions were represented without the “=”), which we had to fix.
 By fixing our input games, we ensured a consistent format for our machine learning system.
@@
 \subsection{Parsing}
-We tried various open source PGN parsers, each succumbing to flaws like mishandling uncommon cases (like promotions and special moves like the En Passant).
-Finally we chose chesspresso parser, which we fixed and modified for our needs.
+We tried various open source PGN parsers, each succumbing to flaws like mishandling uncommon cases (like promotions and special moves like en passant).
+Finally we chose Chesspresso parser, which we fixed and modified for our needs.
 Chesspresso takes a PGN file as input and returns a data structure that represents the initial position of the board and the subsequent moves (and all metadata necessary to recreate the PGN file).
@@
 \section{Features}
+
 \subsection{Sparse Autoencoding}
 \subsection{Manual baseline}
+As a first step towards classification, we selected chess-specific manual features from a given chess position (and a previous move):
+\begin{table*}
+\begin{tabular}{lp{0.6\textwidth}}
+\hline
+\textbf{Name} & \textbf{Description} \\ \hline
+\tt{materialCount()} & \\
+\tt{squareDominationCount()} & \\
+\tt{isCapturing()} & \\
+\tt{isEnPessant()} & \\
+\tt{isLongCastle()} & \\
+\tt{isCheck()} & \\
+\tt{isStalemate()} & \\
+\tt{isMate()} & \\
+\tt{isPromotion()} & \\
+
+\end{tabular} 
+\caption{List of chess-specific manual features per position.}
+\end{table*}
+
+\section{Classification}
+We built a support vector machine for structured classification of games.
+We filtered our tagged games by the number of plies (number of individual moves), generated mappings between tags and game prefixes (some initial position plus a sequence of moves), segmented our tagged games into three groups (training, validation and test), and called SVMLight with that data, an external library for SVM.
+
+\subsection{High-level descriptions}
+We classify segments of the chess game based on various high-level descriptions from chess theory and folklore:
+
+\begin{table*}
+\begin{tabular}{lp{0.6\textwidth}}
+\hline
+\textbf{Name} & \textbf{Description} \\ \hline
+\tt{attackPossible} & The current player has opportunity to attack. \\
+\tt{backrankWeakness} & There exists a future checkmate by a rook or queen on the edges of the board. \\
+\tt{bishopPair} & There exists a decisive strategic advantage for the player with two bishops. \\
+\tt{clearance} & There exists a maneuver to break the pawn structure. \\ 
+\tt{closure} & There exists a maneuver to fix the pawn structure. \\
+\tt{decoy} & There exists a maneuver to distract a crucial piece. \\
+\tt{defensePressure} & There exists a maneuver to break the opposing King’s defense. \\
+\tt{deflection} & There exists a maneuver to deflect an attack.  \\ 
+\tt{discoveredAttack} & There exists a maneuver to allow for a discovered attack (an attack that was blocked by a friendly piece.) \\
+\tt{doubleAttack} & There exists a move that simultaneously attacks at least two crucial squares. \\
+\tt{drawCombinationsStalemate} & There exists a combination that allows for decisive material advantage or immediate checkmate. \\
+\tt{enclosedKings} & There exists a maneuver to trap the opposing King. \\
+\tt{freeingBlockingSquares} & There exists a move to free a crucial square. \\
+\tt{goodEndgameChances} & There exists a move to allow for a favorable endgame. \\
+\tt{inBetweenMove} & There exists an interrupting move that allows for decisive material advantage of immediate checkmate. \\ 
+\tt{interference} & There exists a maneuver to block a crucial piece. \\ 
+\tt{interruption} & There exists a maneuver that interrupts the opponent’s tactics. \\
+\tt{kingInCenter} & There exists a decisive strategic advantage with the player with the King in the center of the board. \\
+\tt{kingToBadSquares} & There exists a combination that forces the opposing King to move to a bad square. \\
+\tt{longDiagonals} & There exists a strategic advantage to the player that controls squares on the long diagonal. \\
+\tt{matingAttack} & There exists a decisive mating attack. \\
+\tt{openingClosingDiagonals} & There exists a maneuver to clear a diagonal. \\
+\tt{openingClosingFiles} & There exists a maneuver to clear a file. \\
+\tt{overload} & There exists a combination that makes a defending piece defend too many squares at once. \\
+\tt{passedPawn} & There exists a combination that allows for a passed pawn (a pawn with no opposing pawns in front of it) \\
+\tt{pawnPromotion} & There exists a combination that allows for a pawn promotion. \\
+\tt{perpetualCheck} & There exists a combination that forces a perpetual check/draw. \\
+\tt{pin} & There exists a combination that allows for a pin. \\
+\tt{pullingPiecesToBadSquares} & There exists a combination that forces an opposing piece to move to a bad square. \\
+\tt{queenCapture} & There exists a combination to capture the opposing Queen. \\
+\tt{quietMates} & There exists a quiet move (one that looks inconsequential on the surface) leading to mate. \\
+\tt{removalOfDefence} & There exists a maneuver to remove an opponent’s defender. \\
+\tt{unpinning} & There exists a maneuver to unpin a piece. \\
+\tt{wrongBishop} & The opposing player’s only bishop is limited in use because of the pawn structure of the game. \\ \hline
+\end{tabular} 
+\caption{List of high-level themes we attempt to automatically identify.}
+\end{table*}
 
 \section{Implementation} % Parallelization and Scaling
-
-\section{Classification}
+Running serially, the autoencoder takes a long time to complete, even on simple inputs.
+Therefore, we examined ways to parallelize the task in order to fully leverage the capacity of our 8-core server with hyper-threading (16 virtual cores).
+We implemented a local version of MapReduce that creates 16 execution threads.
 
@@
 \subsection{Discussion}
+We discovered that there is an inherent trade off between how general the approach is and the quality of the outcomes.
+In pursuing this project, we attempted to maintain an approach that is as general as possible so that it can be applied to other contexts.
+However, in attempting to maintain generality, we sacrificed access to powerful but chess-specific tools like game engines, resulting in a classifier that is more surface-level (yet slower) than existing chess analysis tools that leverage the game engine.
+
+\section{Future Work}
+There are three areas in which future work can make progress on automated sportscasting.
+It could improve the performance of the autoencoder so that it can be trained on a larger data set.
+Furthermore, it could focus on trying different methods to improve the classification quality by leveraging chess-specific tools and features like chess engines and evaluation tools.
+Moreover, it could investigate an orthogonal path by applying our general approach to other games, including more complex ones like StarCraft and simpler ones like Checkers.
+
+\subsection{Performance}
+One of the main limitations of our work is performance.
+Due to how slowly the autoencoder performs, we were not able to fully exploit the millions of games that we have available to generate features for our unsupervised system.
+Therefore, extensions to our work should investigate ways to improve the performance so that the autoencoder can process more inputs and thus generate more relevant feature functions.
+
+\paragraph{Distributed MapReduce.}
+Our autoencoder takes advantage of some of the parallel processing capacity of the host system through a local MapReduce scheme.
+Extending that with a distributed MapReduce framework like Hadoop may improve performance if the additional cost of network I/O are outweighed by the benefits of parallelization.
+
+\paragraph{Using the GPU.}
+Another option that we partially explored is the exploitation of the processing power of the graphics processing unit (GPU).
+Optimization is a computation-heavy task, and the GPU is specifically designed for massively-parallel floating point computations.
+We investigated using JavaCL to parallelize the autoencoder, and demonstrated that it is definitely superior at parallelizable and computation-intensive tasks like matrix multiplication, because the benefits of greater throughput through greater parallelization outweigh the costs of memory transfer.
+
+\subsection{Classification Quality}
 

### 2011-05-18 src/edu/berkeley/nlp/classification/ClassifyLabeledData.java

@@ -10,6 +10,7 @@ import java.text.ParseException;
 import java.util.Collection;
 import java.util.List;
 import java.util.Map;
+import java.util.Random;
 import java.util.Scanner;
 import java.util.Set;
 
@@ -31,13 +32,30 @@ import edu.berkeley.nlp.chess.util.GzipFiles;
 import edu.berkeley.nlp.util.CommandLineUtils;
 
 public class ClassifyLabeledData {
+	String dataPath, autoEncoderPath;
+
+	Multimap<String, List<PositionWithMoves>> allPositiveExamples, allNegativeExamples;
+	Multimap<List<PositionWithMoves>, String> tags, validationSet, testSet;
+	
 	public static void main(String[] args) throws IOException,
-			ClassNotFoundException, ParseException, URISyntaxException,
+			ClassNotFoundException, ParseException,
 			InterruptedException {
 		Map<String, String> argMap = CommandLineUtils.simpleCommandLineParser(args);
 		String dataPath = argMap.get("-data");
 		String autoEncoderPath = argMap.get("-autoencoder");
 
+		ClassifyLabeledData classifier = new ClassifyLabeledData(dataPath, autoEncoderPath);
+		classifier.classify();
+	}
+	
+	public ClassifyLabeledData(String dataPath, String autoEncoderPath) {
+		this.dataPath = dataPath;
+		this.autoEncoderPath = autoEncoderPath;
+	}
+	
+	public void classify() throws IOException,
+	ClassNotFoundException, ParseException,
+	InterruptedException {
 		Featurizer<List<PositionWithMoves>> featurizer;
 		
 		if (autoEncoderPath != null) {
@@ -79,25 +97,11 @@ public class ClassifyLabeledData {
 		ObjectInputStream ois = GzipFiles.newObjectInputStreamSupplier(
 				new File(dataPath)).getInput();
 
-		@SuppressWarnings("unchecked")
-		Multimap<List<PositionWithMoves>, String> tags = (Multimap<List<PositionWithMoves>, String>) ois
-				.readObject();
-
-		@SuppressWarnings("unchecked")
-		Multimap<String, List<PositionWithMoves>> allPositiveExamples = (Multimap<String, List<PositionWithMoves>>) ois
-				.readObject();
-
-		@SuppressWarnings("unchecked")
-		Multimap<String, List<PositionWithMoves>> allNegativeExamples = (Multimap<String, List<PositionWithMoves>>) ois
-				.readObject();
-
[…]

### 2011-05-18 paper/paper.tex
Partition based on first player's color

@@
 
-Unfortunately, we found that our approaches did not deliver accurate
+Unfortunately, we found that our approaches did not deliver accurate results in our experiments.
+In particular, the auto-encoder's output failed to serve as a useful criterion for distinguishing between different types of themes.
+We examine the types of correct predictions made by our classifiers and discuss how we may improve our results in the future.
 
@@
 \section{Related Work}
-\paragraph{Deep learning.}
+\paragraph{Deep learning.} These techniques attempt to capture complex relationships between input variables in order to transform the input into a different representation which better expresses the relevant structures that make up the input.
+Specifically, they compose many layers of non-linearities in order to compactly represent complex functions which can model these relationships.
+They have been used successfully for many different types of tasks \cite{ng-mnist,ng-vision,ng-nlp}.
+
 \paragraph{Automated sportscasting.}
+
 \paragraph{Automatic chess commentary.}
@@
 \subsection{Manual baseline}
-As a first step towards classification, we selected chess-specific manual features from a given chess position (and a previous move):
+As a first step towards classification, we selected chess-specific manual features from a given chess position-previousmove pair.
+The position-based features involve information such as where the pieces are on the board and potential moves, while the move-based features involve binary information such as if the move is a capturing move or a check.
+We list all our hand-selected manual features in Table REF.
+
 \begin{table*}
@@
 \hline
-\textbf{Name} & \textbf{Description} \\ \hline
-\tt{materialCount()} & \\
-\tt{squareDominationCount()} & \\
-\tt{isCapturing()} & \\
-\tt{isEnPessant()} & \\
-\tt{isLongCastle()} & \\
-\tt{isCheck()} & \\
-\tt{isStalemate()} & \\
-\tt{isMate()} & \\
-\tt{isPromotion()} & \\
-
+\textbf{Category} & \textbf{Description} \\ \hline
+Position features & Piece location, potential moves per piece, material count, square control. \\
+Move features & Capturing move, recapturing move, en passant move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move.
 \end{tabular} 

### 2011-05-18 src/edu/berkeley/nlp/chess/pgn/LabeledGamesProcessor.java

@@ -36,13 +36,11 @@ import edu.berkeley.nlp.classification.Featurizers;
 import edu.berkeley.nlp.util.CommandLineUtils;
 
 public class LabeledGamesProcessor {
-	public static String getTag(String filename) {
-		return filename.substring(filename.lastIndexOf('/') + 1, filename.lastIndexOf('.'));
+	public static String getTag(String filename, boolean isWhiteMove) {
+		return filename.substring(filename.lastIndexOf('/') + 1, filename.lastIndexOf('.')) + (isWhiteMove ? " (white)" : " (black)");
 	}
 	
 	public static void main(String[] args) throws IOException {
-		assert getTag("themes/t/deflection.pgn-fixed").equals("deflection") : "getTag doesn't work";
-
 		Injector injector = Guice.createInjector(new AbstractModule() {
 			@Override
 			protected void configure() {
@@ -84,18 +82,26 @@ public class LabeledGamesProcessor {
 					if (g == null) break;
 
 					lineNumber = reader.getLineNumber();
+					
 					List<PositionWithMoves> boards = Games.flatten(g);
 					if (boards.size() > maxLength) {
 						++largeGames;
 						continue;
 					}
+					
+					
+					
 					counts[boards.size() - 1]++;
 					
 //					for (int i = 1; i <= sliceLength && i <= boards.size(); ++i) {
 					if (boards.size() >= sliceLength) {
-						String tag = getTag(filename);
-						tags.put(PositionWithMoves.getPrefix(boards, sliceLength), tag);
-						tagsSet.add(tag);
+						try {
+							String tag = getTag(filename, boards.get(0).nextMove.isWhiteMove());
+							tags.put(PositionWithMoves.getPrefix(boards, sliceLength), tag);
+							tagsSet.add(tag);
+						} catch (NullPointerException e) {
+							System.out.printf("Missing game (line %d of good game; error line %d): %s", lineNumber, reader.getLineNumber(), e.getMessage());
+						}
 					}
 
 					++games;

### 2011-05-18 paper/paper.tex
Added segmented data

@@
 
+\section{Related Work}
 \begin{itemize}
-
 \item Different applications of similar techniques
@@
 
-\section{Related Work}
 \paragraph{Deep learning.} These techniques attempt to capture complex relationships between input variables in order to transform the input into a different representation which better expresses the relevant structures that make up the input.
@@
 
-\paragraph{Automatic chess commentary.}
+\paragraph{Automatic chess commentary.} There exist techniques to automatically annotate chess games which are heavily dependent on rule-based logic and use of a chess-engine to determine which parts of a specific chess game are salient~\cite{cambridge-chess-annotation}.
+As a departure from the rule-based approach to automatic chess commentary, Guid et al.
+(\cite{learning-positional-features}) discuss the use of argument based machine learning, which analyzes reasoning where arguments for and against a certain claim are produced and evaluated (in the context of chess, why a certain move was played over others.) Their system, much like the one presented in this paper, identifies complex positional features of a given chess game.
 
-\section{Data Sources}
+However, we identified two shortcomings of the AMBL system.
+First, it relies on experts to provide arguments as to why a certain position exemplifies (or doesn’t) a particular positional feature.
+While this is a step up from defining a classification rule by hand, it is infeasible to iterate this procedure multiple times.
+Moreover, Guid described that
 
-We obtained chess games, both annotated and unannotated, from various websites in the PGN format \cite{pgn}, the universal format for compactly representing chess games, because we anticipated that many open source PGN parsers would be available.
+\section{Data Sources and Processing}
+We obtained both annotated and unannotated records of chess games from various web sites in the PGN format \cite{pgn}, a widely-used format for compactly representing chess games.
+A large number of games as well as software and tools that use the PGN format are freely available.
 
-The following websites contain chess games with annotations: \url{http://wwwu.uni-klu.ac.at/gossimit/c/tactic.htm} \url{http://webplaza.pt.lu/public/ckaber/Chess.htm#Chess training}.
-We obtained example games from them to feed into our classifier for the supervised component of our system.
+The already-annotated games come from databases of short snippets of chess games,\footnote{\url{http://wwwu.uni-klu.ac.at/gossimit/c/tactic.htm}} \footnote{\url{http://webplaza.pt.lu/public/ckaber/Chess.htm#Chess training}} intended for chess players to learn from by example, which are organized by particular themes that the students are to identify.
+These provided a useful repository of both interesting high-level concepts we could identify (we provide a listing in Table~\ref{table:themes}), as well as boards and moves corresponding to each one.
 
-For the unsupervised component, we obtained as many chess games as possible from a variety of websites like chessgames.com \cite{other sites}.
-In all, we obtained over five million untagged games spread over 600,000 files and 5 GB of data.
-On top of that, we obtained over 5000 tagged games.
+We also looked for as many chess games as possible from a variety of web sites, such as \url{chessgames.com}.
+We easily obtained over five million untagged games spread over 600,000 files and 5 GB of data.
+The number of annotated game snippets we obtained was slightly over 5,000.
 
-\subsection{Data Sanitation}	
+\subsection{Data Pre-Processing}	
 In order to ensure consistency and integrity over the games obtained from disparate sources, we sanitized them in multiple ways.
-We parsed each game in the data set and removed information that would confused the parser.
-For example, some games had English or German comments by humans that were not structured and thus meaningful for machine learning.
-Others had alternate branches, sometimes up to multiple levels (alternate branches within alternate branches).
-Still others represented games in a non-standard way (For example, pawn promotions were represented without the “=”), which we had to fix.
-By fixing our input games, we ensured a consistent format for our machine learning system.
+We parsed each game in the data set and removed information that would confuse the parser.
+For example, some games had English or German comments by humans which we did not plan to use, and sometimes resulted in parsing errors due to specific characters present in the comments.
+Others had alternate branches, sometimes up to multiple levels (alternate branches within alternate branches), which the parser could not handle properly.
+Still others represented games in a non-standard way (for example, pawn promotions were represented without the ``=''), which we had to fixed.
 
-In addition, because the games were from multiple sources, there was significant overlap in games (approximately 50%).
-Therefore, we also went through all of the downloaded games and eliminated duplicate games: Multiple games that had the same initial position and move sequence were considered redundant and only one of each were kept.
+Since we obtained the games from many sources, there was significant overlap in games (approximately 50% of the unlabeled games collected were redundant).
+Therefore, we also went through all of the downloaded games and eliminated duplicate games: multiple games that had the same initial position and move sequence were considered redundant and only one of each were kept.
 
@@
 We tried various open source PGN parsers, each succumbing to flaws like mishandling uncommon cases (like promotions and special moves like en passant).
-Finally we chose Chesspresso parser, which we fixed and modified for our needs.
-Chesspresso takes a PGN file as input and returns a data structure that represents the initial position of the board and the subsequent moves (and all metadata necessary to recreate the PGN file).
-We also extended it to be more robust to variations in input format, in addition to being able to export PGNs in a canonical form, which was crucial for deduplication.
+We based our system on Chesspresso,\footnote{url{http://www.chesspresso.org/}} a Java library for writing chess programs, which we fixed and modified for our needs.
+Chesspresso takes a PGN file as input and returns a data structure that represents the initial position of the board and the subsequent moves (as well as all metadata necessary to recreate the PGN file).
+We also extended it to be more robust to variations in input format, in addition to being able to export PGNs in a canonical form, which was crucial for removing duplicates.
 
 \section{Features}
-
-\subsection{Sparse Autoencoding}
-\subsection{Manual baseline}
-As a first step towards classification, we selected chess-specific manual features from a given chess position-previousmove pair.
+\subsection{Manually Generated Features}
+As a first step towards classification, we selected chess-specific manual features from a given position-move pair.
 The position-based features involve information such as where the pieces are on the board and potential moves, while the move-based features involve binary information such as if the move is a capturing move or a check.
@@
 \end{table*}
+
+\subsection{Sparse Autoencoding}
 
@@
 \end{tabular}
-\caption{List of high-level themes we attempt to automatically identify.}
+\caption{List of high-level themes we attempt to automatically identify, along with explanations}
+\label{table:themes}
 \end{table*}
@@
 \section{Experiments}
+\subsection{SVM Based Structured Classifier}
+
+In our experiments, we processed the input data into multimaps from game prefixes to tags.
+We then segmented the data into three disjoint sets: 90% for training, 5% as the validation set (for tuning our classifiers), and 5% as the test set.
+The segmenting was done by randomly sampling from the set of all labeled games.
+
+Afterwards, we constructed lists of positive and negative labels (mappings from tags to games that have that tag, and games that do not have that tag).
+We then built SVM models using the training data.
+In testing, we fed that to the SVM to generate, for each test game and tag, a boolean prediction of whether that tag applies to that game.
+As a result, we constructed a SVM-based multiple-class classifier that attempts to predict the tags that characterize an input game.
+
+We measured the accuracy of our model using various measures like false positive rate, false negative rate, precision, recall, and F1.
+In addition, we compared the predictions made by SVM to those made randomly where each game was assigned a tag at random based on the proportion of the games that have that tag.
+The random predictions were made 1,000 times, with the predictions made independently at random, and the error rates were averaged.
+
 \subsection{Results}
+Since each game has very few tags, most of the binary classifications are negative. As a result, the model is heavily biased toward negative results. Our results bear this disposition:
+
+INSERT TABLE HERE
+
+The basic classifier generates classifications with relatively high precision compared to a random classifier.
+However, the recall is much lower than that of the random classifier.
+As a result, the F1 score (a combination of the precision and recall scores) of the SVM-based model is no better than that of the random model.
+
+We then extended the tagging to also incorporate information about the player
+
 \subsection{Discussion}
@@
 Furthermore, it could focus on trying different methods to improve the classification quality by leveraging chess-specific tools and features like chess engines and evaluation tools.
-Moreover, it could investigate an orthogonal path by applying our general approach to other games, including more complex ones like StarCraft and simpler ones like Checkers.
+Moreover, it could investigate an orthogonal path by applying our general approach to other games, including complex ones like StarCraft and simpler ones like Checkers.
 
@@
 \end{document}
+
+stack of work:
+Segmented:
+True positive: 13 (0.001252)
+False positive: 2 (0.000193)
+True negative: 10213 (0.983438)
+False negative: 157 (0.015118)
+Precision/Recall: 0.866667, 0.076471
+F1 Score: 0.140541
+True positive: 12841 (0.001236)
+False positive: 154930 (0.014919)
+True negative: 10060070 (0.968712)
+False negative: 157159 (0.015133)
+Precision/Recall: 0.076539, 0.075535
+F1 Score: 0.076034
+
+Unsegmented:
+True positive: 11 (0.002087)
+False positive: 0 (0.000000)
+True negative: 5106 (0.968880)
+False negative: 153 (0.029032)
+Precision/Recall: 1.000000, 0.067073
+F1 Score: 0.125714
+True positive: 22387 (0.004248)
+False positive: 145688 (0.027645)
+True negative: 4960312 (0.941236)
+False negative: 141613 (0.026872)
+Precision/Recall: 0.133196, 0.136506
+F1 Score: 0.134831
+
+\begin{table*}
+\begin{tabular}{lp{0.6\textwidth}}
+\hline
+\textbf{True Positive} & \textbf{False Positive} & \textbf{True Negative} & \textbf{False Negative} & \textbf{Precision} & \textbf{Recall} & \textbf{F1} \\ \hline
+Position features & Piece location, potential moves per piece, material count, square control. \\
+Move features & Capturing move, recapturing move, en passant move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move.
+\end{tabular} 
+\caption{List of chess-specific manual features per position.}
+\end{table*}

### 2011-05-18 src/edu/berkeley/nlp/chess/pgn/LabeledGamesProcessor.java

@@ -61,7 +61,6 @@ public class LabeledGamesProcessor {
 		String prefix = argMap.get("-prefix");
 		String outputPath = argMap.get("-output");
 		int sliceLength = Integer.parseInt(argMap.get("-sliceLength"));
-		String extraFeatures = argMap.get("-extra-features"); 
 		
 		long games = 0, largeGames = 0;
 		Multimap<List<PositionWithMoves>, String> tags = HashMultimap.create();

### 2011-05-18 paper/paper (copy).tex
Broken paper

--- 2011-05-18 paper/paper.tex
@@
 Even after years of play, each game is new challenge, a new problem to solve.
-Historically, chess has been prominent as a sport where the battles of grandmasters engulf thousands; given the highly intricate nature of chess (and other similar abstract strategy games), experts have scrutinized these games to devise a standard high-level structures like openings, strategies, tactics, and time advantage, which they use to explain maneuvers and blunders that occur in the games.
+Historically, chess has been prominent as a sport where the battles of grandmasters engulf thousands; given the highly intricate nature of chess (and similar abstract strategy games), experts have scrutinized these games to devise a standard high-level structures like openings, strategies, tactics, and time advantage, which they use to explain maneuvers and blunders that occur in the games.
 Devotees of chess also attempt to improve their own performance in terms of these high-level structures, reading chess theorists' comments about games by high-ranking players.
@@
 \section{Related Work}
-\begin{itemize}
-\item Different applications of similar techniques
-\item Differences from sportscasting for, e.g., RoboCup
-\item Text about approach
-\end{itemize}
 
@@
 
-\paragraph{Automated sportscasting.}
+\paragraph{Automated sportscasting.} Chen and Mooney (\cite{learning-to-sportscast}) present a commentator system that learns language from sportscasts of simulated RoboCup games.
+Their system, which is trained on ambiguous supervision in the form of textual human commentaries and simulation states of the soccer games, simultaneously tries to establish correspondences between the commentaries and the simulation states.
+Their work focuses on generating meaningful English commentary from relatively simple topics such as ‘Player 1 kicked the ball to Player 2’, while our work focuses more on identifying complex topics.
 
-\paragraph{Automatic chess commentary.} There exist techniques to automatically annotate chess games which are heavily dependent on rule-based logic and use of a chess-engine to determine which parts of a specific chess game are salient~\cite{cambridge-chess-annotation}.
+\paragraph{Automatic chess commentary.} There exist techniques for automatic annotation of chess games which are heavily dependent on rule-based logic and use of a chess-engine to determine which parts of a specific chess game are salient~\cite{cambridge-chess-annotation}.
 As a departure from the rule-based approach to automatic chess commentary, Guid et al.
@@
 
-However, we identified two shortcomings of the AMBL system.
+However, we identified some shortcomings of the AMBL system.
 First, it relies on experts to provide arguments as to why a certain position exemplifies (or doesn’t) a particular positional feature.
-While this is a step up from defining a classification rule by hand, it is infeasible to iterate this procedure multiple times.
-Moreover, Guid described that
+While this is an improvement from defining a classification rule manually, it is infeasible to repeat this procedure multiple times.
+Second, the arguments that the experts are allowed to provide rely on a preset list of features.
 
@@
 Others had alternate branches, sometimes up to multiple levels (alternate branches within alternate branches), which the parser could not handle properly.
-Still others represented games in a non-standard way (for example, pawn promotions were represented without the ``=''), which we had to fixed.
+Still others represented games in a non-standard way (for example, pawn promotions were represented without the ``=''), which we had to fix.
 
@@
 \subsection{Parsing}
-We tried various open source PGN parsers, each succumbing to flaws like mishandling uncommon cases (like promotions and special moves like en passant).
+We tried various open source PGN parsers, each succumbing to flaws like mishandling uncommon cases (such as  promotions and special moves like \emph{en passant}).
 We based our system on Chesspresso,\footnote{url{http://www.chesspresso.org/}} a Java library for writing chess programs, which we fixed and modified for our needs.
@@
 
+\subsection{Preparing the Data}
+After parsing each game, we filtered out the tagged games that had more than 10-plies
+JINGHAO: Write about slicing and tag generation
+
 \section{Features}
-\subsection{Manually Generated Features}
-As a first step towards classification, we selected chess-specific manual features from a given position-move pair.
-The position-based features involve information such as where the pieces are on the board and potential moves, while the move-based features involve binary information such as if the move is a capturing move or a check.
-We list all our hand-selected manual features in Table REF.
+\subsection{Sparse auto-encoder}
+The sparse auto-encoder is one of many methods in use for unsupervised feature learning~\cite{single-layer-networks-unsupervised-feature-learning}.
+It takes the form of a feedforward neural network, which consists of an input layer of N nodes, some number of hidden layers, and output layer of M nodes.
+The value of each layer $\vec{l_i}$ is determined with a weight matrix $W_i$ and a bias vector $b_i$: \[L_i = f(W_i l_{i-1} + b_i),\] where $f(\cdot)$ is a non-linearity like the sigmoid $f(x) = \frac{1}{1 + e^{-z}}$ applied component-wise.
 
+An auto-encoder is such a neural network trained so as to minimize the error between the output layer and the input layer. If we use a short
+
+\subsection{Manually-generated features}
 \begin{table*}
@@
 Move features & Capturing move, recapturing move, en passant move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move.
+\hline
 \end{tabular} 
@@
 
-\subsection{Sparse Autoencoding}
+As a first step towards classification, we selected chess-specific manual features from a given position-move pair.
+The position-based features involve information such as where the pieces are on the board and potential moves, while the move-based features involve binary information such as if the move is a capturing move or a check.
+We list all our hand-selected manual features in Table REF.
 
@@
 
+TALK ABOUT THE AUTOENCODER EXPERIMENTS
+
 \subsection{Results}
@@
 
-We then extended the tagging to also incorporate information about the player
+We then extended the tagging to also incorporate information about the player’s color because there is a bias.
+For the predictions from the SVM model, this modification improved recall at a minor cost to precision, resulting in a higher F1 score.
+In contrast, the random predictions got much worse: Both the precision and recall scores fell as a result of this change, which adds further bias and reduces the amount of data per tag (since games are now spread over twice as many tags).
+
+TALK ABOUT AUTOENCODER RESULTS
+
+TALK ABOUT USING AUTOENCODER inputs for the SVM
 
@@
 \subsection{Classification Quality}
+Future work should emphasize the quality of classifications.
+We demonstrated that general approaches may not be effective enough to be useful for complex games like chess.
+Therefore, future projects should seek improvements by better taking advantage of the chess-specific tools and libraries that are available like game engines and weight functions.
+If the classifier incorporated chess-specific features like those generated by chess engines, it could be much better at identifying the types of strategies that are available to the player.
+For example, a deep traversal of many future paths by a chess engine may reveal that an attack is possible (attackPossible) or that the player can force a pawn promotion (pawnPromotion).
+While a purely computational approach to identifying opportunities and strategies may provide higher-quality or more reliable assessments, a statistical approach may be more robust.
+%talk about what i mean by robust
+
+One potential improvement that does not involve increased reliance on chess-specific technologies is to implement hierarchical tags so that matingAttack, doubleAttack, and similar tags imply a more general tag like attackPossible.
+By doing so, there will be a greater balance of positives and negatives, and labels that applied to games but were previously considered wrong are now considered correct.
+By incorporating hierarchical structure to tags, the system can jointly train and classify games.
+
+Another possible improvement is to use a different classifier than SVM, such as the perceptron or maximum entropy classifiers.
+One issue with the SVM is that often the tags are not so clear-cut: What does it mean for example that an endgame is “good”?
+Rather, there are degrees of how good an endgame is to the player.
+Therefore, a softer approach like maximum entropy may produce better results.
+
+\subsection{Generalization to Other Games.}
+One goal of this project was to develop a statistical machine learning approach that was general enough to be easily adapted to other applications like StarCraft or Checkers.
+A future work can explore this possibility by simply creating features for other applications.
 

### 2011-05-18 paper/paper.tex
removed the paper(copy) thing

@@
 
+\subsection{Preparing the Data}
+JINGHAO: Write about slicing and tag generation
+
 \section{Features}
-\subsection{Manually Generated Features}
-As a first step towards classification, we selected chess-specific manual features from a given position-move pair.
-The position-based features involve information such as where the pieces are on the board and potential moves, while the move-based features involve binary information such as if the move is a capturing move or a check.
-We list all our hand-selected manual features in Table REF.
+\subsection{Sparse auto-encoder}
+The sparse auto-encoder is one of many methods in use for unsupervised feature learning~\cite{single-layer-networks-unsupervised-feature-learning}.
+It takes the form of a feedforward neural network, which consists of an input layer of N nodes, some number of hidden layers, and output layer of M nodes.
+The value of each layer $\vec{l_i}$ is determined with a weight matrix $W_i$ and a bias vector $b_i$: \[L_i = f(W_i l_{i-1} + b_i),\] where $f(\cdot)$ is a non-linearity like the sigmoid $f(x) = \frac{1}{1 + e^{-z}}$, applied component-wise.
 
+We want to
+
+\subsection{Manually-generated features}
 \begin{table*}

### 2011-05-18 paper/paper.tex
Fixed refs.bib

@@
 Specifically, they compose many layers of non-linearities in order to compactly represent complex functions which can model these relationships.
-They have been used successfully for many different types of tasks \cite{ng-mnist,ng-vision,ng-nlp}.
+They have been used successfully for many different types of tasks \cite{Vincent,Lee,Socher}.
 
-\paragraph{Automated sportscasting.} Chen and Mooney (\cite{learning-to-sportscast}) present a commentator system that learns language from sportscasts of simulated RoboCup games.
+\paragraph{Automated sportscasting.} Chen and Mooney (\cite{chen}) present a commentator system that learns language from sportscasts of simulated RoboCup games.
 Their system, which is trained on ambiguous supervision in the form of textual human commentaries and simulation states of the soccer games, simultaneously tries to establish correspondences between the commentaries and the simulation states.
@@
 As a departure from the rule-based approach to automatic chess commentary, Guid et al.
-(\cite{learning-positional-features}) discuss the use of argument based machine learning, which analyzes reasoning where arguments for and against a certain claim are produced and evaluated (in the context of chess, why a certain move was played over others.) Their system, much like the one presented in this paper, identifies complex positional features of a given chess game.
+(\cite{Guid}) discuss the use of argument based machine learning, which analyzes reasoning where arguments for and against a certain claim are produced and evaluated (in the context of chess, why a certain move was played over others.) Their system, much like the one presented in this paper, identifies complex positional features of a given chess game.
 
@@
 \subsection{Preparing the Data}
-JINGHAO: Write about slicing and tag generation
+After parsing each game, we filtered out the tagged games that had more than 10 plies because it will be very difficult to identify the tags that apply to a complex sequence of moves.
+Among the games that we kept, we also kept only the first few positions and moves (the ``prefix'' of each game, as described earlier), a parameter that can be adjusted and is one by default.
+We then generate tags for each game based on the information from the source websites.
+For example, games in a file that corresponds to a lesson called ``Double Attack'' are labeled doubleAttack.
+In addition, some of those games may appear in other lessons, and consequently will have multiple labels.
+An additional step that we take during experimentation is to segment the games further by the color of the player in addition to the tags of the game in order to account for the white-bias in the training data.
 
@@
 \subsection{Sparse auto-encoder}
-The sparse auto-encoder is one of many methods in use for unsupervised feature learning~\cite{single-layer-networks-unsupervised-feature-learning}.
+The sparse auto-encoder is one of many methods in use for unsupervised feature learning~\cite{Coates}.
 It takes the form of a feedforward neural network, which consists of an input layer of N nodes, some number of hidden layers, and output layer of M nodes.
-The value of each layer $\vec{l_i}$ is determined with a weight matrix $W_i$ and a bias vector $b_i$: \[L_i = f(W_i l_{i-1} + b_i),\] where $f(\cdot)$ is a non-linearity like the sigmoid $f(x) = \frac{1}{1 + e^{-z}}$, applied component-wise.
+The value of each layer $\vec{l_i}$ is determined with a weight matrix $W_i$ and a bias vector $b_i$: \[L_i = f(W_i \cdot \vec{l_{i-1}} + b_i),\] where $f(\cdot)$ is a non-linearity like the sigmoid $f(x) = \frac{1}{1 + e^{-z}}$ applied component-wise.
 
-We want to
+An auto-encoder is such a neural network trained so as to minimize the error (specifically, the sum-squared difference) between the output layer and the input layer, over some set of input examples.
+If we have fewer nodes in the hidden layers than in the input or output layer, the neural network is forced to learn a compressed representation of the data (represented by the activations, or outputs, of the smaller number of hidden nodes).
+Instead, we keep a larger number of hidden nodes while imposing a sparsity constraint that the average activation of each one be small (e.g., 0.05).
+Other work has shown that such a constraint better forces the auto-encoder to learn interesting structures from the data~\cite{bengio}.
+
+A normal auto-encoder only has one layer of hidden nodes, but we can also add more hidden layers to form a stacked auto-encoder.
+Per~\cite{Bengio07greedylayer-wise}, we use a greedy strategy to train the extra hidden layers in the auto-encoder.
+We train an auto-encoder with one hidden layer as usual; once this completes, we obtain the output of the hidden layer over all the examples in the input set, and use that as the new set of examples.
+This allows us to learn higher-level representations of the data~\cite{Vincent}.
+
+Other strategies for unsupervised feature learning include restricted Boltzmann machines (which are undirected), deep belief nets (stacked RBMs, akin to stacked auto-encoders), and denoising autoencoders (which try to reconstruct some input from a corrupted version of it).
+We only evaluated sparse stacked auto-encoders in this paper.
 
@@
 \begin{table*}
+\centering
 \begin{tabular}{lp{0.6\textwidth}}
@@
 Position features & Piece location, potential moves per piece, material count, square control. \\
-Move features & Capturing move, recapturing move, en passant move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move.
+Move features & Capturing move, recapturing move, en passant move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move. \\
 \hline
@@
 
-As a first step towards classification, we selected chess-specific manual features from a given position-move pair.
+As a point of comparison to auto-encoders, we selected chess-specific manual features from a given position-move pair.
 The position-based features involve information such as where the pieces are on the board and potential moves, while the move-based features involve binary information such as if the move is a capturing move or a check.
@@
 \subsection{SVM Based Structured Classifier}
-
 In our experiments, we processed the input data into multimaps from game prefixes to tags.
@@
 
-TALK ABOUT AUTOENCODER RESULTS
+When we delved deeper to analyze the sources of errors and true positives, we arrived at a surprising result.
+We found that deflection and removalOfDefence are much more likely to cause false negatives than other tags (Table \ref{table:negatives}).
+In addition, we found that all of the true positives are for the quietMates tag.
+
+\begin{table*}
+\centering
+\begin{tabular}{lp{0.6\textwidth}}
+\hline
+\textbf{Category} & \textbf{Description} \\ \hline
+Position features & Piece location, potential moves per piece, material count, square control. \\
+Move features & Capturing move, recapturing move, en passant move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move. \\
+\hline
+\end{tabular} 
+\caption{Sources of error in the SVM classifier.}
+\label{table:negatives}
+\end{table*}
+
+Upon closer examination, it was clear that the “checkmate move” feature contributed to that: Any board that is in a checkmate situation is likely to be a quietMate.
+This interesting result demonstrates the limitations of pursuing a general approach for a highly-complex and specific game like chess.
+
+We sought to improve upon this baseline with an autoencoder. TALK ABOUT AUTOENCODER RESULTS
 
@@
 
+\bibliographystyle{acl}
+\bibliography{paper}
+
 \end{document}
-
-stack of work:
-Segmented:
-True positive: 13 (0.001252)
-False positive: 2 (0.000193)
-True negative: 10213 (0.983438)
-False negative: 157 (0.015118)
-Precision/Recall: 0.866667, 0.076471
-F1 Score: 0.140541
-True positive: 12841 (0.001236)
-False positive: 154930 (0.014919)
-True negative: 10060070 (0.968712)
-False negative: 157159 (0.015133)
-Precision/Recall: 0.076539, 0.075535
-F1 Score: 0.076034
-
-Unsegmented:
-True positive: 11 (0.002087)
-False positive: 0 (0.000000)
-True negative: 5106 (0.968880)
-False negative: 153 (0.029032)
-Precision/Recall: 1.000000, 0.067073
-F1 Score: 0.125714
-True positive: 22387 (0.004248)
-False positive: 145688 (0.027645)
-True negative: 4960312 (0.941236)
-False negative: 141613 (0.026872)
-Precision/Recall: 0.133196, 0.136506
-F1 Score: 0.134831
-
-\begin{table*}
-\begin{tabular}{lp{0.6\textwidth}}
-\hline
-\textbf{True Positive} & \textbf{False Positive} & \textbf{True Negative} & \textbf{False Negative} & \textbf{Precision} & \textbf{Recall} & \textbf{F1} \\ \hline
-Position features & Piece location, potential moves per piece, material count, square control. \\
-Move features & Capturing move, recapturing move, en passant move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move.
-\end{tabular} 
-\caption{List of chess-specific manual features per position.}
-\end{table*}

### 2011-05-18 paper/paper.tex
added svmlight cite

@@
 \begin{abstract}
-We investigate applications of statistical learning to automated sportscasting for chess. In particular, we 
+We investigate applications of statistical learning to automated sportscasting for chess.
+In particular, this project studies methods to identify high level structures in chess games using support vector machines and autoencoders.
 \end{abstract}
@@
 Even after years of play, each game is new challenge, a new problem to solve.
-Historically, chess has been prominent as a sport where the battles of grandmasters engulf thousands; given the highly intricate nature of chess (and similar abstract strategy games), experts have scrutinized these games to devise a standard high-level structures like openings, strategies, tactics, and time advantage, which they use to explain maneuvers and blunders that occur in the games.
+Historically, chess has been prominent as a sport where the battles of grandmasters engulf thousands; given the highly intricate nature of chess (and similar abstract strategy games), experts have scrutinized these games to devise standard high-level structures like openings, strategies, tactics, and time advantage, which they use to explain maneuvers and blunders that occur in the games.
 Devotees of chess also attempt to improve their own performance in terms of these high-level structures, reading chess theorists' comments about games by high-ranking players.
@@
 
-\paragraph{Automated sportscasting.} Chen and Mooney (\cite{chen}) present a commentator system that learns language from sportscasts of simulated RoboCup games.
+\paragraph{Automated sportscasting.} Chen and Mooney~\shortcite{chen} present a commentator system that learns language from sportscasts of simulated RoboCup games.
 Their system, which is trained on ambiguous supervision in the form of textual human commentaries and simulation states of the soccer games, simultaneously tries to establish correspondences between the commentaries and the simulation states.
@@
 \paragraph{Automatic chess commentary.} There exist techniques for automatic annotation of chess games which are heavily dependent on rule-based logic and use of a chess-engine to determine which parts of a specific chess game are salient~\cite{cambridge-chess-annotation}.
-As a departure from the rule-based approach to automatic chess commentary, Guid et al.
-(\cite{Guid}) discuss the use of argument based machine learning, which analyzes reasoning where arguments for and against a certain claim are produced and evaluated (in the context of chess, why a certain move was played over others.) Their system, much like the one presented in this paper, identifies complex positional features of a given chess game.
+As a departure from the rule-based approach to automatic chess commentary, Guid et al.~\shortcite{Guid} discuss the use of argument based machine learning, which analyzes reasoning where arguments for and against a certain claim are produced and evaluated (in the context of chess, why a certain move was played over others.) Their system, much like the one presented in this paper, identifies complex positional features of a given chess game.
 
@@
 A large number of games as well as software and tools that use the PGN format are freely available.
-
-The already-annotated games come from databases of short snippets of chess games,\footnote{\url{http://wwwu.uni-klu.ac.at/gossimit/c/tactic.htm}} \footnote{\url{http://webplaza.pt.lu/public/ckaber/Chess.htm#Chess training}} intended for chess players to learn from by example, which are organized by particular themes that the students are to identify.
-These provided a useful repository of both interesting high-level concepts we could identify (we provide a listing in Table~\ref{table:themes}), as well as boards and moves corresponding to each one.
-
-We also looked for as many chess games as possible from a variety of web sites, such as \url{chessgames.com}.
-We easily obtained over five million untagged games spread over 600,000 files and 5 GB of data.
-The number of annotated game snippets we obtained was slightly over 5,000.
-
-\subsection{Data Pre-Processing}	
-In order to ensure consistency and integrity over the games obtained from disparate sources, we sanitized them in multiple ways.
-We parsed each game in the data set and removed information that would confuse the parser.
-For example, some games had English or German comments by humans which we did not plan to use, and sometimes resulted in parsing errors due to specific characters present in the comments.
-Others had alternate branches, sometimes up to multiple levels (alternate branches within alternate branches), which the parser could not handle properly.
-Still others represented games in a non-standard way (for example, pawn promotions were represented without the ``=''), which we had to fix.
-
-Since we obtained the games from many sources, there was significant overlap in games (approximately 50% of the unlabeled games collected were redundant).
-Therefore, we also went through all of the downloaded games and eliminated duplicate games: multiple games that had the same initial position and move sequence were considered redundant and only one of each were kept.
-
-\subsection{Parsing}
-We tried various open source PGN parsers, each succumbing to flaws like mishandling uncommon cases (such as  promotions and special moves like \emph{en passant}).
-We based our system on Chesspresso,\footnote{url{http://www.chesspresso.org/}} a Java library for writing chess programs, which we fixed and modified for our needs.
-Chesspresso takes a PGN file as input and returns a data structure that represents the initial position of the board and the subsequent moves (as well as all metadata necessary to recreate the PGN file).
-We also extended it to be more robust to variations in input format, in addition to being able to export PGNs in a canonical form, which was crucial for removing duplicates.
-
-\subsection{Preparing the Data}
-After parsing each game, we filtered out the tagged games that had more than 10 plies because it will be very difficult to identify the tags that apply to a complex sequence of moves.
-Among the games that we kept, we also kept only the first few positions and moves (the ``prefix'' of each game, as described earlier), a parameter that can be adjusted and is one by default.
-We then generate tags for each game based on the information from the source websites.
-For example, games in a file that corresponds to a lesson called ``Double Attack'' are labeled doubleAttack.
-In addition, some of those games may appear in other lessons, and consequently will have multiple labels.
-An additional step that we take during experimentation is to segment the games further by the color of the player in addition to the tags of the game in order to account for the white-bias in the training data.
-
-\section{Features}
-\subsection{Sparse auto-encoder}
-The sparse auto-encoder is one of many methods in use for unsupervised feature learning~\cite{Coates}.
-It takes the form of a feedforward neural network, which consists of an input layer of N nodes, some number of hidden layers, and output layer of M nodes.
-The value of each layer $\vec{l_i}$ is determined with a weight matrix $W_i$ and a bias vector $b_i$: \[L_i = f(W_i \cdot \vec{l_{i-1}} + b_i),\] where $f(\cdot)$ is a non-linearity like the sigmoid $f(x) = \frac{1}{1 + e^{-z}}$ applied component-wise.
-
-An auto-encoder is such a neural network trained so as to minimize the error (specifically, the sum-squared difference) between the output layer and the input layer, over some set of input examples.
-If we have fewer nodes in the hidden layers than in the input or output layer, the neural network is forced to learn a compressed representation of the data (represented by the activations, or outputs, of the smaller number of hidden nodes).
-Instead, we keep a larger number of hidden nodes while imposing a sparsity constraint that the average activation of each one be small (e.g., 0.05).
-Other work has shown that such a constraint better forces the auto-encoder to learn interesting structures from the data~\cite{bengio}.
-
-A normal auto-encoder only has one layer of hidden nodes, but we can also add more hidden layers to form a stacked auto-encoder.
-Per~\cite{Bengio07greedylayer-wise}, we use a greedy strategy to train the extra hidden layers in the auto-encoder.
-We train an auto-encoder with one hidden layer as usual; once this completes, we obtain the output of the hidden layer over all the examples in the input set, and use that as the new set of examples.
-This allows us to learn higher-level representations of the data~\cite{Vincent}.
-
-Other strategies for unsupervised feature learning include restricted Boltzmann machines (which are undirected), deep belief nets (stacked RBMs, akin to stacked auto-encoders), and denoising autoencoders (which try to reconstruct some input from a corrupted version of it).
-We only evaluated sparse stacked auto-encoders in this paper.
-
-\subsection{Manually-generated features}
-\begin{table*}
-\centering
-\begin{tabular}{lp{0.6\textwidth}}
-\hline
-\textbf{Category} & \textbf{Description} \\ \hline
-Position features & Piece location, potential moves per piece, material count, square control. \\
-Move features & Capturing move, recapturing move, en passant move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move. \\
-\hline
-\end{tabular} 
-\caption{List of chess-specific manual features per position.}
-\end{table*}
-
-As a point of comparison to auto-encoders, we selected chess-specific manual features from a given position-move pair.
-The position-based features involve information such as where the pieces are on the board and potential moves, while the move-based features involve binary information such as if the move is a capturing move or a check.
-We list all our hand-selected manual features in Table REF.
-
-\section{Classification}
-We built a support vector machine for structured classification of games.
-We filtered our tagged games by the number of plies (number of individual moves), generated mappings between tags and game prefixes (some initial position plus a sequence of moves), segmented our tagged games into three groups (training, validation and test), and called SVMLight with that data, an external library for SVM.
-
-\subsection{High-level descriptions}
-We classify segments of the chess game based on various high-level descriptions from chess theory and folklore:
 
@@
 
-\section{Implementation} % Parallelization and Scaling
-Running serially, the autoencoder takes a long time to complete, even on simple inputs.
-Therefore, we examined ways to parallelize the task in order to fully leverage the capacity of our 8-core server with hyper-threading (16 virtual cores).
-We implemented a local version of MapReduce that creates 16 execution threads.
+The already-annotated games come from databases of short snippets of chess games,\footnote{\url{http://wwwu.uni-klu.ac.at/gossimit/c/tactic.htm}} \footnote{\url{http://webplaza.pt.lu/public/ckaber/Chess.htm#Chess training}} intended for chess players to learn from by example, which are organized by particular themes that the students are to identify.
+These provided a useful repository of both interesting high-level concepts we could identify (we provide a listing in Table~\ref{table:themes}), as well as boards and moves corresponding to each one.
 
-\section{Experiments}
-\subsection{SVM Based Structured Classifier}
-In our experiments, we processed the input data into multimaps from game prefixes to tags.
-We then segmented the data into three disjoint sets: 90% for training, 5% as the validation set (for tuning our classifiers), and 5% as the test set.
-The segmenting was done by randomly sampling from the set of all labeled games.
+We also looked for as many chess games as possible from a variety of web sites, such as \url{chessgames.com}.
+We easily obtained over five million untagged games spread over 600,000 files and 5 GB of data.
+The number of annotated game snippets we obtained was slightly over 5,000.
 
-Afterwards, we constructed lists of positive and negative labels (mappings from tags to games that have that tag, and games that do not have that tag).
-We then built SVM models using the training data.
-In testing, we fed that to the SVM to generate, for each test game and tag, a boolean prediction of whether that tag applies to that game.
-As a result, we constructed a SVM-based multiple-class classifier that attempts to predict the tags that characterize an input game.
+\subsection{Data Pre-Processing}	
+In order to ensure consistency and integrity over the games obtained from disparate sources, we sanitized them in multiple ways.
+We parsed each game in the data set and removed information that would confuse the parser.
+For example, some games had English or German comments by humans which we did not plan to use, and sometimes resulted in parsing errors due to specific characters present in the comments.
+Others had alternate branches, sometimes up to multiple levels (alternate branches within alternate branches), which the parser could not handle properly.
+Still others represented games in a non-standard way (for example, pawn promotions were represented without the ``=''), which we had to fix.
 
-We measured the accuracy of our model using various measures like false positive rate, false negative rate, precision, recall, and F1.
+Since we obtained the games from many sources, there was significant overlap in games (approximately 50% of the unlabeled games collected were redundant).
+Therefore, we also went through all of the downloaded games and eliminated duplicate games: multiple games that had the same initial position and move sequence were considered redundant and only one of each were kept.
+
+\subsection{Parsing}
+We tried various open source PGN parsers, each succumbing to flaws like mishandling uncommon cases (such as  promotions and special moves like \emph{en passant}).
+We based our system on Chesspresso,\footnote{\url{http://www.chesspresso.org/}} a Java library for writing chess programs, which we fixed and modified for our needs.
+Chesspresso takes a PGN file as input and returns a data structure that represents the initial position of the board and the subsequent moves (as well as all metadata necessary to recreate the PGN file).
+We also extended it to be more robust to variations in input format, in addition to being able to export PGNs in a canonical form, which was crucial for removing duplicates.
+
+\subsection{Preparing the Data}
+After parsing each game, we filtered out the tagged games that had more than 10 plies because it will be very difficult to identify the tags that apply to a complex sequence of moves.
+Among the games that we kept, we also kept only the first few positions and moves (the ``prefix'' of each game, as described earlier), a parameter that can be adjusted; for our experiments, we kept just the initial board specified in each game.
+We then generate labels for each game based on the information from the source websites.
+For example, games in a file that corresponds to a lesson called ``Double Attack'' are labeled doubleAttack.
+In addition, some of those games may appear in other lessons, and consequently will have multiple labels.
+We also segment the games further by the color of the player that makes the first move in the game, in addition to the tags of the game in order to account for the white-bias in the training data.
+
+\section{Feature Extraction}
+\subsection{Sparse auto-encoder}
+The sparse auto-encoder is one of many methods in use for unsupervised feature learning~\cite{Coates}.
+It takes the form of a feedforward neural network, which consists of an input layer of N nodes, some number of hidden layers, and output layer of M nodes.
+The value of each layer $\vec{l_i}$ is determined with a weight matrix $W_i$ and a bias vector $b_i$: \[L_i = f(W_i \cdot \vec{l_{i-1}} + b_i),\] where $f(\cdot)$ is a non-linearity like the sigmoid $f(x) = \frac{1}{1 + e^{-z}}$ applied component-wise.
+
+An auto-encoder is such a neural network trained so as to minimize the error (specifically, the sum-squared difference) between the output layer and the input layer, over some set of input examples.
+If we have fewer nodes in the hidden layers than in the input or output layer, the neural network is forced to learn a compressed representation of the data (represented by the activations, or outputs, of the smaller number of hidden nodes).
+Instead, we keep a larger number of hidden nodes while imposing a sparsity constraint that the average activation of each one be small (e.g., 0.05).
+Other work has shown that such a constraint better forces the auto-encoder to learn interesting structures from the data~\cite{bengio}.
+
+A normal auto-encoder only has one layer of hidden nodes, but we can also add more hidden layers to form a stacked auto-encoder.
+Per Bengio et al.~\shortcite{Bengio07greedylayer-wise}, we use a greedy strategy to train the extra hidden layers in the auto-encoder.
+We train an auto-encoder with one hidden layer as usual; once this completes, we obtain the output of the hidden layer over all the examples in the input set, and use that as the new set of examples.
+This allows us to learn higher-level representations of the data~\cite{Vincent}.
+
+Other strategies for unsupervised feature learning include restricted Boltzmann machines (which are undirected), deep belief nets (stacked RBMs, akin to stacked auto-encoders), and denoising autoencoders (which try to reconstruct some input from a corrupted version of it).
+We only evaluated sparse stacked auto-encoders in this paper.
+
+\subsection{Manually-generated features}
+\begin{table*}
+\centering
+\begin{tabular}{lp{0.6\textwidth}}
+\hline
+\textbf{Category} & \textbf{Description} \\ \hline
+Position & Piece location, potential moves per piece, material count, square control. \\
+Previous move & Capturing move, recapturing move, en passant move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move. \\
+\hline
+\end{tabular} 
+\caption{List of chess-specific manual features per position.}
+\label{table:features}
+\end{table*}
+
+As a point of comparison to auto-encoders, we extracted chess-specific features manually from a given chess position (configuration of pieces on the board) and the move that resulted in that position; Table~\ref{table:features} contains a list.
+The position-based features include information such as where the pieces are on the board and potential moves, while the move-based features involve binary information such as if the move that resulted in this position was a capturing move or resulted in a check.
+
+\section{Experimental methodology}
+We processed all of the labeled training data to obtain a mapping from each unique game prefix (for most of the experiments, just the initial specified configuration of the board in the labeled game) to the labels that apply to it; most of the prefixes only had one label, while others had two or three.
+We then segmented the data into three disjoint sets: 90% for training, 5% as the validation set (for use with tuning our classifiers), and 5% as the test set.
+The segmenting was done by random selection from the set of all labeled games.
+
+In our experiments, we first use the labeled data in the training set (and possibly also unlabeled data) to train the classifiers.
+Then, for every test input chess game (or a prefix of a game), we use our classifiers to apply some number of labels to it that come from Table~\ref{table:themes}.
+We measure the performance of the predictions made by our classifiers by counting the number of true positives (a predicted label was among the set of gold labels for a test input), true negatives (a label in Table~\ref{table:themes} was neither in the set of predicted labels nor the set of gold labels), false positives (a predicted label was not in the set of gold labels), and false negatives (a gold label was not predicted).
+We compute and report the precision, recall, and F1 score using these measures of accuracy.
+
+\subsection{Support vector machine classifier}
+We note that any number of labels may apply to some test input that we want to classify; a straightforward use of a binary (or even multiclass) classifier, as provided by most implementations of support vector machines, will not suffice as it can only assign one label.
+Therefore, we take the one-versus-all strategy of training many binary SVMs; for each label that we consider, we train a binary classifier with all training games that have the given label as positive examples, and all other games as negative examples.
+We used SVMlight~\cite{svmlight}, a popular third-party implementation of SVMs, for our experiments, using the manually-generated features extracted from each labeled data point.
+
+\subsection{Classifying with the auto-encoder}
+There exists several methods for using the auto-encoder for the task at hand.
+First, we can consider using the output of the last hidden layer as the feature vector for input into the SVM classifier, just as we did with the manually-generated features.
+To do this,
+
+\subsection{Random baseline}
 In addition, we compared the predictions made by SVM to those made randomly where each game was assigned a tag at random based on the proportion of the games that have that tag.
 The random predictions were made 1,000 times, with the predictions made independently at random, and the error rates were averaged.
+
+\section{Experimental results}
+\subsection{SVM Based Structured Classifier}
+We built a support vector machine for structured classification of games.
+We filtered our tagged games by the number of plies (number of individual moves), generated mappings between tags and game prefixes (some initial position plus a sequence of moves), segmented our tagged games into three groups (training, validation and test), and called SVMLight with that data, an external library for SVM.
 
@@
 
-\begin{table*}
-\centering
-\begin{tabular}{lp{0.6\textwidth}}
+\begin{table}
+\begin{tabular}{lr}
 \hline
-\textbf{Category} & \textbf{Description} \\ \hline
-Position features & Piece location, potential moves per piece, material count, square control. \\
-Move features & Capturing move, recapturing move, en passant move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move. \\
-\hline
+\textbf{Tag} & \textbf{Error rate} \\ \hline
+\tt{queenCapture}  & 0.006452 \\
+\tt{kingInCenter}  & 0.006452 \\
+\tt{perpetualCheck}  & 0.006452 \\
+\tt{drawCombinationsStalemate}  & 0.038710 \\
+\tt{bishopPair}  & 0.006452 \\
+\tt{doubleAttack}  & 0.006452 \\
+\tt{clearance}  & 0.025806 \\
+\tt{openingClosingDiagonals}  & 0.006452 \\
+\tt{openingClosingFiles}  & 0.006452 \\
+\tt{enclosedKings}  & 0.012903 \\
+\tt{discoveredAttack}  & 0.025806 \\
+\tt{passedPawn}  & 0.012903 \\
+\tt{quietMates}  & 0.038710 \\
+\tt{pin}  & 0.051613 \\
+\tt{removalOfDefence}  & 0.225806 \\
+\tt{deflection}  & 0.251613 \\
+\tt{decoy}  & 0.096774 \\
+\tt{interference}  & 0.038710 \\
+\tt{backrankWeakness}  & 0.032258 \\
+\tt{freeingBlockingSquares}  & 0.019355 \\
+\tt{pawnPromotion}  & 0.064516 \\
+\tt{pullingPiecesToBadSquares}  & 0.006452
 \end{tabular} 
-\caption{Sources of error in the SVM classifier.}
+\caption{False negative and false positive rates for various tags.}
 \label{table:negatives}
-\end{table*}
+\end{table}
 
@@
 
-TALK ABOUT USING AUTOENCODER inputs for the SVM
+\section{Implementation} % Parallelization and Scaling
+Running serially, the autoencoder takes a long time to complete, even on simple inputs.
+Therefore, we examined ways to parallelize the task in order to fully leverage the capacity of our 8-core server with hyper-threading (16 virtual cores).
+We implemented a local version of MapReduce that creates 16 execution threads.
 
@@
 \subsection{Classification Quality}
-Future work should emphasize the quality of classifications.
+Future work should emphasize the classification quality.
 We demonstrated that general approaches may not be effective enough to be useful for complex games like chess.
@@
 
-Another possible improvement is to use a different classifier than SVM, such as the perceptron or maximum entropy classifiers.
-One issue with the SVM is that often the tags are not so clear-cut: What does it mean for example that an endgame is “good”?
+Another possible improvement is to use a different classifier than SVM, such as the perceptron or a maximum entropy classifier.
+One issue with the SVM is that often the tags are not so clear-cut: what does it mean that an endgame has “good chances”?
 Rather, there are degrees of how good an endgame is to the player.
@@
 
-\subsection{Generalization to Other Games.}
+\subsection{Generalization to Other Games}
 One goal of this project was to develop a statistical machine learning approach that was general enough to be easily adapted to other applications like StarCraft or Checkers.

### 2011-05-18 paper/paper.tex
Updated ClassifyLabeledData to identify false positives/negatives; + paper tex and pdf

@@
   Charles Chen \\
-  {\tt charleschen@berkeley.edu} \\
+	UC Berkeley \\
+	Berkeley, CA 94720 \\
+  {\tt \small{charleschen@berkeley.edu}} \\
 \And
   Richard Shin \\
-  {\tt ricshin@berkeley.edu} \\
+	UC Berkeley \\
+	Berkeley, CA 94720 \\
+  {\tt \small{ricshin@berkeley.edu}} \\
 \And
   Jinghao Yan \\
-  {\tt jinghao@berkeley.edu} \\
+	UC Berkeley \\
+	Berkeley, CA 94720 \\
+  {\tt \small{jinghao@berkeley.edu}} \\
 }
-\date{}
 
@@
 We investigate applications of statistical learning to automated sportscasting for chess.
-In particular, this project studies methods to identify high level structures in chess games using support vector machines and autoencoders.
+In particular, this project studies methods to automatically identify high level structures in chess games using support vector machines and auto-encoders.
+We then compare the results of the two different approaches and showed improvements over the SVM model using a stacked auto-encoder.
+Finally, we conclude with results, analysis of errors, and ideas for future work.
 \end{abstract}
@@
 Their system, which is trained on ambiguous supervision in the form of textual human commentaries and simulation states of the soccer games, simultaneously tries to establish correspondences between the commentaries and the simulation states.
-Their work focuses on generating meaningful English commentary from relatively simple topics such as ‘Player 1 kicked the ball to Player 2’, while our work focuses more on identifying complex topics.
+Their work focuses on generating meaningful English commentary from relatively simple topics such as ``Player 1 kicked the ball to Player 2'', while our work focuses less on text generation and more on identifying complex topics.
 
@@
 \tt{interruption} & There exists a maneuver that interrupts the opponent’s tactics. \\
-\tt{kingInCenter} & There exists a decisive strategic advantage with the player with the King in the center of the board. \\
+\tt{kingInCenter} & There exists a decisive strategic advantage for the player with the King in the center of the board. \\
 \tt{kingToBadSquares} & There exists a combination that forces the opposing King to move to a bad square. \\
-\tt{longDiagonals} & There exists a strategic advantage to the player that controls squares on the long diagonal. \\
+\tt{longDiagonals} & There exists a strategic advantage for the player that controls squares on the long diagonal. \\
 \tt{matingAttack} & There exists a decisive mating attack. \\
@@
 \tt{overload} & There exists a combination that makes a defending piece defend too many squares at once. \\
-\tt{passedPawn} & There exists a combination that allows for a passed pawn (a pawn with no opposing pawns in front of it) \\
+\tt{passedPawn} & There exists a combination that allows for a passed pawn (a pawn with no opposing pawns in front of it.) \\
 \tt{pawnPromotion} & There exists a combination that allows for a pawn promotion. \\
@@
 \tt{unpinning} & There exists a maneuver to unpin a piece. \\
-\tt{wrongBishop} & The opposing player’s only bishop is limited in use because of the pawn structure of the game. \\ \hline
+\tt{wrongBishop} & The opponent’s only bishop is limited in use because of the pawn structure of the game. \\ \hline
 \end{tabular}
@@
 
-Since we obtained the games from many sources, there was significant overlap in games (approximately 50% of the unlabeled games collected were redundant).
+Since we obtained the games from many sources, there was significant overlap in games (approximately 50\% of the unlabeled games collected were redundant).
 Therefore, we also went through all of the downloaded games and eliminated duplicate games: multiple games that had the same initial position and move sequence were considered redundant and only one of each were kept.
@@
 We then generate labels for each game based on the information from the source websites.
-For example, games in a file that corresponds to a lesson called ``Double Attack'' are labeled doubleAttack.
+For example, games in a file that corresponds to a lesson called ``Double Attack'' are labeled {\tt doubleAttack}.
 In addition, some of those games may appear in other lessons, and consequently will have multiple labels.
-We also segment the games further by the color of the player that makes the first move in the game, in addition to the tags of the game in order to account for the white-bias in the training data.
+We also segment the games further by the color of the player that makes the first move in the game, in addition to the labels of the game, since most of the labels in Table~\ref{table:themes} only refer to a specific player; for instance, some of the games labeled with {\tt backrankWeakness} may have the white player with the weakness, while others show the black player with the weakness.
 
@@
 
-Other strategies for unsupervised feature learning include restricted Boltzmann machines (which are undirected), deep belief nets (stacked RBMs, akin to stacked auto-encoders), and denoising autoencoders (which try to reconstruct some input from a corrupted version of it).
+Other strategies for unsupervised feature learning include restricted Boltzmann machines (which are undirected), deep belief nets (stacked RBMs, akin to stacked auto-encoders), and denoising auto-encoders (which try to reconstruct some input from a corrupted version of it).
 We only evaluated sparse stacked auto-encoders in this paper.
+
+The auto-encoder requires training data in the form of vectors where all the elements are between 0 and 1, or the range of the sigmoid function.
+We generated the training vectors by, for some chess board and for every type of piece (knight, rook, etc.), whether a white one (valued 1.0), a black one (valued 0.0), or none at all (valued 0.5) was present in every square of the chessboard.
 
@@
 The position-based features include information such as where the pieces are on the board and potential moves, while the move-based features involve binary information such as if the move that resulted in this position was a capturing move or resulted in a check.
+The resulting feature vector was then L2-normalized to fit the specifications of the support vector machine.
 
@@
 We processed all of the labeled training data to obtain a mapping from each unique game prefix (for most of the experiments, just the initial specified configuration of the board in the labeled game) to the labels that apply to it; most of the prefixes only had one label, while others had two or three.
-We then segmented the data into three disjoint sets: 90% for training, 5% as the validation set (for use with tuning our classifiers), and 5% as the test set.
+We then segmented the data into three disjoint sets: 90\% for training, 5\% as the validation set (for use with tuning our classifiers), and 5\% as the test set.
 The segmenting was done by random selection from the set of all labeled games.
@@
 \subsection{Classifying with the auto-encoder}
+\label{sec:classifying-auto-encoder}
 There exists several methods for using the auto-encoder for the task at hand.
 First, we can consider using the output of the last hidden layer as the feature vector for input into the SVM classifier, just as we did with the manually-generated features.
-To do this,
+To do this, we first perform unsupervised fine-tuning; we combine all the greedily-trained auto-encoders to make a many-layered neural network intended to reconstruct the input layer in its output layer, after which we again minimize the sum-squared error between the input and the output layers.
+We can then use the output of the hidden layer (i.e., not the one that reconstructs the input, but the middle layer which provides the most high-level representation of the input) as the feature vector.
+
+We can also use semi-supervised learning methods, in which after greedily training the auto-encoders, we combine all the hidden layers of the auto-encoders and add an output layer at the end.
+For every labeled input game, we construct input-output vectors such that the output encodes the gold labels of the input.
+We then train the resulting multi-layer neural network to minimize the sum-squared error; this acts similarly to a multiclass logistic regression, except with a different objective at the last layer.
+
+To train these neural networks, we used batch learning over all the examples at once.
+The loss function contained components for the sum-squared error as well as an L2 regularization term and a sparsity penalty based on the Kullback-Leibler divergence between the desired average activation of the hidden nodes and the actual average activation.
+We used backpropagation to compute the gradient of the loss function and L-BFGS to optimize the parameters as to minimize the loss function~\cite{ng-lecturenotes}.
+Running serially, the training takes a long time to complete, even on simple inputs.
+Therefore, we examined ways to parallelize the task in order to fully leverage the capacity of modern many-core machines; we achieved a significant speedup by parallelizing the loss function evaluation to 16 execution threads.
 
 \subsection{Random baseline}
-In addition, we compared the predictions made by SVM to those made randomly where each game was assigned a tag at random based on the proportion of the games that have that tag.
+To use as a baseline and ensure that our classifiers are giving sensible results, we randomly made predictions for every test game, where each one assigned each possible label at random, with the possibility of the label being given based on the proportion of the training games which have that label.
 The random predictions were made 1,000 times, with the predictions made independently at random, and the error rates were averaged.
@@
 \section{Experimental results}
-\subsection{SVM Based Structured Classifier}
-We built a support vector machine for structured classification of games.
-We filtered our tagged games by the number of plies (number of individual moves), generated mappings between tags and game prefixes (some initial position plus a sequence of moves), segmented our tagged games into three groups (training, validation and test), and called SVMLight with that data, an external library for SVM.
+\subsection{SVM classifier}
+\begin{table*}
+\centering
+\begin{tabular}{lccccccc}
+\hline
+\textbf{Type} & \textbf{TP} & \textbf{FP} & \textbf{TN} & \textbf{FN} & \textbf{Precision} & \textbf{Recall} & \textbf{F1} \\ \hline
+Unsegmented & 0.21\% & 0.00\% & 96.89\% & 2.90\% & 100\% & 6.71\% & 12.57\% \\
+Baseline (Random) & 0.42\% & 2.76\% & 94.12\% & 2.69\% & 13.32\% & 13.65\% & 13.48\% \\
+\hline
+Segmented & 0.13\% & 0.02\% & 98.34\% & 1.51\% & 86.67\% & 7.65\% & 14.05\% \\
+Baseline (Random) & 0.12\% & 1.49\% & 96.87\% & 1.51\% & 7.65\% & 7.55\% & 7.60\% \\
+\end{tabular} 
+\caption{Classification results for the games that have not been segmented by the color of the active player, and those that have been.
+In the former, precision clearly outperforms a random baseline, but recall falls short, resulting in a lower F1 score.
+In the latter, precision has fallen relative to the unsegmented version, but recall has risen, resulting in a slightly higher F1 score.
+In contrast, the random model became much worse in both ways.}
+\label{table:svmResults}
+\end{table*}
 
-TALK ABOUT THE AUTOENCODER EXPERIMENTS
+We first describe the results achieved when we trained using the manually-specified feature extractors.
+Since each game has very few labels, most of the binary classification results are negative.
+Therefore, the model is heavily biased toward negative results, as shown in Table~\ref{table:svmResults}.
 
-\subsection{Results}
-Since each game has very few tags, most of the binary classifications are negative. As a result, the model is heavily biased toward negative results. Our results bear this disposition:
-
-INSERT TABLE HERE
-
-The basic classifier generates classifications with relatively high precision compared to a random classifier.
+This classification method had relatively high precision compared to the random baseline.
 However, the recall is much lower than that of the random classifier.
@@
 
-We then extended the tagging to also incorporate information about the player’s color because there is a bias.
-For the predictions from the SVM model, this modification improved recall at a minor cost to precision, resulting in a higher F1 score.
-In contrast, the random predictions got much worse: Both the precision and recall scores fell as a result of this change, which adds further bias and reduces the amount of data per tag (since games are now spread over twice as many tags).
-
-When we delved deeper to analyze the sources of errors and true positives, we arrived at a surprising result.
-We found that deflection and removalOfDefence are much more likely to cause false negatives than other tags (Table \ref{table:negatives}).
-In addition, we found that all of the true positives are for the quietMates tag.
-
+When we extended the labeling to also incorporate the first mover's color information, the SVM model's recall was improved at a minor cost to precision, resulting in a higher F1 score.
+In contrast, the random predictions became much worse: Both the precision and recall scores fell as a result of this change, which reduced the evenness of the label distribution among the training games, and reduced the amount of games with each label (since games are now spread over twice as many tags).
 \begin{table}
+\centering
 \begin{tabular}{lr}
@@
 \textbf{Tag} & \textbf{Error rate} \\ \hline
-\tt{queenCapture}  & 0.006452 \\
-\tt{kingInCenter}  & 0.006452 \\
-\tt{perpetualCheck}  & 0.006452 \\
-\tt{drawCombinationsStalemate}  & 0.038710 \\
-\tt{bishopPair}  & 0.006452 \\
-\tt{doubleAttack}  & 0.006452 \\
-\tt{clearance}  & 0.025806 \\
-\tt{openingClosingDiagonals}  & 0.006452 \\
-\tt{openingClosingFiles}  & 0.006452 \\
-\tt{enclosedKings}  & 0.012903 \\
-\tt{discoveredAttack}  & 0.025806 \\
-\tt{passedPawn}  & 0.012903 \\
-\tt{quietMates}  & 0.038710 \\
-\tt{pin}  & 0.051613 \\
-\tt{removalOfDefence}  & 0.225806 \\
-\tt{deflection}  & 0.251613 \\
-\tt{decoy}  & 0.096774 \\
-\tt{interference}  & 0.038710 \\
-\tt{backrankWeakness}  & 0.032258 \\
-\tt{freeingBlockingSquares}  & 0.019355 \\
-\tt{pawnPromotion}  & 0.064516 \\
-\tt{pullingPiecesToBadSquares}  & 0.006452
+\tt{queenCapture}  & 0.6452\% \\
+\tt{kingInCenter}  & 0.6452\% \\
+\tt{perpetualCheck}  & 0.6452\% \\
+\tt{drawCombinationsStalemate}  & 3.8710\% \\
+\tt{bishopPair}  & 0.6452\% \\
+\tt{doubleAttack}  & 0.6452\% \\
+\tt{clearance}  & 2.5806\% \\
+\tt{openingClosingDiagonals}  & 0.6452\% \\
+\tt{openingClosingFiles}  & 0.6452\% \\
+\tt{enclosedKings}  & 1.2903\% \\
+\tt{discoveredAttack}  & 2.5806\% \\
+\tt{passedPawn}  & 1.2903\% \\
+\tt{quietMates}  & 3.8710\% \\
+\tt{pin}  & 5.1613\% \\
+\tt{removalOfDefence}  & 22.5806\% \\
+\tt{deflection}  & 25.1613\% \\
+\tt{decoy}  & 9.6774\% \\
+\tt{interference}  & 3.8710\% \\
+\tt{backrankWeakness}  & 3.2258\% \\
+\tt{freeingBlockingSquares}  & 1.9355\% \\
+\tt{pawnPromotion}  & 6.4516\% \\
+\tt{pullingPiecesToBadSquares}  & 0.6452\%
 \end{tabular} 
-\caption{False negative and false positive rates for various tags.}
+\caption{Percent of incorrect classification (false negatives and false positives) for various tags.}
 \label{table:negatives}
@@
 
-Upon closer examination, it was clear that the “checkmate move” feature contributed to that: Any board that is in a checkmate situation is likely to be a quietMate.
-This interesting result demonstrates the limitations of pursuing a general approach for a highly-complex and specific game like chess.
+Given the high precision but low recall, we delved deeper to analyze the sources of errors and true positives.
+As we show in Table~\ref{table:negatives}, we found that {\tt deflection} and {\tt removalOfDefence} are much more likely to cause false negatives than other tags.
+In addition, we found that all of the true positives are for the {\tt quietMates} tag.
+9 of the 11 true positive {\tt quietMates} games were classified correctly using only the features that describe the contents of the board.
+On the other hand, 5 true positive {\\tt quietMates} games were correctly classified when using only features about the possible moves in that position.
+Interestingly, disabling the ``checkmate move'' feature did not affect the number of true positives.
 
-We sought to improve upon this baseline with an autoencoder. TALK ABOUT AUTOENCODER RESULTS
+\subsection{Classifying with the auto-encoder}
+\begin{table*}
+\centering
+\begin{tabular}{lcccccccc}
+\hline
+\textbf{Type} & \textbf{Hidden layers} & \textbf{TP} & \textbf{FP} & \textbf{TN} & \textbf{FN} & \textbf{Precision} & \textbf{Recall} & \textbf{F1} \\ \hline
+Unsegmented & 0 & 1.49\% & 3.62\% & 93.26\% & 1.61\% & 29.26\% & 48.17\% & 36.40\% \\
+Unsegmented & 1 & 1.35\% & 3.83\% & 93.06\% & 1.76\% & 26.01\% & 43.29\% & 32.49\% \\
+Unsegmented & 2 & 1.40\% & 4.48\% & 92.41\% & 1.71\% & 23.87\% & 45.12\% & 31.22\% \\
+Unsegmented & 3 & 1.40\% & 4.48\% & 92.41\% & 1.71\% & 23.87\% & 45.12\% & 31.22\% \\
+\hline
+Segmented & 0 & 0.38\% & 1.01\% & 97.35\% & 1.26\% & 27.08\% & 22.94\% & 24.84\% \\
+Segmented & 1 & 0.49\% & 2.90\% & 95.46\% & 1.15\% & 14.49\% & 30.00\% & 19.54\% \\
+Segmented & 2 & 0.22\% & 1.27\% & 97.09\% & 1.42\% & 14.84\% & 13.53\% & 14.15\% \\
+Segmented & 3 & 0.59\% & 3.89\% & 94.47\% & 1.05\% & 13.12\% & 35.88\% & 19.21\% \\
+\end{tabular} 
+\caption{Auto-encoder classification results for the games that have not been segmented by the color of the active player, and those that have been, as the number of hidden layers varies.}
+\label{table:auto-encoder-results}
+\end{table*}
 
-\section{Implementation} % Parallelization and Scaling
-Running serially, the autoencoder takes a long time to complete, even on simple inputs.
-Therefore, we examined ways to parallelize the task in order to fully leverage the capacity of our 8-core server with hyper-threading (16 virtual cores).
-We implemented a local version of MapReduce that creates 16 execution threads.
+We first tried to use the unsupervised fine-tuned stacked auto-encoder's hidden outputs as the input to the SVM classifiers, as with the manually-generated feature vectors.
+However, we found that the resulting feature vectors were not very useful for SVM, as the classifiers did not predict any labels as true for all training or test input games.
+
+Instead, we used the semi-supervised classification method as described in Section~\ref{sec:classifying-auto-encoder}.
+First, we greedily train each hidden layer as an auto-encoder on the training, test, and validation sets.
+We trained the combined neural network composed of greedily-trained auto-encoders to, for each input, predict an output vector corresponding to the labels that apply to the input---0 if a label did not apply, and 1 if it did.
+From the output vectors computed with the elements in the validation set, we tuned the threshold we use in considering whether a label is present or not present to maximize the F1 score.
+In Table~\ref{table:auto-encoder-results}, we report the performance achieved on the test set when using the threshold tuned to maximize the F1 score on the validation set.
+
+These results show higher F1 score than the SVM classifier, with significantly lower precision but also much higher recall.
+Unfortunately, the results indicate that the representation of the board learned by the auto-encoder does not correlates very well with the concepts we wanted to identify, given that having no hidden layers resulted in higher scores.
+Adding extra hidden layers seems to not have helped improve the classification results, except in the last instance, where an extra layer improved the F1 score to original levels.
+It is hard to tell whether some slight change in the auto-encoder training procedure will be able to drastically improve the results, or whether significant changes in the architecture will be required.
+We also were not able to use the unlabeled chess games we collected with the auto-encoder, due to the long training time required and our relatively slow implementation of its training.
 
 \subsection{Discussion}
-We discovered that there is an inherent trade off between how general the approach is and the quality of the outcomes.
-In pursuing this project, we attempted to maintain an approach that is as general as possible so that it can be applied to other contexts.
-However, in attempting to maintain generality, we sacrificed access to powerful but chess-specific tools like game engines, resulting in a classifier that is more surface-level (yet slower) than existing chess analysis tools that leverage the game engine.
+In pursuing this project, we attempted to maintain an approach that is as general as possible so that it can be applied to other contexts, and to minimize the amount of engineering required very specific to chess.
+However, in attempting to maintain generality, we sacrificed access to powerful but chess-specific tools like chess engines.%, resulting in a classifier that is more surface-level (yet slower) than existing chess analysis tools that leverage the game engine.
+
+A chess-specific approach would be able to perfectly identify the more precisely-described tags like {\tt doubleAttack} and {\tt queenCapture} by analyzing the future branches and looking for specific outcomes.
+In addition, it would be able to identify less concrete tags like {\tt clearance} and {\tt kingToBadSquares}.
+However, one advantage of the statistical approach is that feature functions can be written once and reused for all inputs.
+In contrast, in order to effectively identify tags using a chess engine, each tag must have a specific heuristic function written for that express purpose to identify certain properties in future states.
+A better automated sportscasting for chess will leverage the strengths of each approach by using the chess-specific features when they are available, and combining that with results from statistical learning.
+For instance, while the chess engine can easily identify imminent mate combinations, it would require a coordination of both the engine and the statistical learning tool to identify threshold functions like {\tt kingToBadSquares}.
+Machine learning would be used to optimize weights and thresholds (e.g.
+for whether a king’s position is bad) while the engine would be used to explore future paths.
+
+Based on the performance of our general approach, we believe that there is an inherent trade off between how general the approach is and the quality of the outcomes.
 
@@
 There are three areas in which future work can make progress on automated sportscasting.
-It could improve the performance of the autoencoder so that it can be trained on a larger data set.
+It could improve the performance of the auto-encoder so that it can be trained on a larger data set.
 Furthermore, it could focus on trying different methods to improve the classification quality by leveraging chess-specific tools and features like chess engines and evaluation tools.
@@
 One of the main limitations of our work is performance.
-Due to how slowly the autoencoder performs, we were not able to fully exploit the millions of games that we have available to generate features for our unsupervised system.
-Therefore, extensions to our work should investigate ways to improve the performance so that the autoencoder can process more inputs and thus generate more relevant feature functions.
+Due to how slowly the auto-encoder performs, we were not able to fully exploit the millions of games that we have available to generate features for our unsupervised system.
+Therefore, extensions to our work should investigate ways to improve the performance so that the auto-encoder can process more inputs and thus generate more relevant feature functions.
 
 \paragraph{Distributed MapReduce.}
-Our autoencoder takes advantage of some of the parallel processing capacity of the host system through a local MapReduce scheme.
+Our auto-encoder takes advantage of some of the parallel processing capacity of the host system through a local MapReduce scheme.
 Extending that with a distributed MapReduce framework like Hadoop may improve performance if the additional cost of network I/O are outweighed by the benefits of parallelization.
@@
 Optimization is a computation-heavy task, and the GPU is specifically designed for massively-parallel floating point computations.
-We investigated using JavaCL to parallelize the autoencoder, and demonstrated that it is definitely superior at parallelizable and computation-intensive tasks like matrix multiplication, because the benefits of greater throughput through greater parallelization outweigh the costs of memory transfer.
+We investigated using JavaCL to parallelize the auto-encoder, and demonstrated that it is definitely superior at parallelizable and computation-intensive tasks like matrix multiplication, because the benefits of greater throughput through greater parallelization outweigh the costs of memory transfer.
 
@@
 If the classifier incorporated chess-specific features like those generated by chess engines, it could be much better at identifying the types of strategies that are available to the player.
-For example, a deep traversal of many future paths by a chess engine may reveal that an attack is possible (attackPossible) or that the player can force a pawn promotion (pawnPromotion).
-While a purely computational approach to identifying opportunities and strategies may provide higher-quality or more reliable assessments, a statistical approach may be more robust.
-%talk about what i mean by robust
+For example, a deep traversal of many future paths by a chess engine may reveal that an attack is possible ({\tt attackPossible}) or that the player can force a pawn promotion ({\tt pawnPromotion}).
+While a purely computational approach to identifying opportunities and strategies may provide higher-quality or more reliable assessments, a statistical approach may be more robust to data sparsity.
 
-One potential improvement that does not involve increased reliance on chess-specific technologies is to implement hierarchical tags so that matingAttack, doubleAttack, and similar tags imply a more general tag like attackPossible.
+One potential improvement that does not involve increased reliance on chess-specific technologies is to implement hierarchical tags so that {\tt matingAttack}, {\tt doubleAttack}, and similar tags imply a more general tag like {\tt attackPossible}.
 By doing so, there will be a greater balance of positives and negatives, and labels that applied to games but were previously considered wrong are now considered correct.
@@
 
-Another possible improvement is to use a different classifier than SVM, such as the perceptron or a maximum entropy classifier.
+Another possible improvement is to use a different classifier than an SVM, such as the perceptron or a maximum entropy classifier.
 One issue with the SVM is that often the tags are not so clear-cut: what does it mean that an endgame has “good chances”?
@@
 \section{Conclusion}
+We built two statistical learning systems to classify high-level positional concepts for chess, and described various issues involved in implementing these systems.
+Both classifiers are feature driven and classified by a support vector machine.
+The first system uses manually generated features for a chessboard configuration and the move that led to it, while the second system uses a sparse auto-encoder to automatically discover high-level structures (and hence features) for a given chessboard configuration in an unsupervised fashion.
+We found that the auto-encoder-based classifier outperformed the manual feature vector-based classifier, which outperformed the random baseline classifier: on the same segmented data set, the random baseline classifier had an F1 score of $\mathbf{7.60}$, the best manual feature vector-based classifier had an F1 score of $\mathbf{14.05}$, and the best auto-encoder-based classifier had an F1 score of $\mathbf{24.84}$.
 
 \section{Acknowledgements}
-We would like to thank Dan Klein for teaching us statistical machine learning techniques.
-\section*{Appendix}
+We would like to thank Dan Klein for teaching us statistical machine learning techniques, and Soda Hall for accommodating us for many nights.
 

### 2011-05-18 src/edu/berkeley/nlp/classification/ClassifyLabeledData.java

@@ -8,11 +8,13 @@ import java.io.ObjectInputStream;
 import java.net.URISyntaxException;
 import java.text.ParseException;
 import java.util.Collection;
+import java.util.HashMap;
 import java.util.List;
 import java.util.Map;
 import java.util.Random;
 import java.util.Scanner;
 import java.util.Set;
+import java.util.Map.Entry;
 
 import chesspresso.position.Position;
 
@@ -188,6 +190,9 @@ public class ClassifyLabeledData {
 		int[] errors = new int[4];
 
 		File predictions = File.createTempFile("predictions", "");
+		
+		Map<String, Integer> errorCounts = Maps.newHashMap();
+		Map<String, Integer> truePositives = Maps.newHashMap();
 
 		for (String tag : models.keySet()) {
 			File model = models.get(tag);
@@ -211,11 +216,31 @@ public class ClassifyLabeledData {
 			for (List<PositionWithMoves> sample : testSet.keySet()) {
 				double prediction = predictionScanner.nextDouble();				
 				boolean trueValue = testSet.containsEntry(sample, tag);
-
-				++errors[errorType(trueValue, prediction > 0)];
+				
+				int errorType = errorType(trueValue, prediction > 0);
+				
+				if (errorType % 2 == 1) { // error of some kind
+					int errorCount = errorCounts.containsKey(tag) ? (errorCounts.get(tag) + 1) : 1;
+					errorCounts.put(tag, errorCount);
+				} else if (errorType == 0) {
+					truePositives.put(tag, truePositives.containsKey(tag) ? (truePositives.get(tag) + 1) : 1);
+				}
+				
+				++errors[errorType];
 			}
 		}
 		
+		
+		
+		System.out.println("Sources of negatives");
+		for (Entry<String, Integer> entry : errorCounts.entrySet()) {
+			System.out.printf("\t%s: %d/%d error (%f)\n", entry.getKey(), entry.getValue(), testSet.keySet().size(), (double)entry.getValue()/testSet.keySet().size());
+		}
+		
+		System.out.println("Sources of positives");
+		for (Entry<String, Integer> entry : truePositives.entrySet()) {
+			System.out.printf("\t%s: %d/%d true positive (%f)\n", entry.getKey(), entry.getValue(), testSet.keySet().size(), (double)entry.getValue()/testSet.keySet().size());
+		}
 		return errors;
 	}
 	

### 2011-05-18 paper/paper.tex

@@
 Position & Piece location, potential moves per piece, material count, square control. \\
-Previous move & Capturing move, recapturing move, en passant move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move. \\
+Previous move & Capturing move, recapturing move, \emph{en passant} move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move. \\
 \hline
@@
 
-\subsection{Discussion}
-In pursuing this project, we attempted to maintain an approach that is as general as possible so that it can be applied to other contexts, and to minimize the amount of engineering required very specific to chess.
-However, in attempting to maintain generality, we sacrificed access to powerful but chess-specific tools like chess engines.%, resulting in a classifier that is more surface-level (yet slower) than existing chess analysis tools that leverage the game engine.
-
-A chess-specific approach would be able to perfectly identify the more precisely-described tags like {\tt doubleAttack} and {\tt queenCapture} by analyzing the future branches and looking for specific outcomes.
-In addition, it would be able to identify less concrete tags like {\tt clearance} and {\tt kingToBadSquares}.
-However, one advantage of the statistical approach is that feature functions can be written once and reused for all inputs.
-In contrast, in order to effectively identify tags using a chess engine, each tag must have a specific heuristic function written for that express purpose to identify certain properties in future states.
-A better automated sportscasting for chess will leverage the strengths of each approach by using the chess-specific features when they are available, and combining that with results from statistical learning.
-For instance, while the chess engine can easily identify imminent mate combinations, it would require a coordination of both the engine and the statistical learning tool to identify threshold functions like {\tt kingToBadSquares}.
-Machine learning would be used to optimize weights and thresholds (e.g.
-for whether a king’s position is bad) while the engine would be used to explore future paths.
-
-Based on the performance of our general approach, we believe that there is an inherent trade off between how general the approach is and the quality of the outcomes.
-
 \section{Future Work}
-There are three areas in which future work can make progress on automated sportscasting.
+We identify three areas in which future work can make progress on automated sportscasting for chess.
 It could improve the performance of the auto-encoder so that it can be trained on a larger data set.
@@
 
+\subsection{Discussion}
+In pursuing this project, we attempted to maintain an approach that is as general as possible so that it can be applied to other contexts, and to minimize the amount of engineering required very specific to chess.
+However, in attempting to maintain generality, we sacrificed access to powerful but chess-specific tools like chess engines.%, resulting in a classifier that is more surface-level (yet slower) than existing chess analysis tools that leverage the game engine.
+
+A chess-specific rule-based approach should be able to easily identify the more precisely-described themes like {\tt doubleAttack} and {\tt queenCapture} by analyzing the future branches of the game and looking for specific outcomes.
+However, one advantage of the statistical approach is that feature functions can be written once and reused for all possible themes.
+In addition, it would help in identifying more subjective themes like {\tt clearance} and {\tt kingToBadSquares} based more on human judgement.
+In contrast, in order to effectively identify tags using a chess engine, each theme must have a specific heuristic function written for that express purpose to identify certain properties in future states.
+A better automated sportscasting for chess will leverage the strengths of each approach by using the chess-specific features when they are available, and combining that with results from statistical learning.
+For instance, while the chess engine can easily identify imminent mate combinations, it would require a coordination of both the engine and the statistical learning tool to identify threshold functions like {\tt kingToBadSquares}.
+Machine learning would be used to optimize weights and thresholds (e.g.
+for whether a king’s position is bad) while the engine would be used to explore future paths.
+
+Based on the performance of our general approach, we believe that there is an inherent trade off between how general the approach is and the quality of the outcomes.
+
 \subsection{Generalization to Other Games}

### 2011-05-18 paper/paper.tex
Final paper

@@
 \setlength\titlebox{6.5cm}    % Expanding the titlebox
+\frenchspacing
 
@@
 \begin{abstract}
-We investigate applications of statistical learning to automated sportscasting for chess.
-In particular, this project studies methods to automatically identify high level structures in chess games using support vector machines and auto-encoders.
-We then compare the results of the two different approaches and showed improvements over the SVM model using a stacked auto-encoder.
-Finally, we conclude with results, analysis of errors, and ideas for future work.
+We investigate applications of statistical learning to automated 
+sportscasting for chess. In particular, this project studies two methods 
+to automatically identify high level structures in chess games using 
+support vector machines. Our first classifier is based on manual feature 
+selection, while our second classifier employs sparse auto-encoding to 
+automatically generate a set of feature vectors for a chess game. We 
+then compare the results of the two different approaches and showed 
+improvements using auto-encoding techniques. Finally, we conclude with 
+results, analysis of errors, and ideas for potential future work.
 \end{abstract}
@@
 \section{Introduction}
-Chess has a long history of intriguing enthusiasts due to its complexity.
-Even after years of play, each game is new challenge, a new problem to solve.
-Historically, chess has been prominent as a sport where the battles of grandmasters engulf thousands; given the highly intricate nature of chess (and similar abstract strategy games), experts have scrutinized these games to devise standard high-level structures like openings, strategies, tactics, and time advantage, which they use to explain maneuvers and blunders that occur in the games.
-Devotees of chess also attempt to improve their own performance in terms of these high-level structures, reading chess theorists' comments about games by high-ranking players.
+Chess has a long history of intriguing enthusiasts due to its 
+complexity. Even after years of play, each game is new challenge, a new 
+problem to solve. Historically, chess has been prominent as a sport 
+where the battles of grandmasters engulf thousands; given the highly 
+intricate nature of chess (and similar abstract strategy games), experts 
+have scrutinized these games to devise standard high-level structures 
+like openings, strategies, tactics, and time advantage, which they use 
+to explain maneuvers and blunders that occur in the games. Devotees of 
+chess also attempt to improve their own performance in terms of these 
+high-level structures, reading chess theorists' comments about games by 
+high-ranking players.
 
-In this paper, with the end goal of automatically producing textual commentary similar to those produced by experts for grandmaster games, we investigate whether we might automatically identify such high-level patterns in real games.
-Specifically, we want to see how statistical machine learning techniques can help us automatically identify such structures using a large number of real chess games as training data, where some of the games may be already annotated with these high-level structures, while others may not be.
+In this paper, with the end goal of automatically producing textual 
+commentary similar to those produced by experts for grandmaster games, 
+we investigate whether we might automatically identify such high-level 
+patterns in real games. Specifically, we want to see how statistical 
+machine learning techniques can help us automatically identify such 
+structures using a large number of real chess games as training data, 
+where some of the games may be already annotated with these high-level 
+structures, while others may not be.
 
-While others have created highly rigid and rule-based approaches to annotating chess games~\cite{cambridge-chess-annotation}, we specifically sought to avoid such approaches which require a lot of manual engineering and chess-specific expertise in order to obtain good results.
-In particular, for chess, there exists a large number of computer chess tools like chess engines, which search the game tree to examine the breadth of options available and identify the outcomes of each.
-We would like our approaches to have general applicability to other domains, so we chose to specifically avoid using such tools as much as possible.
+While others have created highly rigid and rule-based approaches to 
+annotating chess games~\cite{cambridge-chess-annotation}, we 
+specifically sought to avoid such approaches which require a lot of 
+manual engineering and chess-specific expertise in order to obtain good 
+results. In particular, for chess, there exists a large number of 
+computer chess tools like chess engines, which search the game tree to 
+examine the breadth of options available and identify the outcomes of 
+each. We would like our approaches to have general applicability to 
+other domains, so we chose to specifically avoid using such tools as 
+much as possible.
 
-To that end, we investigate the use of unsupervised feature learning and deep learning techniques on transcripts of chess games to see if they could automatically identify the high-level structures that make up a chess game.
-Deep learning techniques, like the sparse stacked auto-encoder, learns a more efficient way to represent some input given a large set of examples.
-In other fields such as computer vision, researchers have found that these techniques can give state-of-the-art results rivaling or exceeding features hand-engineered by experts.
-While learning a more compact way to represent parts of chess games, auto-encoders may be able to discover high-level structures similar to those used by human experts.
+To that end, we investigate the use of unsupervised feature learning and 
+deep learning techniques on transcripts of chess games to see if they 
+could automatically identify the high-level structures that make up a 
+chess game. Deep learning techniques, like the sparse stacked 
+auto-encoder, learns a more efficient way to represent some input given 
+a large set of examples. In other fields such as computer vision, 
+researchers have found that these techniques can give state-of-the-art 
+results rivaling or exceeding features hand-engineered by experts. While 
+learning a more compact way to represent parts of chess games, 
+auto-encoders may be able to discover high-level structures similar to 
+those used by human experts.
 
-As a comparison, we also implement a small number of chess-specific, hand-engineered features based on information from a chess engine, such as the number of moves available from a specific board position.
-Given feature vectors made from hand-labeled games, we train SVM classifiers to predict the label of an unlabeled feature vector.
-We compare its performance to automatically learned features to see which one can more successfully predict high-level structures.
+As a comparison, we also implement a small number of chess-specific, 
+hand-engineered features based on information from a chess engine, such 
+as the number of moves available from a specific board position. Given 
+feature vectors made from hand-labeled games, we train SVM classifiers 
+to predict the label of an unlabeled feature vector. We compare its 
+performance to automatically learned features to see which one can more 
+successfully predict high-level structures.
 
-To conduct our experiments, we collect several thousand ``tactical training positions'' organized by theme, intended for use by human chess players to learn how to recognize such situations and improve their playing skills.
-We also collect a large number of unannotated chess games (about 2.5 million) from various sources on the web for use with semi-supervised and unsupervised learning techniques.
-After training our models, for each of the positions in the training data labeled with a theme (or high-level structure), we attempt to predict the correct theme(s) for a position using the models.
+To conduct our experiments, we collect several thousand ``tactical 
+training positions'' organized by theme, intended for use by human chess 
+players to learn how to recognize such situations and improve their 
+playing skills. We also collect a large number of unannotated chess 
+games (about 2.5 million) from various sources on the web for use with 
+semi-supervised and unsupervised learning techniques. After training our 
+models, for each of the positions in the training data labeled with a 
+theme (or high-level structure), we attempt to predict the correct 
+theme(s) for a position using the models.
 
-Unfortunately, we found that our approaches did not deliver accurate results in our experiments.
-In particular, the auto-encoder's output failed to serve as a useful criterion for distinguishing between different types of themes.
-We examine the types of correct predictions made by our classifiers and discuss how we may improve our results in the future.
+Unfortunately, we found that our approaches did not deliver accurate 
+results in our experiments. In particular, the auto-encoder's output did 
+not help significantly as a useful criterion for distinguishing between 
+different types of themes. We examine the types of correct predictions 
+made by our classifiers and discuss how we may improve our results in 
+the future.
 
@@
 
-\paragraph{Deep learning.} These techniques attempt to capture complex relationships between input variables in order to transform the input into a different representation which better expresses the relevant structures that make up the input.
-Specifically, they compose many layers of non-linearities in order to compactly represent complex functions which can model these relationships.
-They have been used successfully for many different types of tasks \cite{Vincent,Lee,Socher}.
+\paragraph{Deep learning.} These techniques attempt to capture complex 
+relationships between input variables in order to transform the input 
+into a different representation which better expresses the relevant 
+structures that make up the input. Specifically, they compose many 
+layers of non-linearities in order to compactly represent complex 
+functions which can model these relationships. They have been used 
+successfully for many different types of tasks, especially in computer 
+vision~\cite{Vincent,Lee,Socher}.
 
-\paragraph{Automated sportscasting.} Chen and Mooney~\shortcite{chen} present a commentator system that learns language from sportscasts of simulated RoboCup games.
-Their system, which is trained on ambiguous supervision in the form of textual human commentaries and simulation states of the soccer games, simultaneously tries to establish correspondences between the commentaries and the simulation states.
-Their work focuses on generating meaningful English commentary from relatively simple topics such as ``Player 1 kicked the ball to Player 2'', while our work focuses less on text generation and more on identifying complex topics.
+\paragraph{Automated sportscasting.} Chen and Mooney~\shortcite{chen} 
+present a commentator system that learns language from sportscasts of 
+simulated RoboCup games. Their system, which is trained on ambiguous 
+supervision in the form of textual human commentaries and simulation 
+states of the soccer games, simultaneously tries to establish 
+correspondences between the commentaries and the simulation states. 
+Their work focuses on generating meaningful English commentary from 
+relatively simple topics such as ``Player 1 kicked the ball to Player 
+2'', while our work focuses less on text generation and more on 
+identifying complex topics.
 
-\paragraph{Automatic chess commentary.} There exist techniques for automatic annotation of chess games which are heavily dependent on rule-based logic and use of a chess-engine to determine which parts of a specific chess game are salient~\cite{cambridge-chess-annotation}.
-As a departure from the rule-based approach to automatic chess commentary, Guid et al.~\shortcite{Guid} discuss the use of argument based machine learning, which analyzes reasoning where arguments for and against a certain claim are produced and evaluated (in the context of chess, why a certain move was played over others.) Their system, much like the one presented in this paper, identifies complex positional features of a given chess game.
+\paragraph{Automatic chess commentary.} There exist techniques for 
+automatic annotation of chess games which are heavily dependent on 
+rule-based logic and use of a chess-engine to determine which parts of a 
+specific chess game are salient~\cite{cambridge-chess-annotation}. As a 
+departure from the rule-based approach to automatic chess commentary, 
+Guid et al.~\shortcite{Guid} discuss the use of argument based machine 
+learning, which analyzes reasoning where arguments for and against a 
+certain claim are produced and evaluated (in the context of chess, why a 
+certain move was played over others.) Their system, much like the one 
+presented in this paper, identifies complex positional features of a 
+given chess game.
 
-However, we identified some shortcomings of the AMBL system.
-First, it relies on experts to provide arguments as to why a certain position exemplifies (or doesn’t) a particular positional feature.
-While this is an improvement from defining a classification rule manually, it is infeasible to repeat this procedure multiple times.
-Second, the arguments that the experts are allowed to provide rely on a preset list of features.
+However, we identified some shortcomings of the AMBL system. First, it 
+relies on experts to provide arguments as to why a certain position 
+exemplifies (or doesn.t) a particular positional feature. While this is 
+an improvement from defining a classification rule manually, it is 
+infeasible to repeat this procedure multiple times. Second, the 
+arguments that the experts are allowed to provide rely on a preset list 
+of features.
 
 \section{Data Sources and Processing}
-We obtained both annotated and unannotated records of chess games from various web sites in the PGN format \cite{pgn}, a widely-used format for compactly representing chess games.
-A large number of games as well as software and tools that use the PGN format are freely available.
+We obtained both annotated and unannotated records of chess games from 
+various web sites in the PGN format \cite{pgn}, a widely-used format for 
+compactly representing chess games. A large number of games as well as 
+software and tools that use the PGN format are freely available.
 
-\begin{table*}
+\begin{table*}[p]
 \centering
+\vspace{-0.7cm}
 \begin{tabular}{lp{0.6\textwidth}}
@@
 \tt{attackPossible} & The current player has opportunity to attack. \\
-\tt{backrankWeakness} & There exists a future checkmate by a rook or queen on the edges of the board. \\
-\tt{bishopPair} & There exists a decisive strategic advantage for the player with two bishops. \\
+\tt{backrankWeakness} & There exists a future checkmate by a rook or 
+queen on the edges of the board. \\
+\tt{bishopPair} & There exists a decisive strategic advantage for the 
+player with two bishops. \\
 \tt{clearance} & There exists a maneuver to break the pawn structure. \\ 
@@
 \tt{decoy} & There exists a maneuver to distract a crucial piece. \\
-\tt{defensePressure} & There exists a maneuver to break the opposing King’s defense. \\
+\tt{defensePressure} & There exists a maneuver to break the opposing 
+King.s defense. \\
 \tt{deflection} & There exists a maneuver to deflect an attack.  \\ 
-\tt{discoveredAttack} & There exists a maneuver to allow for a discovered attack (an attack that was blocked by a friendly piece.) \\
-\tt{doubleAttack} & There exists a move that simultaneously attacks at least two crucial squares. \\
-\tt{drawCombinationsStalemate} & There exists a combination that allows for decisive material advantage or immediate checkmate. \\
-\tt{enclosedKings} & There exists a maneuver to trap the opposing King. \\
-\tt{freeingBlockingSquares} & There exists a move to free a crucial square. \\
-\tt{goodEndgameChances} & There exists a move to allow for a favorable endgame. \\
-\tt{inBetweenMove} & There exists an interrupting move that allows for decisive material advantage of immediate checkmate. \\ 
+\tt{discoveredAttack} & There exists a maneuver to allow for a 
+discovered attack (an attack that was blocked by a friendly piece.) \\
+\tt{doubleAttack} & There exists a move that simultaneously attacks at 
+least two crucial squares. \\
+\tt{drawCombinationsStalemate} & There exists a combination that allows 
+for decisive material advantage or immediate checkmate. \\
+\tt{enclosedKings} & There exists a maneuver to trap the opposing King. 
+\\
+\tt{freeingBlockingSquares} & There exists a move to free a crucial 
+square. \\
+\tt{goodEndgameChances} & There exists a move to allow for a favorable 
+endgame. \\
+\tt{inBetweenMove} & There exists an interrupting move that allows for 
+decisive material advantage of immediate checkmate. \\ 
 \tt{interference} & There exists a maneuver to block a crucial piece. \\ 
-\tt{interruption} & There exists a maneuver that interrupts the opponent’s tactics. \\
-\tt{kingInCenter} & There exists a decisive strategic advantage for the player with the King in the center of the board. \\
-\tt{kingToBadSquares} & There exists a combination that forces the opposing King to move to a bad square. \\
-\tt{longDiagonals} & There exists a strategic advantage for the player that controls squares on the long diagonal. \\
+\tt{interruption} & There exists a maneuver that interrupts the 
+opponent.s tactics. \\
+\tt{kingInCenter} & There exists a decisive strategic advantage for the 
+player with the King in the center of the board. \\
+\tt{kingToBadSquares} & There exists a combination that forces the 
+opposing King to move to a bad square. \\
+\tt{longDiagonals} & There exists a strategic advantage for the player 
+that controls squares on the long diagonal. \\
 \tt{matingAttack} & There exists a decisive mating attack. \\
-\tt{openingClosingDiagonals} & There exists a maneuver to clear a diagonal. \\
+\tt{openingClosingDiagonals} & There exists a maneuver to clear a 
+diagonal. \\
 \tt{openingClosingFiles} & There exists a maneuver to clear a file. \\
-\tt{overload} & There exists a combination that makes a defending piece defend too many squares at once. \\
-\tt{passedPawn} & There exists a combination that allows for a passed pawn (a pawn with no opposing pawns in front of it.) \\
-\tt{pawnPromotion} & There exists a combination that allows for a pawn promotion. \\
-\tt{perpetualCheck} & There exists a combination that forces a perpetual check/draw. \\
+\tt{overload} & There exists a combination that makes a defending piece 
+defend too many squares at once. \\
+\tt{passedPawn} & There exists a combination that allows for a passed 
+pawn (a pawn with no opposing pawns in front of it.) \\
+\tt{pawnPromotion} & There exists a combination that allows for a pawn 
+promotion. \\
+\tt{perpetualCheck} & There exists a combination that forces a perpetual 
+check/draw. \\
 \tt{pin} & There exists a combination that allows for a pin. \\
-\tt{pullingPiecesToBadSquares} & There exists a combination that forces an opposing piece to move to a bad square. \\
-\tt{queenCapture} & There exists a combination to capture the opposing Queen. \\
-\tt{quietMates} & There exists a quiet move (one that looks inconsequential on the surface) leading to mate. \\
-\tt{removalOfDefence} & There exists a maneuver to remove an opponent’s defender. \\
+\tt{pullingPiecesToBadSquares} & There exists a combination that forces 
+an opposing piece to move to a bad square. \\
+\tt{queenCapture} & There exists a combination to capture the opposing 
+Queen. \\
+\tt{quietMates} & There exists a quiet move (one that looks 
+inconsequential on the surface) leading to mate. \\
+\tt{removalOfDefence} & There exists a maneuver to remove an opponent.s 
+defender. \\
 \tt{unpinning} & There exists a maneuver to unpin a piece. \\
-\tt{wrongBishop} & The opponent’s only bishop is limited in use because of the pawn structure of the game. \\ \hline
+\tt{wrongBishop} & The opponent.s only bishop is limited in use because 
+of the pawn structure of the game. \\ \hline
 \end{tabular}
-\caption{List of high-level themes we attempt to automatically identify, along with explanations}
+\caption{List of high-level themes we attempt to automatically identify, 
+along with explanations}
 \label{table:themes}
@@
 
-The already-annotated games come from databases of short snippets of chess games,\footnote{\url{http://wwwu.uni-klu.ac.at/gossimit/c/tactic.htm}} \footnote{\url{http://webplaza.pt.lu/public/ckaber/Chess.htm#Chess training}} intended for chess players to learn from by example, which are organized by particular themes that the students are to identify.
-These provided a useful repository of both interesting high-level concepts we could identify (we provide a listing in Table~\ref{table:themes}), as well as boards and moves corresponding to each one.
+The already-annotated games come from databases of short snippets of 
+chess 
+games,\footnote{\url{http://wwwu.uni-klu.ac.at/gossimit/c/tactic.htm}} 
+\footnote{\url{http://webplaza.pt.lu/public/ckaber/Chess.htm#Chess 
+training}} intended for chess players to learn from by example, which 
+are organized by particular themes that the students are to identify. 
+These provided a useful repository of both interesting high-level 
+concepts we could identify (we provide a listing in 
+Table~\ref{table:themes}), as well as boards and moves corresponding to 
+each one.
 
-We also looked for as many chess games as possible from a variety of web sites, such as \url{chessgames.com}.
-We easily obtained over five million untagged games spread over 600,000 files and 5 GB of data.
-The number of annotated game snippets we obtained was slightly over 5,000.
+We also looked for as many chess games as possible from a variety of web 
+sites, such as \url{chessgames.com}. We easily obtained over five 
+million untagged games spread over 600,000 files and 5 GB of data. The 
+number of annotated game snippets we obtained was slightly over 5,000.
 
 \subsection{Data Pre-Processing}	
-In order to ensure consistency and integrity over the games obtained from disparate sources, we sanitized them in multiple ways.
-We parsed each game in the data set and removed information that would confuse the parser.
-For example, some games had English or German comments by humans which we did not plan to use, and sometimes resulted in parsing errors due to specific characters present in the comments.
-Others had alternate branches, sometimes up to multiple levels (alternate branches within alternate branches), which the parser could not handle properly.
-Still others represented games in a non-standard way (for example, pawn promotions were represented without the ``=''), which we had to fix.
+In order to ensure consistency and integrity over the games obtained 
+from disparate sources, we sanitized them in multiple ways. We parsed 
+each game in the data set and removed information that would confuse the 
+parser. For example, some games had English or German comments by humans 
+which we did not plan to use, and sometimes resulted in parsing errors 
+due to specific characters present in the comments. Others had alternate 
+branches, sometimes up to multiple levels (alternate branches within 
+alternate branches), which the parser could not handle properly. Still 
+others represented games in a non-standard way (for example, pawn 
+promotions were represented without the ``=''), which we had to fix.
 
-Since we obtained the games from many sources, there was significant overlap in games (approximately 50\% of the unlabeled games collected were redundant).
-Therefore, we also went through all of the downloaded games and eliminated duplicate games: multiple games that had the same initial position and move sequence were considered redundant and only one of each were kept.
+Since we obtained the games from many sources, there was significant 
+overlap in games (approximately 50\% of the unlabeled games collected 
+were redundant). Therefore, we also went through all of the downloaded 
+games and eliminated duplicate games: multiple games that had the same 
+initial position and move sequence were considered redundant and only 
+one of each were kept.
 
 \subsection{Parsing}
-We tried various open source PGN parsers, each succumbing to flaws like mishandling uncommon cases (such as  promotions and special moves like \emph{en passant}).
-We based our system on Chesspresso,\footnote{\url{http://www.chesspresso.org/}} a Java library for writing chess programs, which we fixed and modified for our needs.
-Chesspresso takes a PGN file as input and returns a data structure that represents the initial position of the board and the subsequent moves (as well as all metadata necessary to recreate the PGN file).
-We also extended it to be more robust to variations in input format, in addition to being able to export PGNs in a canonical form, which was crucial for removing duplicates.
+We tried various open source PGN parsers, each succumbing to flaws like 
+mishandling uncommon cases (such as  promotions and special moves like 
+\emph{en passant}). We based our system on 
+Chesspresso,\footnote{\url{http://www.chesspresso.org/}} a Java library 
+for writing chess programs, which we fixed and modified for our needs. 
+Chesspresso takes a PGN file as input and returns a data structure that 
+represents the initial position of the board and the subsequent moves 
+(as well as all metadata necessary to recreate the PGN file). We also 
+extended it to be more robust to variations in input format, in addition 
+to being able to export PGNs in a canonical form, which was crucial for 
+removing duplicates.
 
 \subsection{Preparing the Data}
-After parsing each game, we filtered out the tagged games that had more than 10 plies because it will be very difficult to identify the tags that apply to a complex sequence of moves.
-Among the games that we kept, we also kept only the first few positions and moves (the ``prefix'' of each game, as described earlier), a parameter that can be adjusted; for our experiments, we kept just the initial board specified in each game.
-We then generate labels for each game based on the information from the source websites.
-For example, games in a file that corresponds to a lesson called ``Double Attack'' are labeled {\tt doubleAttack}.
-In addition, some of those games may appear in other lessons, and consequently will have multiple labels.
-We also segment the games further by the color of the player that makes the first move in the game, in addition to the labels of the game, since most of the labels in Table~\ref{table:themes} only refer to a specific player; for instance, some of the games labeled with {\tt backrankWeakness} may have the white player with the weakness, while others show the black player with the weakness.
+After parsing each game, we filtered out the tagged games that had more 
+than 10 plies because it will be very difficult to identify the tags 
+that apply to a complex sequence of moves. Among the games that we kept, 
+we also kept only the first few positions and moves (the ``prefix'' of 
+each game, as described earlier), a parameter that can be adjusted; for 
+our experiments, we kept just the initial board specified in each game. 
+We then generate labels for each game based on the information from the 
+source websites. For example, games in a file that corresponds to a 
+lesson called ``Double Attack'' are labeled {\tt doubleAttack}. In 
+addition, some of those games may appear in other lessons, and 
+consequently will have multiple labels. We also segment the games 
+further by the color of the player that makes the first move in the 
+game, in addition to the labels of the game, since most of the labels in 
+Table~\ref{table:themes} only refer to a specific player; for instance, 
+some of the games labeled with {\tt backrankWeakness} may have the white 
+player with the weakness, while others show the black player with the 
+weakness.
 
@@
 \subsection{Sparse auto-encoder}
-The sparse auto-encoder is one of many methods in use for unsupervised feature learning~\cite{Coates}.
-It takes the form of a feedforward neural network, which consists of an input layer of N nodes, some number of hidden layers, and output layer of M nodes.
-The value of each layer $\vec{l_i}$ is determined with a weight matrix $W_i$ and a bias vector $b_i$: \[L_i = f(W_i \cdot \vec{l_{i-1}} + b_i),\] where $f(\cdot)$ is a non-linearity like the sigmoid $f(x) = \frac{1}{1 + e^{-z}}$ applied component-wise.
+The sparse auto-encoder is one of many methods in use for unsupervised 
+feature learning~\cite{Coates}. It takes the form of a feedforward 
+neural network, which consists of an input layer of N nodes, some number 
+of hidden layers, and output layer of M nodes. The value of each layer 
+$\mathbf{l_i}$ is determined with a weight matrix $W_i$ and a bias 
+vector $b_i$: \[\mathbf{l_i} = f(W_i \cdot \mathbf{l_{i-1}} + b_i),\] 
+where $f(\cdot)$ is a non-linearity like the sigmoid $f(x) = \frac{1}{1 
++ e^{-z}}$ applied component-wise.
 
-An auto-encoder is such a neural network trained so as to minimize the error (specifically, the sum-squared difference) between the output layer and the input layer, over some set of input examples.
-If we have fewer nodes in the hidden layers than in the input or output layer, the neural network is forced to learn a compressed representation of the data (represented by the activations, or outputs, of the smaller number of hidden nodes).
-Instead, we keep a larger number of hidden nodes while imposing a sparsity constraint that the average activation of each one be small (e.g., 0.05).
-Other work has shown that such a constraint better forces the auto-encoder to learn interesting structures from the data~\cite{bengio}.
+An auto-encoder is such a neural network trained so as to minimize the 
+error (specifically, the sum-squared difference) between the output 
+layer and the input layer, over some set of input examples. If we have 
+fewer nodes in the hidden layers than in the input or output layer, the 
+neural network is forced to learn a compressed representation of the 
+data (represented by the activations, or outputs, of the smaller number 
+of hidden nodes). Instead, we keep a larger number of hidden nodes while 
+imposing a sparsity constraint that the average activation of each one 
+be small (e.g., 0.05). Other work has shown that such a constraint 
+better forces the auto-encoder to learn interesting structures from the 
+data~\cite{bengio}.
 
-A normal auto-encoder only has one layer of hidden nodes, but we can also add more hidden layers to form a stacked auto-encoder.
-Per Bengio et al.~\shortcite{Bengio07greedylayer-wise}, we use a greedy strategy to train the extra hidden layers in the auto-encoder.
-We train an auto-encoder with one hidden layer as usual; once this completes, we obtain the output of the hidden layer over all the examples in the input set, and use that as the new set of examples.
-This allows us to learn higher-level representations of the data~\cite{Vincent}.
+A normal auto-encoder only has one layer of hidden nodes, but we can 
+also add more hidden layers to form a stacked auto-encoder. Per Bengio 
+et al.~\shortcite{Bengio07greedylayer-wise}, we use a greedy strategy to 
+train the extra hidden layers in the auto-encoder. We train an 
+auto-encoder with one hidden layer as usual; once this completes, we 
+obtain the output of the hidden layer over all the examples in the input 
+set, and use that as the new set of examples. This allows us to learn 
+higher-level representations of the data~\cite{Vincent}.
 
-Other strategies for unsupervised feature learning include restricted Boltzmann machines (which are undirected), deep belief nets (stacked RBMs, akin to stacked auto-encoders), and denoising auto-encoders (which try to reconstruct some input from a corrupted version of it).
-We only evaluated sparse stacked auto-encoders in this paper.
+Other strategies for unsupervised feature learning include restricted 
+Boltzmann machines (which are undirected), deep belief nets (stacked 
+RBMs, akin to stacked auto-encoders), and denoising auto-encoders (which 
+try to reconstruct some input from a corrupted version of it). We only 
+evaluated sparse stacked auto-encoders in this paper.
 
-The auto-encoder requires training data in the form of vectors where all the elements are between 0 and 1, or the range of the sigmoid function.
-We generated the training vectors by, for some chess board and for every type of piece (knight, rook, etc.), whether a white one (valued 1.0), a black one (valued 0.0), or none at all (valued 0.5) was present in every square of the chessboard.
+The auto-encoder requires training data in the form of vectors where all 
+the elements are between 0 and 1, or the range of the sigmoid function. 
+We generated the training vectors by, for some chess board and for every 
+type of piece (knight, rook, etc.), whether a white one (valued 1.0), a 
+black one (valued 0.0), or none at all (valued 0.5) was present in every 
+square of the chessboard.
 
@@
 \textbf{Category} & \textbf{Description} \\ \hline
-Position & Piece location, potential moves per piece, material count, square control. \\
-Previous move & Capturing move, recapturing move, \emph{en passant} move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move. \\
+Position & Piece location, potential moves per piece, material count, 
+square control. \\
+Previous move & Capturing move, recapturing move, \emph{en passant} 
+move, long castling move, short castling move, check move, stalemate 
+move, checkmate move, promotion move. \\
 \hline
@@
 
-As a point of comparison to auto-encoders, we extracted chess-specific features manually from a given chess position (configuration of pieces on the board) and the move that resulted in that position; Table~\ref{table:features} contains a list.
-The position-based features include information such as where the pieces are on the board and potential moves, while the move-based features involve binary information such as if the move that resulted in this position was a capturing move or resulted in a check.
-The resulting feature vector was then L2-normalized to fit the specifications of the support vector machine.
+As a point of comparison to auto-encoders, we extracted chess-specific 
+features manually from a given chess position (configuration of pieces 
+on the board) and the move that resulted in that position; 
+Table~\ref{table:features} contains a list. The position-based features 
+include information such as where the pieces are on the board and 
+potential moves, while the move-based features involve binary 
+information such as if the move that resulted in this position was a 
+capturing move or resulted in a check.  The resulting feature vector was 
+then L2-normalized to fit the specifications of the support vector 
+machine.
 
 \section{Experimental methodology}
-We processed all of the labeled training data to obtain a mapping from each unique game prefix (for most of the experiments, just the initial specified configuration of the board in the labeled game) to the labels that apply to it; most of the prefixes only had one label, while others had two or three.
-We then segmented the data into three disjoint sets: 90\% for training, 5\% as the validation set (for use with tuning our classifiers), and 5\% as the test set.
-The segmenting was done by random selection from the set of all labeled games.
+We processed all of the labeled training data to obtain a mapping from 
+each unique game prefix (for most of the experiments, just the initial 
+specified configuration of the board in the labeled game) to the labels 
+that apply to it; most of the prefixes only had one label, while others 
+had two or three. We then segmented the data (consisting of about 3000 
+board configurations) into three disjoint sets: 90\% for training, 5\% 
+as the validation set (for use with tuning our classifiers), and 5\% as 
+the test set. The segmenting was done by random selection from the set 
+of all labeled games.
 
-In our experiments, we first use the labeled data in the training set (and possibly also unlabeled data) to train the classifiers.
-Then, for every test input chess game (or a prefix of a game), we use our classifiers to apply some number of labels to it that come from Table~\ref{table:themes}.
-We measure the performance of the predictions made by our classifiers by counting the number of true positives (a predicted label was among the set of gold labels for a test input), true negatives (a label in Table~\ref{table:themes} was neither in the set of predicted labels nor the set of gold labels), false positives (a predicted label was not in the set of gold labels), and false negatives (a gold label was not predicted).
-We compute and report the precision, recall, and F1 score using these measures of accuracy.
+In our experiments, we first use the labeled data in the training set 
+(and possibly also unlabeled data) to train the classifiers. Then, for 
+every test input chess game (or a prefix of a game), we use our 
+classifiers to apply some number of labels to it that come from 
+Table~\ref{table:themes}. We measure the performance of the predictions 
+made by our classifiers by counting the number of true positives (a 
+predicted label was among the set of gold labels for a test input), true 
+negatives (a label in Table~\ref{table:themes} was neither in the set of 
+predicted labels nor the set of gold labels), false positives (a 
+predicted label was not in the set of gold labels), and false negatives 
+(a gold label was not predicted). We compute and report the precision, 
+recall, and F1 score using these measures of accuracy.
 
 \subsection{Support vector machine classifier}
-We note that any number of labels may apply to some test input that we want to classify; a straightforward use of a binary (or even multiclass) classifier, as provided by most implementations of support vector machines, will not suffice as it can only assign one label.
-Therefore, we take the one-versus-all strategy of training many binary SVMs; for each label that we consider, we train a binary classifier with all training games that have the given label as positive examples, and all other games as negative examples.
-We used SVMlight~\cite{svmlight}, a popular third-party implementation of SVMs, for our experiments, using the manually-generated features extracted from each labeled data point.
+We note that any number of labels may apply to some test input that we 
+want to classify; a straightforward use of a binary (or even multiclass) 
+classifier, as provided by most implementations of support vector 
+machines, will not suffice as it can only assign one label. Therefore, 
+we take the one-versus-all strategy of training many binary SVMs; for 
+each label that we consider, we train a binary classifier with all 
+training games that have the given label as positive examples, and all 
+other games as negative examples. We used {\tt 
+SVMlight}~\cite{svmlight}, a popular third-party implementation of SVMs, 
+for our experiments, using the manually-generated features extracted 
+from each labeled data point.
 
@@
 \label{sec:classifying-auto-encoder}
-There exists several methods for using the auto-encoder for the task at hand.
-First, we can consider using the output of the last hidden layer as the feature vector for input into the SVM classifier, just as we did with the manually-generated features.
-To do this, we first perform unsupervised fine-tuning; we combine all the greedily-trained auto-encoders to make a many-layered neural network intended to reconstruct the input layer in its output layer, after which we again minimize the sum-squared error between the input and the output layers.
-We can then use the output of the hidden layer (i.e., not the one that reconstructs the input, but the middle layer which provides the most high-level representation of the input) as the feature vector.
+There exists several methods for using the auto-encoder for the task at 
+hand. First, we can consider using the output of the last hidden layer 
+as the feature vector for input into the SVM classifier, just as we did 
+with the manually-generated features. To do this, we first perform 
+unsupervised fine-tuning; we combine all the greedily-trained 
+auto-encoders to make a many-layered neural network intended to 
+reconstruct the input layer in its output layer, after which we again 
+minimize the sum-squared error between the input and the output layers. 
+We can then use the output of the hidden layer (i.e., not the one that 
+reconstructs the input, but the middle layer which provides the most 
+high-level representation of the input) as the feature vector.
 
-We can also use semi-supervised learning methods, in which after greedily training the auto-encoders, we combine all the hidden layers of the auto-encoders and add an output layer at the end.
-For every labeled input game, we construct input-output vectors such that the output encodes the gold labels of the input.
-We then train the resulting multi-layer neural network to minimize the sum-squared error; this acts similarly to a multiclass logistic regression, except with a different objective at the last layer.
+We can also use semi-supervised learning methods, in which after 
+greedily training the auto-encoders, we combine all the hidden layers of 
+the auto-encoders and add an output layer at the end. For every labeled 
+input game, we construct input-output vectors such that the output 
+encodes the gold labels of the input. We then train the resulting 
+multi-layer neural network to minimize the sum-squared error; this acts 
+similarly to a multiclass logistic regression, except with a different 
+objective at the last layer.
 
-To train these neural networks, we used batch learning over all the examples at once.
-The loss function contained components for the sum-squared error as well as an L2 regularization term and a sparsity penalty based on the Kullback-Leibler divergence between the desired average activation of the hidden nodes and the actual average activation.
-We used backpropagation to compute the gradient of the loss function and L-BFGS to optimize the parameters as to minimize the loss function~\cite{ng-lecturenotes}.
-Running serially, the training takes a long time to complete, even on simple inputs.
-Therefore, we examined ways to parallelize the task in order to fully leverage the capacity of modern many-core machines; we achieved a significant speedup by parallelizing the loss function evaluation to 16 execution threads.
+To train these neural networks, we used batch learning over all the 
+examples at once. The loss function contained components for the 
+sum-squared error as well as an L2 regularization term and a sparsity 
+penalty based on the Kullback-Leibler divergence between the desired 
+average activation of the hidden nodes and the actual average 
+activation. We used backpropagation to compute the gradient of the loss 
+function and L-BFGS to optimize the parameters as to minimize the loss 
+function~\cite{ng-lecturenotes}. Running serially, the training takes a 
+long time to complete, even on simple inputs. Therefore, we examined 
+ways to parallelize the task in order to fully leverage the capacity of 
+modern many-core machines; we achieved a significant speedup by 
+parallelizing the loss function evaluation to 16 execution threads.
 
 \subsection{Random baseline}
-To use as a baseline and ensure that our classifiers are giving sensible results, we randomly made predictions for every test game, where each one assigned each possible label at random, with the possibility of the label being given based on the proportion of the training games which have that label.
-The random predictions were made 1,000 times, with the predictions made independently at random, and the error rates were averaged.
+To use as a baseline and ensure that our classifiers are giving sensible 
+results, we randomly made predictions for every test game, where each 
+one assigned each possible label at random, with the possibility of the 
+label being given based on the prior probability of the label, or the 
+proportion of the training games which have that label. The random 
+predictions were made 1,000 times, with the predictions made 
+independently at random, and the error rates were averaged.
 
@@
 \hline
-\textbf{Type} & \textbf{TP} & \textbf{FP} & \textbf{TN} & \textbf{FN} & \textbf{Precision} & \textbf{Recall} & \textbf{F1} \\ \hline
-Unsegmented & 0.21\% & 0.00\% & 96.89\% & 2.90\% & 100\% & 6.71\% & 12.57\% \\
-Baseline (Random) & 0.42\% & 2.76\% & 94.12\% & 2.69\% & 13.32\% & 13.65\% & 13.48\% \\
+\textbf{Type} & \textbf{TP} & \textbf{FP} & \textbf{TN} & \textbf{FN} & 
+\textbf{Precision} & \textbf{Recall} & \textbf{F1} \\ \hline
+Unsegmented & 0.21\% & 0.00\% & 96.89\% & 2.90\% & 100\% & 6.71\% & 
+12.57\% \\
+Baseline (Random) & 0.42\% & 2.76\% & 94.12\% & 2.69\% & 13.32\% & 
+13.65\% & 13.48\% \\
 \hline
-Segmented & 0.13\% & 0.02\% & 98.34\% & 1.51\% & 86.67\% & 7.65\% & 14.05\% \\
-Baseline (Random) & 0.12\% & 1.49\% & 96.87\% & 1.51\% & 7.65\% & 7.55\% & 7.60\% \\
+Segmented & 0.13\% & 0.02\% & 98.34\% & 1.51\% & 86.67\% & 7.65\% & 
+14.05\% \\
+Baseline (Random) & 0.12\% & 1.49\% & 96.87\% & 1.51\% & 7.65\% & 7.55\% 
+& 7.60\% \\ \hline
 \end{tabular} 
-\caption{Classification results for the games that have not been segmented by the color of the active player, and those that have been.
-In the former, precision clearly outperforms a random baseline, but recall falls short, resulting in a lower F1 score.
-In the latter, precision has fallen relative to the unsegmented version, but recall has risen, resulting in a slightly higher F1 score.
-In contrast, the random model became much worse in both ways.}
+\caption{Classification results for the games that have not been 
+segmented by the color of the active player, and those that have been. 
+In the former, precision clearly outperforms a random baseline, but 
+recall falls short, resulting in a lower F1 score. In the latter, 
+precision has fallen relative to the unsegmented version, but recall has 
+risen, resulting in a slightly higher F1 score. In contrast, the random 
+model became much worse in both ways.}
 \label{table:svmResults}
@@
 
-We first describe the results achieved when we trained using the manually-specified feature extractors.
-Since each game has very few labels, most of the binary classification results are negative.
-Therefore, the model is heavily biased toward negative results, as shown in Table~\ref{table:svmResults}.
+We first describe the results achieved when we trained using the 
+manually-specified feature extractors. Since each game has very few 
+labels, most of the binary classification results are negative. 
+Therefore, the model is heavily biased toward negative results, as shown 
+in Table~\ref{table:svmResults}.
 
-This classification method had relatively high precision compared to the random baseline.
-However, the recall is much lower than that of the random classifier.
-As a result, the F1 score (a combination of the precision and recall scores) of the SVM-based model is no better than that of the random model.
+This classification method had relatively high precision compared to the 
+random baseline. However, the recall is much lower than that of the 
+random classifier. As a result, the F1 score (a combination of the 
+precision and recall scores) of the SVM-based model is no better than 
+that of the random model.
 
-When we extended the labeling to also incorporate the first mover's color information, the SVM model's recall was improved at a minor cost to precision, resulting in a higher F1 score.
-In contrast, the random predictions became much worse: Both the precision and recall scores fell as a result of this change, which reduced the evenness of the label distribution among the training games, and reduced the amount of games with each label (since games are now spread over twice as many tags).
+When we extended the labeling to also incorporate the first mover's 
+color information, the SVM model's recall was improved at a minor cost 
+to precision, resulting in a higher F1 score. In contrast, the random 
+predictions became much worse: Both the precision and recall scores fell 
+as a result of this change, which reduced the evenness of the label 
+distribution among the training games, and reduced the amount of games 
+with each label (since games are now spread over twice as many tags).
 \begin{table}
@@
 \tt{pawnPromotion}  & 6.4516\% \\
-\tt{pullingPiecesToBadSquares}  & 0.6452\%
+\tt{pullingPiecesToBadSquares}  & 0.6452\% \\ \hline
 \end{tabular} 
-\caption{Percent of incorrect classification (false negatives and false positives) for various tags.}
+\caption{Percent of incorrect classification (false negatives and false 
+positives) for various tags.}
 \label{table:negatives}
@@
 
-Given the high precision but low recall, we delved deeper to analyze the sources of errors and true positives.
-As we show in Table~\ref{table:negatives}, we found that {\tt deflection} and {\tt removalOfDefence} are much more likely to cause false negatives than other tags.
-In addition, we found that all of the true positives are for the {\tt quietMates} tag.
-9 of the 11 true positive {\tt quietMates} games were classified correctly using only the features that describe the contents of the board.
-On the other hand, 5 true positive {\\tt quietMates} games were correctly classified when using only features about the possible moves in that position.
-Interestingly, disabling the ``checkmate move'' feature did not affect the number of true positives.
+Given the high precision but low recall, we delved deeper to analyze the 
+sources of errors and true positives. As we show in 
+Table~\ref{table:negatives}, we found that {\tt deflection} and {\tt 
+removalOfDefence} are much more likely to cause false negatives than 
+other tags. In addition, we found that all of the true positives are for 
+the {\tt quietMates} tag. 9 of the 11 true positive {\tt quietMates} 
+games were classified correctly using only the features that describe 
+the contents of the board. On the other hand, 5 true positive {\tt 
+quietMates} games were correctly classified when using only features 
+about the possible moves in that position. Interestingly, disabling the 
+``checkmate move'' feature did not affect the number of true positives.
 
@@
 \centering
-\begin{tabular}{lcccccccc}
+\begin{tabular}{ccccccccc}
 \hline
-\textbf{Type} & \textbf{Hidden layers} & \textbf{TP} & \textbf{FP} & \textbf{TN} & \textbf{FN} & \textbf{Precision} & \textbf{Recall} & \textbf{F1} \\ \hline
-Unsegmented & 0 & 1.49\% & 3.62\% & 93.26\% & 1.61\% & 29.26\% & 48.17\% & 36.40\% \\
-Unsegmented & 1 & 1.35\% & 3.83\% & 93.06\% & 1.76\% & 26.01\% & 43.29\% & 32.49\% \\
-Unsegmented & 2 & 1.40\% & 4.48\% & 92.41\% & 1.71\% & 23.87\% & 45.12\% & 31.22\% \\
-Unsegmented & 3 & 1.40\% & 4.48\% & 92.41\% & 1.71\% & 23.87\% & 45.12\% & 31.22\% \\
+\textbf{Type} & \textbf{Hidden layers} & \textbf{TP} & \textbf{FP} & 
+\textbf{TN} & \textbf{FN} & \textbf{Precision} & \textbf{Recall} & 
+\textbf{F1} \\ \hline
+Unsegmented & 0 & 1.49\% & 3.62\% & 93.26\% & 1.61\% & 29.26\% & 48.17\% 
+& 36.40\% \\
+Unsegmented & 1 & 1.35\% & 3.83\% & 93.06\% & 1.76\% & 26.01\% & 43.29\% 
+& 32.49\% \\
+Unsegmented & 2 & 1.40\% & 4.48\% & 92.41\% & 1.71\% & 23.87\% & 45.12\% 
+& 31.22\% \\
+Unsegmented & 3 & 1.40\% & 4.48\% & 92.41\% & 1.71\% & 23.87\% & 45.12\% 
+& 31.22\% \\
 \hline
-Segmented & 0 & 0.38\% & 1.01\% & 97.35\% & 1.26\% & 27.08\% & 22.94\% & 24.84\% \\
-Segmented & 1 & 0.49\% & 2.90\% & 95.46\% & 1.15\% & 14.49\% & 30.00\% & 19.54\% \\
-Segmented & 2 & 0.22\% & 1.27\% & 97.09\% & 1.42\% & 14.84\% & 13.53\% & 14.15\% \\
-Segmented & 3 & 0.59\% & 3.89\% & 94.47\% & 1.05\% & 13.12\% & 35.88\% & 19.21\% \\
+Segmented & 0 & 0.38\% & 1.01\% & 97.35\% & 1.26\% & 27.08\% & 22.94\% & 
+24.84\% \\
+Segmented & 1 & 0.49\% & 2.90\% & 95.46\% & 1.15\% & 14.49\% & 30.00\% & 
+19.54\% \\
+Segmented & 2 & 0.22\% & 1.27\% & 97.09\% & 1.42\% & 14.84\% & 13.53\% & 
+14.15\% \\
+Segmented & 3 & 0.59\% & 3.89\% & 94.47\% & 1.05\% & 13.12\% & 35.88\% & 
+19.21\% \\ \hline
 \end{tabular} 
-\caption{Auto-encoder classification results for the games that have not been segmented by the color of the active player, and those that have been, as the number of hidden layers varies.}
+\caption{Auto-encoder classification results for the games that have not 
+been segmented by the color of the active player, and those that have 
+been, as the number of hidden layers varies.}
 \label{table:auto-encoder-results}
@@
 
-We first tried to use the unsupervised fine-tuned stacked auto-encoder's hidden outputs as the input to the SVM classifiers, as with the manually-generated feature vectors.
-However, we found that the resulting feature vectors were not very useful for SVM, as the classifiers did not predict any labels as true for all training or test input games.
+We first tried to use the unsupervised fine-tuned stacked auto-encoder's 
+hidden outputs as the input to the SVM classifiers, as with the 
+manually-generated feature vectors. However, we found that the resulting 
+feature vectors were not very useful for SVM, as the classifiers did not 
+predict any labels as true for all training or test input games.
 
-Instead, we used the semi-supervised classification method as described in Section~\ref{sec:classifying-auto-encoder}.
-First, we greedily train each hidden layer as an auto-encoder on the training, test, and validation sets.
-We trained the combined neural network composed of greedily-trained auto-encoders to, for each input, predict an output vector corresponding to the labels that apply to the input---0 if a label did not apply, and 1 if it did.
-From the output vectors computed with the elements in the validation set, we tuned the threshold we use in considering whether a label is present or not present to maximize the F1 score.
-In Table~\ref{table:auto-encoder-results}, we report the performance achieved on the test set when using the threshold tuned to maximize the F1 score on the validation set.
+Instead, we used the semi-supervised classification method as described 
+in Section~\ref{sec:classifying-auto-encoder}. First, we greedily train 
+each hidden layer as an auto-encoder on the training, test, and 
+validation sets. We trained the combined neural network composed of 
+greedily-trained auto-encoders to, for each input, predict an output 
+vector corresponding to the labels that apply to the input---0 if a 
+label did not apply, and 1 if it did. From the output vectors computed 
+with the elements in the validation set, we tuned the threshold we use 
+in considering whether a label is present or not present to maximize the 
+F1 score. In Table~\ref{table:auto-encoder-results}, we report the 
+performance achieved on the test set when using the threshold tuned to 
+maximize the F1 score on the validation set.
 
-These results show higher F1 score than the SVM classifier, with significantly lower precision but also much higher recall.
-Unfortunately, the results indicate that the representation of the board learned by the auto-encoder does not correlates very well with the concepts we wanted to identify, given that having no hidden layers resulted in higher scores.
-Adding extra hidden layers seems to not have helped improve the classification results, except in the last instance, where an extra layer improved the F1 score to original levels.
-It is hard to tell whether some slight change in the auto-encoder training procedure will be able to drastically improve the results, or whether significant changes in the architecture will be required.
-We also were not able to use the unlabeled chess games we collected with the auto-encoder, due to the long training time required and our relatively slow implementation of its training.
+These results show higher F1 score than the SVM classifier, with 
+significantly lower precision but also much higher recall. 
+Unfortunately, the results indicate that the representation of the board 
+learned by the auto-encoder does not correlate very well with the 
+concepts we wanted to identify, given that having no hidden layers 
+resulted in higher scores. Adding extra hidden layers seems to not have 
+helped improve the classification results, except in the last instance, 
+where an extra layer improved the F1 score to original levels. It is 
+hard to tell whether some slight change in the auto-encoder training 
+procedure will be able to drastically improve the results, or whether 
+significant changes in the architecture will be required. We also were 
+not able to use the unlabeled chess games we collected with the 
+auto-encoder, due to the long training time required and our relatively 
+slow implementation of its training.
+
+\subsection{Analysis}
+Given the nature of labeled training data and their assembly from 
+disparate sources, there exists some redundancy among the labels, and 
+games may not have all the labels that they should have in our 
+chess-description ontology. Furthermore, many of the labels had very few 
+(fewer than twenty) games as positive examples, and performance may 
+improve greatly if we had a larger training set to use.
 
 \section{Future Work}
-We identify three areas in which future work can make progress on automated sportscasting for chess.
-It could improve the performance of the auto-encoder so that it can be trained on a larger data set.
-Furthermore, it could focus on trying different methods to improve the classification quality by leveraging chess-specific tools and features like chess engines and evaluation tools.
-Moreover, it could investigate an orthogonal path by applying our general approach to other games, including complex ones like StarCraft and simpler ones like Checkers.
+We identify three areas in which future work can make progress on 
+automated sportscasting. Using a higher-performance auto-encoder 
+training procedure and training it on a larger data set could help 
+improve the results. We could also extend our system with an automated 
+text-generation layer. In addition, leveraging chess-specific tools and 
+features like chess engines and evaluation tools could help improve the 
+quality of the output. Alternatively, an orthogonal path of applying 
+similar general approaches to other games and attempting to learn 
+higher-level constructs, including complex ones like StarCraft and 
+simpler ones like Checkers, merits exploration.
 
 \subsection{Performance}
-One of the main limitations of our work is performance.
-Due to how slowly the auto-encoder performs, we were not able to fully exploit the millions of games that we have available to generate features for our unsupervised system.
-Therefore, extensions to our work should investigate ways to improve the performance so that the auto-encoder can process more inputs and thus generate more relevant feature functions.
+One of the main limitations of our work is performance. Due to the slow 
+performance of the auto-encoder, we were not able to fully exploit the 
+millions of games that we have available to generate features for our 
+unsupervised system. Therefore, extensions to our work should 
+investigate ways to improve the performance so that the auto-encoder can 
+process more inputs and thus generate more relevant feature functions.
 
-\paragraph{Distributed MapReduce.}
-Our auto-encoder takes advantage of some of the parallel processing capacity of the host system through a local MapReduce scheme.
-Extending that with a distributed MapReduce framework like Hadoop may improve performance if the additional cost of network I/O are outweighed by the benefits of parallelization.
+\paragraph{Distributed training.}
+Our auto-encoder takes advantage of some of the parallel processing 
+capacity of the host system through a local MapReduce scheme. Extending 
+that with, for example, a distributed MapReduce framework like Hadoop 
+may improve performance if the additional cost of network I/O are 
+outweighed by the benefits of parallelization.
 
 \paragraph{Using the GPU.}
-Another option that we partially explored is the exploitation of the processing power of the graphics processing unit (GPU).
-Optimization is a computation-heavy task, and the GPU is specifically designed for massively-parallel floating point computations.
-We investigated using JavaCL to parallelize the auto-encoder, and demonstrated that it is definitely superior at parallelizable and computation-intensive tasks like matrix multiplication, because the benefits of greater throughput through greater parallelization outweigh the costs of memory transfer.
+Another option that we partially explored is the exploitation of the 
+processing power of the graphics processing unit (GPU). Function 
+optimization, as used in training the auto-encoder, is a 
+computation-heavy task, and the GPU is specifically designed for 
+massively-parallel floating point computations. We investigated using 
+JavaCL, an interface to using the GPU for general purpose computation, 
+to parallelize the auto-encoder, and demonstrated that it is definitely 
+superior at parallelizable and computation-intensive tasks like matrix 
+multiplication, because the benefits of greater throughput through 
+greater parallelization outweigh the costs of memory transfer.
 
-\subsection{Classification Quality}
-Future work should emphasize the classification quality.
-We demonstrated that general approaches may not be effective enough to be useful for complex games like chess.
-Therefore, future projects should seek improvements by better taking advantage of the chess-specific tools and libraries that are available like game engines and weight functions.
-If the classifier incorporated chess-specific features like those generated by chess engines, it could be much better at identifying the types of strategies that are available to the player.
-For example, a deep traversal of many future paths by a chess engine may reveal that an attack is possible ({\tt attackPossible}) or that the player can force a pawn promotion ({\tt pawnPromotion}).
-While a purely computational approach to identifying opportunities and strategies may provide higher-quality or more reliable assessments, a statistical approach may be more robust to data sparsity.
+\subsection{Text Generation and Linguistic Grounding}
+One further avenue of exploration is a system that automatically 
+generates text given a chess game with tagged positions. The naive 
+approach would be to build a template-based text generation system, 
+which could be as simple as generating a formatted sentence highlighting 
+hand-selected salient features for each tag or something more 
+complicated such as a context-free grammar. In the context of this 
+project, we found that some chess positions were positive examples of 
+multiple tags. As such, another possible avenue of research could be to 
+conglomerate the annotations for each specific position and generate an 
+annotation that makes sense in its context.
 
-One potential improvement that does not involve increased reliance on chess-specific technologies is to implement hierarchical tags so that {\tt matingAttack}, {\tt doubleAttack}, and similar tags imply a more general tag like {\tt attackPossible}.
-By doing so, there will be a greater balance of positives and negatives, and labels that applied to games but were previously considered wrong are now considered correct.
-By incorporating hierarchical structure to tags, the system can jointly train and classify games.
+While we did use high-level structures as defined by experts and 
+discussed in detailed commentary of chess games, we did not attempt to 
+learn and mimic the actual structures used in the natural language 
+commentary. To have a useful commentating system for an intricate game 
+like chess, we need to both discover the high-level structures in the 
+game, as well as know how to talk about the high-level structures. 
+Extending the approach taken by some of the automated sportscasting work 
+discussed in the Related Works section, it may be possible to learn both 
+high-level concepts that are present in a game, and how to realize 
+descriptions of them in language.
 
-Another possible improvement is to use a different classifier than an SVM, such as the perceptron or a maximum entropy classifier.
-One issue with the SVM is that often the tags are not so clear-cut: what does it mean that an endgame has “good chances”?
-Rather, there are degrees of how good an endgame is to the player.
-Therefore, a softer approach like maximum entropy may produce better results.
+\subsection{Improving Results}
+In pursuing this project, we attempted to maintain an approach that is 
+as general as possible so that it can be applied to other contexts, and 
+to minimize the amount of engineering required for components that are 
+very specific to chess. Given the complexity of chess, a relatively 
+straightforward, but chess-specific, way to improve the performance 
+would be to incorporate tools like chess engines. Computer chess has 
+been an active field of research, and much work has gone into game 
+engines and weight functions which excel at the game.
 
-\subsection{Discussion}
-In pursuing this project, we attempted to maintain an approach that is as general as possible so that it can be applied to other contexts, and to minimize the amount of engineering required very specific to chess.
-However, in attempting to maintain generality, we sacrificed access to powerful but chess-specific tools like chess engines.%, resulting in a classifier that is more surface-level (yet slower) than existing chess analysis tools that leverage the game engine.
+For example, a chess-specific rule-based approach should be able to 
+easily identify the more precisely-described themes like {\tt 
+doubleAttack} and {\tt queenCapture} by analyzing the future branches of 
+the game and looking for specific outcomes. However, this approach would 
+require that each theme must have a specific heuristic function written 
+for that express purpose to identify certain properties in future states 
+in order to effectively identify tags using a chess engine. In addition, 
+it may be difficult to identify more subjective themes like {\tt 
+clearance} and {\tt kingToBadSquares} that are based more on human 
+judgement.
 
-A chess-specific rule-based approach should be able to easily identify the more precisely-described themes like {\tt doubleAttack} and {\tt queenCapture} by analyzing the future branches of the game and looking for specific outcomes.
-However, one advantage of the statistical approach is that feature functions can be written once and reused for all possible themes.
-In addition, it would help in identifying more subjective themes like {\tt clearance} and {\tt kingToBadSquares} based more on human judgement.
-In contrast, in order to effectively identify tags using a chess engine, each theme must have a specific heuristic function written for that express purpose to identify certain properties in future states.
-A better automated sportscasting for chess will leverage the strengths of each approach by using the chess-specific features when they are available, and combining that with results from statistical learning.
-For instance, while the chess engine can easily identify imminent mate combinations, it would require a coordination of both the engine and the statistical learning tool to identify threshold functions like {\tt kingToBadSquares}.
-Machine learning would be used to optimize weights and thresholds (e.g.
-for whether a king’s position is bad) while the engine would be used to explore future paths.
-
-Based on the performance of our general approach, we believe that there is an inherent trade off between how general the approach is and the quality of the outcomes.
+This suggests a combined approach which can leverage the strengths of 
+each approach by using the chess-specific features when they are 
+available, and combining that with results from statistical learning. 
+For instance, while the chess engine can easily identify imminent mate 
+combinations, it would require a coordination of both the engine and the 
+statistical learning tool to identify threshold functions like {\tt 
+kingToBadSquares}. Machine learning would be used to optimize weights 
+and thresholds (e.g. for whether a king's position is bad) while the 
+engine would be used to explore future paths. For example, a deep 
+traversal of many future paths by a chess engine may reveal that an 
+attack is possible ({\tt attackPossible}) or that the player can force a 
+pawn promotion ({\tt pawnPromotion}).
 
 \subsection{Generalization to Other Games}
-One goal of this project was to develop a statistical machine learning approach that was general enough to be easily adapted to other applications like StarCraft or Checkers.
-A future work can explore this possibility by simply creating features for other applications.
+One goal of this project was to develop a statistical machine learning 
+approach that was general enough to be easily adapted to other 
+applications like StarCraft or Checkers. Given the light dependence on 
+chess-specific concepts of the approach described in this paper, 
+specifically with the auto-encoder, we should be able to easily adapt 
+this to other turn-based games.
 
 \section{Conclusion}
-We built two statistical learning systems to classify high-level positional concepts for chess, and described various issues involved in implementing these systems.
-Both classifiers are feature driven and classified by a support vector machine.
-The first system uses manually generated features for a chessboard configuration and the move that led to it, while the second system uses a sparse auto-encoder to automatically discover high-level structures (and hence features) for a given chessboard configuration in an unsupervised fashion.
-We found that the auto-encoder-based classifier outperformed the manual feature vector-based classifier, which outperformed the random baseline classifier: on the same segmented data set, the random baseline classifier had an F1 score of $\mathbf{7.60}$, the best manual feature vector-based classifier had an F1 score of $\mathbf{14.05}$, and the best auto-encoder-based classifier had an F1 score of $\mathbf{24.84}$.
+We built two statistical learning systems to classify high-level 
+positional concepts for chess, and described various issues involved in 
+implementing these systems. Both classifiers are feature driven and 
+classified by a support vector machine. The first system uses manually 
+generated features for a chessboard configuration and the move that led 
+to it, while the second system uses a sparse auto-encoder to 
+automatically discover high-level structures (and hence features) for a 
+given chessboard configuration in an unsupervised fashion. We found that 
+the auto-encoder-based classifier outperformed the manual feature 
+vector-based classifier, which outperformed the random baseline 
+classifier: on the same segmented data set, the random baseline 
+classifier had an F1 score of $\mathbf{7.60}$, the best manual feature 
+vector-based classifier had an F1 score of $\mathbf{14.05}$, and the 
+best auto-encoder-based classifier had an F1 score of $\mathbf{24.84}$.
 
 \section{Acknowledgements}
-We would like to thank Dan Klein for teaching us statistical machine learning techniques, and Soda Hall for accommodating us for many nights.
+We would like to thank Dan Klein for teaching us statistical machine 
+learning techniques, Darren Kuo for helping with initial planning and 
+game de-duplication, and Soda Hall for accommodating us for many days 
+and nights.
 

### 2011-05-18 paper/paper.tex
Updated paper stuff...

@@
 \begin{abstract}
-We investigate applications of statistical learning to automated 
-sportscasting for chess. In particular, this project studies two methods 
-to automatically identify high level structures in chess games using 
-support vector machines. Our first classifier is based on manual feature 
-selection, while our second classifier employs sparse auto-encoding to 
-automatically generate a set of feature vectors for a chess game. We 
-then compare the results of the two different approaches and showed 
-improvements using auto-encoding techniques. Finally, we conclude with 
-results, analysis of errors, and ideas for potential future work.
+We investigate applications of statistical learning to automated sportscasting for chess.
+In particular, this project studies two methods to automatically identify high level structures in chess games using support vector machines.
+Our first classifier is based on manual feature selection, while our second classifier employs sparse auto-encoding to automatically generate a set of feature vectors for a chess game.
+We then compare the results of the two different approaches and showed improvements using auto-encoding techniques.
+Finally, we conclude with results, analysis of errors, and ideas for potential future work.
 \end{abstract}
@@
 \section{Introduction}
-Chess has a long history of intriguing enthusiasts due to its 
-complexity. Even after years of play, each game is new challenge, a new 
-problem to solve. Historically, chess has been prominent as a sport 
-where the battles of grandmasters engulf thousands; given the highly 
-intricate nature of chess (and similar abstract strategy games), experts 
-have scrutinized these games to devise standard high-level structures 
-like openings, strategies, tactics, and time advantage, which they use 
-to explain maneuvers and blunders that occur in the games. Devotees of 
-chess also attempt to improve their own performance in terms of these 
-high-level structures, reading chess theorists' comments about games by 
-high-ranking players.
+Chess has a long history of intriguing enthusiasts due to its complexity.
+Even after years of play, each game is new challenge, a new problem to solve.
+Historically, chess has been prominent as a sport where the battles of grandmasters engulf thousands; given the highly intricate nature of chess (and similar abstract strategy games), experts have scrutinized these games to devise standard high-level structures like openings, strategies, tactics, and time advantage, which they use to explain maneuvers and blunders that occur in the games.
+Devotees of chess also attempt to improve their own performance in terms of these high-level structures, reading chess theorists' comments about games by high-ranking players.
 
-In this paper, with the end goal of automatically producing textual 
-commentary similar to those produced by experts for grandmaster games, 
-we investigate whether we might automatically identify such high-level 
-patterns in real games. Specifically, we want to see how statistical 
-machine learning techniques can help us automatically identify such 
-structures using a large number of real chess games as training data, 
-where some of the games may be already annotated with these high-level 
-structures, while others may not be.
+In this paper, with the end goal of automatically producing textual commentary similar to those produced by experts for grandmaster games, we investigate whether we might automatically identify such high-level patterns in real games.
+Specifically, we want to see how statistical machine learning techniques can help us automatically identify such structures using a large number of real chess games as training data, where some of the games may be already annotated with these high-level structures, while others may not be.
 
-While others have created highly rigid and rule-based approaches to 
-annotating chess games~\cite{cambridge-chess-annotation}, we 
-specifically sought to avoid such approaches which require a lot of 
-manual engineering and chess-specific expertise in order to obtain good 
-results. In particular, for chess, there exists a large number of 
-computer chess tools like chess engines, which search the game tree to 
-examine the breadth of options available and identify the outcomes of 
-each. We would like our approaches to have general applicability to 
-other domains, so we chose to specifically avoid using such tools as 
-much as possible.
+While others have created highly rigid and rule-based approaches to annotating chess games~\cite{cambridge-chess-annotation}, we specifically sought to avoid such approaches which require a lot of manual engineering and chess-specific expertise in order to obtain good results.
+In particular, for chess, there exists a large number of computer chess tools like chess engines, which search the game tree to examine the breadth of options available and identify the outcomes of each.
+We would like our approaches to have general applicability to other domains, so we chose to specifically avoid using such tools as much as possible.
 
-To that end, we investigate the use of unsupervised feature learning and 
-deep learning techniques on transcripts of chess games to see if they 
-could automatically identify the high-level structures that make up a 
-chess game. Deep learning techniques, like the sparse stacked 
-auto-encoder, learns a more efficient way to represent some input given 
-a large set of examples. In other fields such as computer vision, 
-researchers have found that these techniques can give state-of-the-art 
-results rivaling or exceeding features hand-engineered by experts. While 
-learning a more compact way to represent parts of chess games, 
-auto-encoders may be able to discover high-level structures similar to 
-those used by human experts.
+To that end, we investigate the use of unsupervised feature learning and deep learning techniques on transcripts of chess games to see if they could automatically identify the high-level structures that make up a chess game.
+Deep learning techniques, like the sparse stacked auto-encoder, learns a more efficient way to represent some input given a large set of examples.
+In other fields such as computer vision, researchers have found that these techniques can give state-of-the-art results rivaling or exceeding features hand-engineered by experts.
+While learning a more compact way to represent parts of chess games, auto-encoders may be able to discover high-level structures similar to those used by human experts.
 
-As a comparison, we also implement a small number of chess-specific, 
-hand-engineered features based on information from a chess engine, such 
-as the number of moves available from a specific board position. Given 
-feature vectors made from hand-labeled games, we train SVM classifiers 
-to predict the label of an unlabeled feature vector. We compare its 
-performance to automatically learned features to see which one can more 
-successfully predict high-level structures.
+As a comparison, we also implement a small number of chess-specific, hand-engineered features based on information from a chess engine, such as the number of moves available from a specific board position.
+Given feature vectors made from hand-labeled games, we train SVM classifiers to predict the label of an unlabeled feature vector.
+We compare its performance to automatically learned features to see which one can more successfully predict high-level structures.
 
-To conduct our experiments, we collect several thousand ``tactical 
-training positions'' organized by theme, intended for use by human chess 
-players to learn how to recognize such situations and improve their 
-playing skills. We also collect a large number of unannotated chess 
-games (about 2.5 million) from various sources on the web for use with 
-semi-supervised and unsupervised learning techniques. After training our 
-models, for each of the positions in the training data labeled with a 
-theme (or high-level structure), we attempt to predict the correct 
-theme(s) for a position using the models.
+To conduct our experiments, we collect several thousand ``tactical training positions'' organized by theme, intended for use by human chess players to learn how to recognize such situations and improve their playing skills.
+We also collect a large number of unannotated chess games (about 2.5 million) from various sources on the web for use with semi-supervised and unsupervised learning techniques.
+After training our models, for each of the positions in the training data labeled with a theme (or high-level structure), we attempt to predict the correct theme(s) for a position using the models.
 
-Unfortunately, we found that our approaches did not deliver accurate 
-results in our experiments. In particular, the auto-encoder's output did 
-not help significantly as a useful criterion for distinguishing between 
-different types of themes. We examine the types of correct predictions 
-made by our classifiers and discuss how we may improve our results in 
-the future.
+Unfortunately, we found that our approaches did not deliver accurate results in our experiments.
+In particular, the auto-encoder's output did not help significantly as a useful criterion for distinguishing between different types of themes.
+We examine the types of correct predictions made by our classifiers and discuss how we may improve our results in the future.
 
@@
 
-\paragraph{Deep learning.} These techniques attempt to capture complex 
-relationships between input variables in order to transform the input 
-into a different representation which better expresses the relevant 
-structures that make up the input. Specifically, they compose many 
-layers of non-linearities in order to compactly represent complex 
-functions which can model these relationships. They have been used 
-successfully for many different types of tasks, especially in computer 
-vision~\cite{Vincent,Lee,Socher}.
+\paragraph{Deep learning.} These techniques attempt to capture complex relationships between input variables in order to transform the input into a different representation which better expresses the relevant structures that make up the input.
+Specifically, they compose many layers of non-linearities in order to compactly represent complex functions which can model these relationships.
+They have been used successfully for many different types of tasks, especially in computer vision~\cite{Vincent,Lee,Socher}.
 
-\paragraph{Automated sportscasting.} Chen and Mooney~\shortcite{chen} 
-present a commentator system that learns language from sportscasts of 
-simulated RoboCup games. Their system, which is trained on ambiguous 
-supervision in the form of textual human commentaries and simulation 
-states of the soccer games, simultaneously tries to establish 
-correspondences between the commentaries and the simulation states. 
-Their work focuses on generating meaningful English commentary from 
-relatively simple topics such as ``Player 1 kicked the ball to Player 
-2'', while our work focuses less on text generation and more on 
-identifying complex topics.
+\paragraph{Automated sportscasting.} Chen and Mooney~\shortcite{chen} present a commentator system that learns language from sportscasts of simulated RoboCup games.
+Their system, which is trained on ambiguous supervision in the form of textual human commentaries and simulation states of the soccer games, simultaneously tries to establish correspondences between the commentaries and the simulation states.
+Their work focuses on generating meaningful English commentary from relatively simple topics such as ``Player 1 kicked the ball to Player 2'', while our work focuses less on text generation and more on identifying complex topics.
 
-\paragraph{Automatic chess commentary.} There exist techniques for 
-automatic annotation of chess games which are heavily dependent on 
-rule-based logic and use of a chess-engine to determine which parts of a 
-specific chess game are salient~\cite{cambridge-chess-annotation}. As a 
-departure from the rule-based approach to automatic chess commentary, 
-Guid et al.~\shortcite{Guid} discuss the use of argument based machine 
-learning, which analyzes reasoning where arguments for and against a 
-certain claim are produced and evaluated (in the context of chess, why a 
-certain move was played over others.) Their system, much like the one 
-presented in this paper, identifies complex positional features of a 
-given chess game.
+\paragraph{Automatic chess commentary.} There exist techniques for automatic annotation of chess games which are heavily dependent on rule-based logic and use of a chess-engine to determine which parts of a specific chess game are salient~\cite{cambridge-chess-annotation}.
+As a departure from the rule-based approach to automatic chess commentary, Guid et al.~\shortcite{Guid} discuss the use of argument based machine learning, which analyzes reasoning where arguments for and against a certain claim are produced and evaluated (in the context of chess, why a certain move was played over others.) Their system, much like the one presented in this paper, identifies complex positional features of a given chess game.
 
-However, we identified some shortcomings of the AMBL system. First, it 
-relies on experts to provide arguments as to why a certain position 
-exemplifies (or doesn.t) a particular positional feature. While this is 
-an improvement from defining a classification rule manually, it is 
-infeasible to repeat this procedure multiple times. Second, the 
-arguments that the experts are allowed to provide rely on a preset list 
-of features.
+However, we identified some shortcomings of the AMBL system.
+First, it relies on experts to provide arguments as to why a certain position exemplifies (or doesn’t) a particular positional feature.
+While this is an improvement from defining a classification rule manually, it is infeasible to repeat this procedure multiple times.
+Second, the arguments that the experts are allowed to provide rely on a preset list of features.
 
 \section{Data Sources and Processing}
-We obtained both annotated and unannotated records of chess games from 
-various web sites in the PGN format \cite{pgn}, a widely-used format for 
-compactly representing chess games. A large number of games as well as 
-software and tools that use the PGN format are freely available.
+We obtained both annotated and unannotated records of chess games from various web sites in the PGN format \cite{pgn}, a widely-used format for compactly representing chess games.
+A large number of games as well as software and tools that use the PGN format are freely available.
 
@@
 \tt{attackPossible} & The current player has opportunity to attack. \\
-\tt{backrankWeakness} & There exists a future checkmate by a rook or 
-queen on the edges of the board. \\
-\tt{bishopPair} & There exists a decisive strategic advantage for the 
-player with two bishops. \\
+\tt{backrankWeakness} & There exists a future checkmate by a rook or queen on the edges of the board. \\
+\tt{bishopPair} & There exists a decisive strategic advantage for the player with two bishops. \\
 \tt{clearance} & There exists a maneuver to break the pawn structure. \\ 
@@
 \tt{decoy} & There exists a maneuver to distract a crucial piece. \\
-\tt{defensePressure} & There exists a maneuver to break the opposing 
-King.s defense. \\
+\tt{defensePressure} & There exists a maneuver to break the opposing King’s defense. \\
 \tt{deflection} & There exists a maneuver to deflect an attack.  \\ 
-\tt{discoveredAttack} & There exists a maneuver to allow for a 
-discovered attack (an attack that was blocked by a friendly piece.) \\
-\tt{doubleAttack} & There exists a move that simultaneously attacks at 
-least two crucial squares. \\
-\tt{drawCombinationsStalemate} & There exists a combination that allows 
-for decisive material advantage or immediate checkmate. \\
-\tt{enclosedKings} & There exists a maneuver to trap the opposing King. 
-\\
-\tt{freeingBlockingSquares} & There exists a move to free a crucial 
-square. \\
-\tt{goodEndgameChances} & There exists a move to allow for a favorable 
-endgame. \\
-\tt{inBetweenMove} & There exists an interrupting move that allows for 
-decisive material advantage of immediate checkmate. \\ 
+\tt{discoveredAttack} & There exists a maneuver to allow for a discovered attack (an attack that was blocked by a friendly piece.) \\
+\tt{doubleAttack} & There exists a move that simultaneously attacks at least two crucial squares. \\
+\tt{drawCombinationsStalemate} & There exists a combination that allows for decisive material advantage or immediate checkmate. \\
+\tt{enclosedKings} & There exists a maneuver to trap the opposing King. \\
+\tt{freeingBlockingSquares} & There exists a move to free a crucial square. \\
+\tt{goodEndgameChances} & There exists a move to allow for a favorable endgame. \\
+\tt{inBetweenMove} & There exists an interrupting move that allows for decisive material advantage of immediate checkmate. \\ 
 \tt{interference} & There exists a maneuver to block a crucial piece. \\ 
-\tt{interruption} & There exists a maneuver that interrupts the 
-opponent.s tactics. \\
-\tt{kingInCenter} & There exists a decisive strategic advantage for the 
-player with the King in the center of the board. \\
-\tt{kingToBadSquares} & There exists a combination that forces the 
-opposing King to move to a bad square. \\
-\tt{longDiagonals} & There exists a strategic advantage for the player 
-that controls squares on the long diagonal. \\
+\tt{interruption} & There exists a maneuver that interrupts the opponent’s tactics. \\
+\tt{kingInCenter} & There exists a decisive strategic advantage for the player with the King in the center of the board. \\
+\tt{kingToBadSquares} & There exists a combination that forces the opposing King to move to a bad square. \\
+\tt{longDiagonals} & There exists a strategic advantage for the player that controls squares on the long diagonal. \\
 \tt{matingAttack} & There exists a decisive mating attack. \\
-\tt{openingClosingDiagonals} & There exists a maneuver to clear a 
-diagonal. \\
+\tt{openingClosingDiagonals} & There exists a maneuver to clear a diagonal. \\
 \tt{openingClosingFiles} & There exists a maneuver to clear a file. \\
-\tt{overload} & There exists a combination that makes a defending piece 
-defend too many squares at once. \\
-\tt{passedPawn} & There exists a combination that allows for a passed 
-pawn (a pawn with no opposing pawns in front of it.) \\
-\tt{pawnPromotion} & There exists a combination that allows for a pawn 
-promotion. \\
-\tt{perpetualCheck} & There exists a combination that forces a perpetual 
-check/draw. \\
+\tt{overload} & There exists a combination that makes a defending piece defend too many squares at once. \\
+\tt{passedPawn} & There exists a combination that allows for a passed pawn (a pawn with no opposing pawns in front of it.) \\
+\tt{pawnPromotion} & There exists a combination that allows for a pawn promotion. \\
+\tt{perpetualCheck} & There exists a combination that forces a perpetual check/draw. \\
 \tt{pin} & There exists a combination that allows for a pin. \\
-\tt{pullingPiecesToBadSquares} & There exists a combination that forces 
-an opposing piece to move to a bad square. \\
-\tt{queenCapture} & There exists a combination to capture the opposing 
-Queen. \\
-\tt{quietMates} & There exists a quiet move (one that looks 
-inconsequential on the surface) leading to mate. \\
-\tt{removalOfDefence} & There exists a maneuver to remove an opponent.s 
-defender. \\
+\tt{pullingPiecesToBadSquares} & There exists a combination that forces an opposing piece to move to a bad square. \\
+\tt{queenCapture} & There exists a combination to capture the opposing Queen. \\
+\tt{quietMates} & There exists a quiet move (one that looks inconsequential on the surface) leading to mate. \\
+\tt{removalOfDefence} & There exists a maneuver to remove an opponent’s defender. \\
 \tt{unpinning} & There exists a maneuver to unpin a piece. \\
-\tt{wrongBishop} & The opponent.s only bishop is limited in use because 
-of the pawn structure of the game. \\ \hline
+\tt{wrongBishop} & The opponent’s only bishop is limited in use because of the pawn structure of the game. \\ \hline
 \end{tabular}
-\caption{List of high-level themes we attempt to automatically identify, 
-along with explanations}
+\caption{List of high-level themes we attempt to automatically identify, along with explanations}
 \label{table:themes}
@@
 
-The already-annotated games come from databases of short snippets of 
-chess 
-games,\footnote{\url{http://wwwu.uni-klu.ac.at/gossimit/c/tactic.htm}} 
-\footnote{\url{http://webplaza.pt.lu/public/ckaber/Chess.htm#Chess 
-training}} intended for chess players to learn from by example, which 
-are organized by particular themes that the students are to identify. 
-These provided a useful repository of both interesting high-level 
-concepts we could identify (we provide a listing in 
-Table~\ref{table:themes}), as well as boards and moves corresponding to 
-each one.
+The already-annotated games come from databases of short snippets of chess games,\footnote{\url{http://wwwu.uni-klu.ac.at/gossimit/c/tactic.htm}} \footnote{\url{http://webplaza.pt.lu/public/ckaber/Chess.htm#Chess training}} intended for chess players to learn from by example, which are organized by particular themes that the students are to identify.
+These provided a useful repository of both interesting high-level concepts we could identify (we provide a listing in Table~\ref{table:themes}), as well as boards and moves corresponding to each one.
 
-We also looked for as many chess games as possible from a variety of web 
-sites, such as \url{chessgames.com}. We easily obtained over five 
-million untagged games spread over 600,000 files and 5 GB of data. The 
-number of annotated game snippets we obtained was slightly over 5,000.
+We also looked for as many chess games as possible from a variety of web sites, such as \url{chessgames.com}.
+We easily obtained over five million untagged games spread over 600,000 files and 5 GB of data.
+The number of annotated game snippets we obtained was slightly over 5,000.
 
 \subsection{Data Pre-Processing}	
-In order to ensure consistency and integrity over the games obtained 
-from disparate sources, we sanitized them in multiple ways. We parsed 
-each game in the data set and removed information that would confuse the 
-parser. For example, some games had English or German comments by humans 
-which we did not plan to use, and sometimes resulted in parsing errors 
-due to specific characters present in the comments. Others had alternate 
-branches, sometimes up to multiple levels (alternate branches within 
-alternate branches), which the parser could not handle properly. Still 
-others represented games in a non-standard way (for example, pawn 
-promotions were represented without the ``=''), which we had to fix.
+In order to ensure consistency and integrity over the games obtained from disparate sources, we sanitized them in multiple ways.
+We parsed each game in the data set and removed information that would confuse the parser.
+For example, some games had English or German comments by humans which we did not plan to use, and sometimes resulted in parsing errors due to specific characters present in the comments.
+Others had alternate branches, sometimes up to multiple levels (alternate branches within alternate branches), which the parser could not handle properly.
+Still others represented games in a non-standard way (for example, pawn promotions were represented without the ``=''), which we had to fix.
 
-Since we obtained the games from many sources, there was significant 
-overlap in games (approximately 50\% of the unlabeled games collected 
-were redundant). Therefore, we also went through all of the downloaded 
-games and eliminated duplicate games: multiple games that had the same 
-initial position and move sequence were considered redundant and only 
-one of each were kept.
+Since we obtained the games from many sources, there was significant overlap in games (approximately 50\% of the unlabeled games collected were redundant).
+Therefore, we also went through all of the downloaded games and eliminated duplicate games: multiple games that had the same initial position and move sequence were considered redundant and only one of each were kept.
 
 \subsection{Parsing}
-We tried various open source PGN parsers, each succumbing to flaws like 
-mishandling uncommon cases (such as  promotions and special moves like 
-\emph{en passant}). We based our system on 
-Chesspresso,\footnote{\url{http://www.chesspresso.org/}} a Java library 
-for writing chess programs, which we fixed and modified for our needs. 
-Chesspresso takes a PGN file as input and returns a data structure that 
-represents the initial position of the board and the subsequent moves 
-(as well as all metadata necessary to recreate the PGN file). We also 
-extended it to be more robust to variations in input format, in addition 
-to being able to export PGNs in a canonical form, which was crucial for 
-removing duplicates.
+We tried various open source PGN parsers, each succumbing to flaws like mishandling uncommon cases (such as  promotions and special moves like \emph{en passant}).
+We based our system on Chesspresso,\footnote{\url{http://www.chesspresso.org/}} a Java library for writing chess programs, which we fixed and modified for our needs.
+Chesspresso takes a PGN file as input and returns a data structure that represents the initial position of the board and the subsequent moves (as well as all metadata necessary to recreate the PGN file).
+We also extended it to be more robust to variations in input format, in addition to being able to export PGNs in a canonical form, which was crucial for removing duplicates.
 
 \subsection{Preparing the Data}
-After parsing each game, we filtered out the tagged games that had more 
-than 10 plies because it will be very difficult to identify the tags 
-that apply to a complex sequence of moves. Among the games that we kept, 
-we also kept only the first few positions and moves (the ``prefix'' of 
-each game, as described earlier), a parameter that can be adjusted; for 
-our experiments, we kept just the initial board specified in each game. 
-We then generate labels for each game based on the information from the 
-source websites. For example, games in a file that corresponds to a 
-lesson called ``Double Attack'' are labeled {\tt doubleAttack}. In 
-addition, some of those games may appear in other lessons, and 
-consequently will have multiple labels. We also segment the games 
-further by the color of the player that makes the first move in the 
-game, in addition to the labels of the game, since most of the labels in 
-Table~\ref{table:themes} only refer to a specific player; for instance, 
-some of the games labeled with {\tt backrankWeakness} may have the white 
-player with the weakness, while others show the black player with the 
-weakness.
+After parsing each game, we filtered out the tagged games that had more than 10 plies because it will be very difficult to identify the tags that apply to a complex sequence of moves.
+Among the games that we kept, we also kept only the first few positions and moves (the ``prefix'' of each game, as described earlier), a parameter that can be adjusted; for our experiments, we kept just the initial board specified in each game.
+We then generate labels for each game based on the information from the source websites.
+For example, games in a file that corresponds to a lesson called ``Double Attack'' are labeled {\tt doubleAttack}.
+In addition, some of those games may appear in other lessons, and consequently will have multiple labels.
+We also segment the games further by the color of the player that makes the first move in the game, in addition to the labels of the game, since most of the labels in Table~\ref{table:themes} only refer to a specific player; for instance, some of the games labeled with {\tt backrankWeakness} may have the white player with the weakness, while others show the black player with the weakness.
 
@@
 \subsection{Sparse auto-encoder}
-The sparse auto-encoder is one of many methods in use for unsupervised 
-feature learning~\cite{Coates}. It takes the form of a feedforward 
-neural network, which consists of an input layer of N nodes, some number 
-of hidden layers, and output layer of M nodes. The value of each layer 
-$\mathbf{l_i}$ is determined with a weight matrix $W_i$ and a bias 
-vector $b_i$: \[\mathbf{l_i} = f(W_i \cdot \mathbf{l_{i-1}} + b_i),\] 
-where $f(\cdot)$ is a non-linearity like the sigmoid $f(x) = \frac{1}{1 
-+ e^{-z}}$ applied component-wise.
+The sparse auto-encoder is one of many methods in use for unsupervised feature learning~\cite{Coates}.
+It takes the form of a feedforward neural network, which consists of an input layer of N nodes, some number of hidden layers, and output layer of M nodes.
+The value of each layer $\mathbf{l_i}$ is determined with a weight matrix $W_i$ and a bias vector $b_i$: \[\mathbf{l_i} = f(W_i \cdot \mathbf{l_{i-1}} + b_i),\] where $f(\cdot)$ is a non-linearity like the sigmoid $f(x) = \frac{1}{1 + e^{-z}}$ applied component-wise.
 
-An auto-encoder is such a neural network trained so as to minimize the 
-error (specifically, the sum-squared difference) between the output 
-layer and the input layer, over some set of input examples. If we have 
-fewer nodes in the hidden layers than in the input or output layer, the 
-neural network is forced to learn a compressed representation of the 
-data (represented by the activations, or outputs, of the smaller number 
-of hidden nodes). Instead, we keep a larger number of hidden nodes while 
-imposing a sparsity constraint that the average activation of each one 
-be small (e.g., 0.05). Other work has shown that such a constraint 
-better forces the auto-encoder to learn interesting structures from the 
-data~\cite{bengio}.
+An auto-encoder is such a neural network trained so as to minimize the error (specifically, the sum-squared difference) between the output layer and the input layer, over some set of input examples.
+If we have fewer nodes in the hidden layers than in the input or output layer, the neural network is forced to learn a compressed representation of the data (represented by the activations, or outputs, of the smaller number of hidden nodes).
+Instead, we keep a larger number of hidden nodes while imposing a sparsity constraint that the average activation of each one be small (e.g., 0.05).
+Other work has shown that such a constraint better forces the auto-encoder to learn interesting structures from the data~\cite{bengio}.
 
-A normal auto-encoder only has one layer of hidden nodes, but we can 
-also add more hidden layers to form a stacked auto-encoder. Per Bengio 
-et al.~\shortcite{Bengio07greedylayer-wise}, we use a greedy strategy to 
-train the extra hidden layers in the auto-encoder. We train an 
-auto-encoder with one hidden layer as usual; once this completes, we 
-obtain the output of the hidden layer over all the examples in the input 
-set, and use that as the new set of examples. This allows us to learn 
-higher-level representations of the data~\cite{Vincent}.
+A normal auto-encoder only has one layer of hidden nodes, but we can also add more hidden layers to form a stacked auto-encoder.
+Per Bengio et al.~\shortcite{Bengio07greedylayer-wise}, we use a greedy strategy to train the extra hidden layers in the auto-encoder.
+We train an auto-encoder with one hidden layer as usual; once this completes, we obtain the output of the hidden layer over all the examples in the input set, and use that as the new set of examples.
+This allows us to learn higher-level representations of the data~\cite{Vincent}.
 
-Other strategies for unsupervised feature learning include restricted 
-Boltzmann machines (which are undirected), deep belief nets (stacked 
-RBMs, akin to stacked auto-encoders), and denoising auto-encoders (which 
-try to reconstruct some input from a corrupted version of it). We only 
-evaluated sparse stacked auto-encoders in this paper.
+Other strategies for unsupervised feature learning include restricted Boltzmann machines (which are undirected), deep belief nets (stacked RBMs, akin to stacked auto-encoders), and denoising auto-encoders (which try to reconstruct some input from a corrupted version of it).
+We only evaluated sparse stacked auto-encoders in this paper.
 
-The auto-encoder requires training data in the form of vectors where all 
-the elements are between 0 and 1, or the range of the sigmoid function. 
-We generated the training vectors by, for some chess board and for every 
-type of piece (knight, rook, etc.), whether a white one (valued 1.0), a 
-black one (valued 0.0), or none at all (valued 0.5) was present in every 
-square of the chessboard.
+The auto-encoder requires training data in the form of vectors where all the elements are between 0 and 1, or the range of the sigmoid function.
+We generated the training vectors by, for some chess board and for every type of piece (knight, rook, etc.), whether a white one (valued 1.0), a black one (valued 0.0), or none at all (valued 0.5) was present in every square of the chessboard.
 
@@
 \textbf{Category} & \textbf{Description} \\ \hline
-Position & Piece location, potential moves per piece, material count, 
-square control. \\
-Previous move & Capturing move, recapturing move, \emph{en passant} 
-move, long castling move, short castling move, check move, stalemate 
-move, checkmate move, promotion move. \\
+Position & Piece location, potential moves per piece, material count, square control. \\
+Previous move & Capturing move, recapturing move, \emph{en passant} move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move. \\
 \hline
@@
 
-As a point of comparison to auto-encoders, we extracted chess-specific 
-features manually from a given chess position (configuration of pieces 
-on the board) and the move that resulted in that position; 
-Table~\ref{table:features} contains a list. The position-based features 
-include information such as where the pieces are on the board and 
-potential moves, while the move-based features involve binary 
-information such as if the move that resulted in this position was a 
-capturing move or resulted in a check.  The resulting feature vector was 
-then L2-normalized to fit the specifications of the support vector 
-machine.
+As a point of comparison to auto-encoders, we extracted chess-specific features manually from a given chess position (configuration of pieces on the board) and the move that resulted in that position; Table~\ref{table:features} contains a list.
+The position-based features include information such as where the pieces are on the board and potential moves, while the move-based features involve binary information such as if the move that resulted in this position was a capturing move or resulted in a check.
+The resulting feature vector was then L2-normalized to fit the specifications of the support vector machine.
 
 \section{Experimental methodology}
-We processed all of the labeled training data to obtain a mapping from 
-each unique game prefix (for most of the experiments, just the initial 
-specified configuration of the board in the labeled game) to the labels 
-that apply to it; most of the prefixes only had one label, while others 
-had two or three. We then segmented the data (consisting of about 3000 
-board configurations) into three disjoint sets: 90\% for training, 5\% 
-as the validation set (for use with tuning our classifiers), and 5\% as 
-the test set. The segmenting was done by random selection from the set 
-of all labeled games.
+We processed all of the labeled training data to obtain a mapping from each unique game prefix (for most of the experiments, just the initial specified configuration of the board in the labeled game) to the labels that apply to it; most of the prefixes only had one label, while others had two or three.
+We then segmented the data (consisting of about 3000 board configurations) into three disjoint sets: 90\% for training, 5\% as the validation set (for use with tuning our classifiers), and 5\% as the test set.
+The segmenting was done by random selection from the set of all labeled games.
 
-In our experiments, we first use the labeled data in the training set 
-(and possibly also unlabeled data) to train the classifiers. Then, for 
-every test input chess game (or a prefix of a game), we use our 
-classifiers to apply some number of labels to it that come from 
-Table~\ref{table:themes}. We measure the performance of the predictions 
-made by our classifiers by counting the number of true positives (a 
-predicted label was among the set of gold labels for a test input), true 
-negatives (a label in Table~\ref{table:themes} was neither in the set of 
-predicted labels nor the set of gold labels), false positives (a 
-predicted label was not in the set of gold labels), and false negatives 
-(a gold label was not predicted). We compute and report the precision, 
-recall, and F1 score using these measures of accuracy.
+In our experiments, we first use the labeled data in the training set (and possibly also unlabeled data) to train the classifiers.
+Then, for every test input chess game (or a prefix of a game), we use our classifiers to apply some number of labels to it that come from Table~\ref{table:themes}.
+We measure the performance of the predictions made by our classifiers by counting the number of true positives (a predicted label was among the set of gold labels for a test input), true negatives (a label in Table~\ref{table:themes} was neither in the set of predicted labels nor the set of gold labels), false positives (a predicted label was not in the set of gold labels), and false negatives (a gold label was not predicted).
+We compute and report the precision, recall, and F1 score using these measures of accuracy.
 
 \subsection{Support vector machine classifier}
-We note that any number of labels may apply to some test input that we 
-want to classify; a straightforward use of a binary (or even multiclass) 
-classifier, as provided by most implementations of support vector 
-machines, will not suffice as it can only assign one label. Therefore, 
-we take the one-versus-all strategy of training many binary SVMs; for 
-each label that we consider, we train a binary classifier with all 
-training games that have the given label as positive examples, and all 
-other games as negative examples. We used {\tt 
-SVMlight}~\cite{svmlight}, a popular third-party implementation of SVMs, 
-for our experiments, using the manually-generated features extracted 
-from each labeled data point.
+We note that any number of labels may apply to some test input that we want to classify; a straightforward use of a binary (or even multiclass) classifier, as provided by most implementations of support vector machines, will not suffice as it can only assign one label.
+Therefore, we take the one-versus-all strategy of training many binary SVMs; for each label that we consider, we train a binary classifier with all training games that have the given label as positive examples, and all other games as negative examples.
+We used {\tt SVMlight}~\cite{svmlight}, a popular third-party implementation of SVMs, for our experiments, using the manually-generated features extracted from each labeled data point.
 
@@
 \label{sec:classifying-auto-encoder}
-There exists several methods for using the auto-encoder for the task at 
-hand. First, we can consider using the output of the last hidden layer 
-as the feature vector for input into the SVM classifier, just as we did 
-with the manually-generated features. To do this, we first perform 
-unsupervised fine-tuning; we combine all the greedily-trained 
-auto-encoders to make a many-layered neural network intended to 
-reconstruct the input layer in its output layer, after which we again 
-minimize the sum-squared error between the input and the output layers. 
-We can then use the output of the hidden layer (i.e., not the one that 
-reconstructs the input, but the middle layer which provides the most 
-high-level representation of the input) as the feature vector.
+There exists several methods for using the auto-encoder for the task at hand.
+First, we can consider using the output of the last hidden layer as the feature vector for input into the SVM classifier, just as we did with the manually-generated features.
+To do this, we first perform unsupervised fine-tuning; we combine all the greedily-trained auto-encoders to make a many-layered neural network intended to reconstruct the input layer in its output layer, after which we again minimize the sum-squared error between the input and the output layers.
+We can then use the output of the hidden layer (i.e., not the one that reconstructs the input, but the middle layer which provides the most high-level representation of the input) as the feature vector.
 
-We can also use semi-supervised learning methods, in which after 
-greedily training the auto-encoders, we combine all the hidden layers of 
-the auto-encoders and add an output layer at the end. For every labeled 
-input game, we construct input-output vectors such that the output 
-encodes the gold labels of the input. We then train the resulting 
-multi-layer neural network to minimize the sum-squared error; this acts 
-similarly to a multiclass logistic regression, except with a different 
-objective at the last layer.
+We can also use semi-supervised learning methods, in which after greedily training the auto-encoders, we combine all the hidden layers of the auto-encoders and add an output layer at the end.
+For every labeled input game, we construct input-output vectors such that the output encodes the gold labels of the input.
+We then train the resulting multi-layer neural network to minimize the sum-squared error; this acts similarly to a multiclass logistic regression, except with a different objective at the last layer.
 
-To train these neural networks, we used batch learning over all the 
-examples at once. The loss function contained components for the 
-sum-squared error as well as an L2 regularization term and a sparsity 
-penalty based on the Kullback-Leibler divergence between the desired 
-average activation of the hidden nodes and the actual average 
-activation. We used backpropagation to compute the gradient of the loss 
-function and L-BFGS to optimize the parameters as to minimize the loss 
-function~\cite{ng-lecturenotes}. Running serially, the training takes a 
-long time to complete, even on simple inputs. Therefore, we examined 
-ways to parallelize the task in order to fully leverage the capacity of 
-modern many-core machines; we achieved a significant speedup by 
-parallelizing the loss function evaluation to 16 execution threads.
+To train these neural networks, we used batch learning over all the examples at once.
+The loss function contained components for the sum-squared error as well as an L2 regularization term and a sparsity penalty based on the Kullback-Leibler divergence between the desired average activation of the hidden nodes and the actual average activation.
+We used backpropagation to compute the gradient of the loss function and L-BFGS to optimize the parameters as to minimize the loss function~\cite{ng-lecturenotes}.
+Running serially, the training takes a long time to complete, even on simple inputs.
+Therefore, we examined ways to parallelize the task in order to fully leverage the capacity of modern many-core machines; we achieved a significant speedup by parallelizing the loss function evaluation to 16 execution threads.
 
 \subsection{Random baseline}
-To use as a baseline and ensure that our classifiers are giving sensible 
-results, we randomly made predictions for every test game, where each 
-one assigned each possible label at random, with the possibility of the 
-label being given based on the prior probability of the label, or the 
-proportion of the training games which have that label. The random 
-predictions were made 1,000 times, with the predictions made 
-independently at random, and the error rates were averaged.
+To use as a baseline and ensure that our classifiers are giving sensible results, we randomly made predictions for every test game, where each one assigned each possible label at random, with the possibility of the label being given based on the prior probability of the label, or the proportion of the training games which have that label.
+The random predictions were made 1,000 times, with the predictions made independently at random, and the error rates were averaged.
 
@@
 \hline
-\textbf{Type} & \textbf{TP} & \textbf{FP} & \textbf{TN} & \textbf{FN} & 
-\textbf{Precision} & \textbf{Recall} & \textbf{F1} \\ \hline
-Unsegmented & 0.21\% & 0.00\% & 96.89\% & 2.90\% & 100\% & 6.71\% & 
-12.57\% \\
-Baseline (Random) & 0.42\% & 2.76\% & 94.12\% & 2.69\% & 13.32\% & 
-13.65\% & 13.48\% \\
+\textbf{Type} & \textbf{TP} & \textbf{FP} & \textbf{TN} & \textbf{FN} & \textbf{Precision} & \textbf{Recall} & \textbf{F1} \\ \hline
+Unsegmented & 0.21\% & 0.00\% & 96.89\% & 2.90\% & 100\% & 6.71\% & 12.57\% \\
+Baseline (Random) & 0.42\% & 2.76\% & 94.12\% & 2.69\% & 13.32\% & 13.65\% & 13.48\% \\
 \hline
-Segmented & 0.13\% & 0.02\% & 98.34\% & 1.51\% & 86.67\% & 7.65\% & 
-14.05\% \\
-Baseline (Random) & 0.12\% & 1.49\% & 96.87\% & 1.51\% & 7.65\% & 7.55\% 
-& 7.60\% \\ \hline
+Segmented & 0.13\% & 0.02\% & 98.34\% & 1.51\% & 86.67\% & 7.65\% & 14.05\% \\
+Baseline (Random) & 0.12\% & 1.49\% & 96.87\% & 1.51\% & 7.65\% & 7.55\% & 7.60\% \\ \hline
 \end{tabular} 
-\caption{Classification results for the games that have not been 
-segmented by the color of the active player, and those that have been. 
-In the former, precision clearly outperforms a random baseline, but 
-recall falls short, resulting in a lower F1 score. In the latter, 
-precision has fallen relative to the unsegmented version, but recall has 
-risen, resulting in a slightly higher F1 score. In contrast, the random 
-model became much worse in both ways.}
+\caption{Classification results for the games that have not been segmented by the color of the active player, and those that have been.
+In the former, precision clearly outperforms a random baseline, but recall falls short, resulting in a lower F1 score.
+In the latter, precision has fallen relative to the unsegmented version, but recall has risen, resulting in a slightly higher F1 score.
+In contrast, the random model became much worse in both ways.}
 \label{table:svmResults}
@@
 
-We first describe the results achieved when we trained using the 
-manually-specified feature extractors. Since each game has very few 
-labels, most of the binary classification results are negative. 
-Therefore, the model is heavily biased toward negative results, as shown 
-in Table~\ref{table:svmResults}.
+We first describe the results achieved when we trained using the manually-specified feature extractors.
+Since each game has very few labels, most of the binary classification results are negative.
+Therefore, the model is heavily biased toward negative results, as shown in Table~\ref{table:svmResults}.
 
-This classification method had relatively high precision compared to the 
-random baseline. However, the recall is much lower than that of the 
-random classifier. As a result, the F1 score (a combination of the 
-precision and recall scores) of the SVM-based model is no better than 
-that of the random model.
+This classification method had relatively high precision compared to the random baseline.
+However, the recall is much lower than that of the random classifier.
+As a result, the F1 score (a combination of the precision and recall scores) of the SVM-based model is no better than that of the random model.
 
-When we extended the labeling to also incorporate the first mover's 
-color information, the SVM model's recall was improved at a minor cost 
-to precision, resulting in a higher F1 score. In contrast, the random 
-predictions became much worse: Both the precision and recall scores fell 
-as a result of this change, which reduced the evenness of the label 
-distribution among the training games, and reduced the amount of games 
-with each label (since games are now spread over twice as many tags).
+When we extended the labeling to also incorporate the first mover's color information, the SVM model's recall was improved at a minor cost to precision, resulting in a higher F1 score.
+In contrast, the random predictions became much worse: Both the precision and recall scores fell as a result of this change, which reduced the evenness of the label distribution among the training games, and reduced the amount of games with each label (since games are now spread over twice as many tags).
 \begin{table}
@@
 \end{tabular} 
-\caption{Percent of incorrect classification (false negatives and false 
-positives) for various tags.}
+\caption{Percent of incorrect classification (false negatives and false positives) for various tags.}
 \label{table:negatives}
@@
 
-Given the high precision but low recall, we delved deeper to analyze the 
-sources of errors and true positives. As we show in 
-Table~\ref{table:negatives}, we found that {\tt deflection} and {\tt 
-removalOfDefence} are much more likely to cause false negatives than 
-other tags. In addition, we found that all of the true positives are for 
-the {\tt quietMates} tag. 9 of the 11 true positive {\tt quietMates} 
-games were classified correctly using only the features that describe 
-the contents of the board. On the other hand, 5 true positive {\tt 
-quietMates} games were correctly classified when using only features 
-about the possible moves in that position. Interestingly, disabling the 
-``checkmate move'' feature did not affect the number of true positives.
+Given the high precision but low recall, we delved deeper to analyze the sources of errors and true positives.
+As we show in Table~\ref{table:negatives}, we found that {\tt deflection} and {\tt removalOfDefence} are much more likely to cause false negatives than other tags.
+In addition, we found that all of the true positives are for the {\tt quietMates} tag.
+9 of the 11 true positive {\tt quietMates} games were classified correctly using only the features that describe the contents of the board.
+On the other hand, 5 true positive {\tt quietMates} games were correctly classified when using only features about the possible moves in that position.
+Interestingly, disabling the ``checkmate move'' feature did not affect the number of true positives.
 
@@
 \hline
-\textbf{Type} & \textbf{Hidden layers} & \textbf{TP} & \textbf{FP} & 
-\textbf{TN} & \textbf{FN} & \textbf{Precision} & \textbf{Recall} & 
-\textbf{F1} \\ \hline
-Unsegmented & 0 & 1.49\% & 3.62\% & 93.26\% & 1.61\% & 29.26\% & 48.17\% 
-& 36.40\% \\
-Unsegmented & 1 & 1.35\% & 3.83\% & 93.06\% & 1.76\% & 26.01\% & 43.29\% 
-& 32.49\% \\
-Unsegmented & 2 & 1.40\% & 4.48\% & 92.41\% & 1.71\% & 23.87\% & 45.12\% 
-& 31.22\% \\
-Unsegmented & 3 & 1.40\% & 4.48\% & 92.41\% & 1.71\% & 23.87\% & 45.12\% 
-& 31.22\% \\
+\textbf{Type} & \textbf{Hidden layers} & \textbf{TP} & \textbf{FP} & \textbf{TN} & \textbf{FN} & \textbf{Precision} & \textbf{Recall} & \textbf{F1} \\ \hline
+Unsegmented & 0 & 1.49\% & 3.62\% & 93.26\% & 1.61\% & 29.26\% & 48.17\% & 36.40\% \\
+Unsegmented & 1 & 1.35\% & 3.83\% & 93.06\% & 1.76\% & 26.01\% & 43.29\% & 32.49\% \\
+Unsegmented & 2 & 1.40\% & 4.48\% & 92.41\% & 1.71\% & 23.87\% & 45.12\% & 31.22\% \\
+Unsegmented & 3 & 1.40\% & 4.48\% & 92.41\% & 1.71\% & 23.87\% & 45.12\% & 31.22\% \\
 \hline
-Segmented & 0 & 0.38\% & 1.01\% & 97.35\% & 1.26\% & 27.08\% & 22.94\% & 
-24.84\% \\
-Segmented & 1 & 0.49\% & 2.90\% & 95.46\% & 1.15\% & 14.49\% & 30.00\% & 
-19.54\% \\
-Segmented & 2 & 0.22\% & 1.27\% & 97.09\% & 1.42\% & 14.84\% & 13.53\% & 
-14.15\% \\
-Segmented & 3 & 0.59\% & 3.89\% & 94.47\% & 1.05\% & 13.12\% & 35.88\% & 
-19.21\% \\ \hline
+Segmented & 0 & 0.38\% & 1.01\% & 97.35\% & 1.26\% & 27.08\% & 22.94\% & 24.84\% \\
+Segmented & 1 & 0.49\% & 2.90\% & 95.46\% & 1.15\% & 14.49\% & 30.00\% & 19.54\% \\
+Segmented & 2 & 0.22\% & 1.27\% & 97.09\% & 1.42\% & 14.84\% & 13.53\% & 14.15\% \\
+Segmented & 3 & 0.59\% & 3.89\% & 94.47\% & 1.05\% & 13.12\% & 35.88\% & 19.21\% \\ \hline
 \end{tabular} 
-\caption{Auto-encoder classification results for the games that have not 
-been segmented by the color of the active player, and those that have 
-been, as the number of hidden layers varies.}
+\caption{Auto-encoder classification results for the games that have not been segmented by the color of the active player, and those that have been, as the number of hidden layers varies.}
 \label{table:auto-encoder-results}
@@
 
-We first tried to use the unsupervised fine-tuned stacked auto-encoder's 
-hidden outputs as the input to the SVM classifiers, as with the 
-manually-generated feature vectors. However, we found that the resulting 
-feature vectors were not very useful for SVM, as the classifiers did not 
-predict any labels as true for all training or test input games.
+We first tried to use the unsupervised fine-tuned stacked auto-encoder's hidden outputs as the input to the SVM classifiers, as with the manually-generated feature vectors.
+However, we found that the resulting feature vectors were not very useful for SVM, as the classifiers did not predict any labels as true for all training or test input games.
 
-Instead, we used the semi-supervised classification method as described 
-in Section~\ref{sec:classifying-auto-encoder}. First, we greedily train 
-each hidden layer as an auto-encoder on the training, test, and 
-validation sets. We trained the combined neural network composed of 
-greedily-trained auto-encoders to, for each input, predict an output 
-vector corresponding to the labels that apply to the input---0 if a 
-label did not apply, and 1 if it did. From the output vectors computed 
-with the elements in the validation set, we tuned the threshold we use 
-in considering whether a label is present or not present to maximize the 
-F1 score. In Table~\ref{table:auto-encoder-results}, we report the 
-performance achieved on the test set when using the threshold tuned to 
-maximize the F1 score on the validation set.
+Instead, we used the semi-supervised classification method as described in Section~\ref{sec:classifying-auto-encoder}.
+First, we greedily train each hidden layer as an auto-encoder on the training, test, and validation sets.
+We trained the combined neural network composed of greedily-trained auto-encoders to, for each input, predict an output vector corresponding to the labels that apply to the input---0 if a label did not apply, and 1 if it did.
+From the output vectors computed with the elements in the validation set, we tuned the threshold we use in considering whether a label is present or not present to maximize the F1 score.
+In Table~\ref{table:auto-encoder-results}, we report the performance achieved on the test set when using the threshold tuned to maximize the F1 score on the validation set.
 
-These results show higher F1 score than the SVM classifier, with 
-significantly lower precision but also much higher recall. 
-Unfortunately, the results indicate that the representation of the board 
-learned by the auto-encoder does not correlate very well with the 
-concepts we wanted to identify, given that having no hidden layers 
-resulted in higher scores. Adding extra hidden layers seems to not have 
-helped improve the classification results, except in the last instance, 
-where an extra layer improved the F1 score to original levels. It is 
-hard to tell whether some slight change in the auto-encoder training 
-procedure will be able to drastically improve the results, or whether 
-significant changes in the architecture will be required. We also were 
-not able to use the unlabeled chess games we collected with the 
-auto-encoder, due to the long training time required and our relatively 
-slow implementation of its training.
+These results show higher F1 score than the SVM classifier, with significantly lower precision but also much higher recall.
+Unfortunately, the results indicate that the representation of the board learned by the auto-encoder does not correlate very well with the concepts we wanted to identify, given that having no hidden layers resulted in higher scores.
+Adding extra hidden layers seems to not have helped improve the classification results, except in the last instance, where an extra layer improved the F1 score to original levels.
+It is hard to tell whether some slight change in the auto-encoder training procedure will be able to drastically improve the results, or whether significant changes in the architecture will be required.
+We also were not able to use the unlabeled chess games we collected with the auto-encoder, due to the long training time required and our relatively slow implementation of its training.
 
 \subsection{Analysis}
-Given the nature of labeled training data and their assembly from 
-disparate sources, there exists some redundancy among the labels, and 
-games may not have all the labels that they should have in our 
-chess-description ontology. Furthermore, many of the labels had very few 
-(fewer than twenty) games as positive examples, and performance may 
-improve greatly if we had a larger training set to use.
+Given the nature of labeled training data and their assembly from disparate sources, there exists some redundancy among the labels, and games may not have all the labels that they should have in our chess-description ontology.
+Furthermore, many of the labels had very few (fewer than twenty) games as positive examples, and performance may improve greatly if we had a larger training set to use.
 
 \section{Future Work}
-We identify three areas in which future work can make progress on 
-automated sportscasting. Using a higher-performance auto-encoder 
-training procedure and training it on a larger data set could help 
-improve the results. We could also extend our system with an automated 
-text-generation layer. In addition, leveraging chess-specific tools and 
-features like chess engines and evaluation tools could help improve the 
-quality of the output. Alternatively, an orthogonal path of applying 
-similar general approaches to other games and attempting to learn 
-higher-level constructs, including complex ones like StarCraft and 
-simpler ones like Checkers, merits exploration.
+We identify three areas in which future work can make progress on automated sportscasting.
+Using a higher-performance auto-encoder training procedure and training it on a larger data set could help improve the results.
+We could also extend our system with an automated text-generation layer.
+In addition, leveraging chess-specific tools and features like chess engines and evaluation tools could help improve the quality of the output.
+Alternatively, an orthogonal path of applying similar general approaches to other games and attempting to learn higher-level constructs, including complex ones like StarCraft and simpler ones like Checkers, merits exploration.
 
 \subsection{Performance}
-One of the main limitations of our work is performance. Due to the slow 
-performance of the auto-encoder, we were not able to fully exploit the 
-millions of games that we have available to generate features for our 
-unsupervised system. Therefore, extensions to our work should 
-investigate ways to improve the performance so that the auto-encoder can 
-process more inputs and thus generate more relevant feature functions.
+One of the main limitations of our work is performance.
+Due to the slow performance of the auto-encoder, we were not able to fully exploit the millions of games that we have available to generate features for our unsupervised system.
+Therefore, extensions to our work should investigate ways to improve the performance so that the auto-encoder can process more inputs and thus generate more relevant feature functions.
 
 \paragraph{Distributed training.}
-Our auto-encoder takes advantage of some of the parallel processing 
-capacity of the host system through a local MapReduce scheme. Extending 
-that with, for example, a distributed MapReduce framework like Hadoop 
-may improve performance if the additional cost of network I/O are 
-outweighed by the benefits of parallelization.
+Our auto-encoder takes advantage of some of the parallel processing capacity of the host system through a local MapReduce scheme.
+Extending that with, for example, a distributed MapReduce framework like Hadoop may improve performance if the additional cost of network I/O are outweighed by the benefits of parallelization.
 
 \paragraph{Using the GPU.}
-Another option that we partially explored is the exploitation of the 
-processing power of the graphics processing unit (GPU). Function 
-optimization, as used in training the auto-encoder, is a 
-computation-heavy task, and the GPU is specifically designed for 
-massively-parallel floating point computations. We investigated using 
-JavaCL, an interface to using the GPU for general purpose computation, 
-to parallelize the auto-encoder, and demonstrated that it is definitely 
-superior at parallelizable and computation-intensive tasks like matrix 
-multiplication, because the benefits of greater throughput through 
-greater parallelization outweigh the costs of memory transfer.
+Another option that we partially explored is the exploitation of the processing power of the graphics processing unit (GPU).
+Function optimization, as used in training the auto-encoder, is a computation-heavy task, and the GPU is specifically designed for massively-parallel floating point computations.
+We investigated using JavaCL, an interface to using the GPU for general purpose computation, to parallelize the auto-encoder, and demonstrated that it is definitely superior at parallelizable and computation-intensive tasks like matrix multiplication, because the benefits of greater throughput through greater parallelization outweigh the costs of memory transfer.
 
 \subsection{Text Generation and Linguistic Grounding}
-One further avenue of exploration is a system that automatically 
-generates text given a chess game with tagged positions. The naive 
-approach would be to build a template-based text generation system, 
-which could be as simple as generating a formatted sentence highlighting 
-hand-selected salient features for each tag or something more 
-complicated such as a context-free grammar. In the context of this 
-project, we found that some chess positions were positive examples of 
-multiple tags. As such, another possible avenue of research could be to 
-conglomerate the annotations for each specific position and generate an 
-annotation that makes sense in its context.
+One further avenue of exploration is a system that automatically generates text given a chess game with tagged positions.
+The naive approach would be to build a template-based text generation system, which could be as simple as generating a formatted sentence highlighting hand-selected salient features for each tag or something more complicated such as a context-free grammar.
+In the context of this project, we found that some chess positions were positive examples of multiple tags.
+As such, another possible avenue of research could be to conglomerate the annotations for each specific position and generate an annotation that makes sense in its context.
 
-While we did use high-level structures as defined by experts and 
-discussed in detailed commentary of chess games, we did not attempt to 
-learn and mimic the actual structures used in the natural language 
-commentary. To have a useful commentating system for an intricate game 
-like chess, we need to both discover the high-level structures in the 
-game, as well as know how to talk about the high-level structures. 
-Extending the approach taken by some of the automated sportscasting work 
-discussed in the Related Works section, it may be possible to learn both 
-high-level concepts that are present in a game, and how to realize 
-descriptions of them in language.
+While we did use high-level structures as defined by experts and discussed in detailed commentary of chess games, we did not attempt to learn and mimic the actual structures used in the natural language commentary.
+To have a useful commentating system for an intricate game like chess, we need to both discover the high-level structures in the game, as well as know how to talk about the high-level structures.
+Extending the approach taken by some of the automated sportscasting work discussed in the Related Works section, it may be possible to learn both high-level concepts that are present in a game, and how to realize descriptions of them in language.
 
 \subsection{Improving Results}
-In pursuing this project, we attempted to maintain an approach that is 
-as general as possible so that it can be applied to other contexts, and 
-to minimize the amount of engineering required for components that are 
-very specific to chess. Given the complexity of chess, a relatively 
-straightforward, but chess-specific, way to improve the performance 
-would be to incorporate tools like chess engines. Computer chess has 
-been an active field of research, and much work has gone into game 
-engines and weight functions which excel at the game.
+In pursuing this project, we attempted to maintain an approach that is as general as possible so that it can be applied to other contexts, and to minimize the amount of engineering required for components that are very specific to chess.
+Given the complexity of chess, a relatively straightforward, but chess-specific, way to improve the performance would be to incorporate tools like chess engines.
+Computer chess has been an active field of research, and much work has gone into game engines and weight functions which excel at the game.
 
-For example, a chess-specific rule-based approach should be able to 
-easily identify the more precisely-described themes like {\tt 
-doubleAttack} and {\tt queenCapture} by analyzing the future branches of 
-the game and looking for specific outcomes. However, this approach would 
-require that each theme must have a specific heuristic function written 
-for that express purpose to identify certain properties in future states 
-in order to effectively identify tags using a chess engine. In addition, 
-it may be difficult to identify more subjective themes like {\tt 
-clearance} and {\tt kingToBadSquares} that are based more on human 
-judgement.
+For example, a chess-specific rule-based approach should be able to easily identify the more precisely-described themes like {\tt doubleAttack} and {\tt queenCapture} by analyzing the future branches of the game and looking for specific outcomes.
+However, this approach would require that each theme must have a specific heuristic function written for that express purpose to identify certain properties in future states in order to effectively identify tags using a chess engine.
+In addition, it may be difficult to identify more subjective themes like {\tt clearance} and {\tt kingToBadSquares} that are based more on human judgement.
 
-This suggests a combined approach which can leverage the strengths of 
-each approach by using the chess-specific features when they are 
-available, and combining that with results from statistical learning. 
-For instance, while the chess engine can easily identify imminent mate 
-combinations, it would require a coordination of both the engine and the 
-statistical learning tool to identify threshold functions like {\tt 
-kingToBadSquares}. Machine learning would be used to optimize weights 
-and thresholds (e.g. for whether a king's position is bad) while the 
-engine would be used to explore future paths. For example, a deep 
-traversal of many future paths by a chess engine may reveal that an 
-attack is possible ({\tt attackPossible}) or that the player can force a 
-pawn promotion ({\tt pawnPromotion}).
+This suggests a combined approach which can leverage the strengths of each approach by using the chess-specific features when they are available, and combining that with results from statistical learning.
+For instance, while the chess engine can easily identify imminent mate combinations, it would require a coordination of both the engine and the statistical learning tool to identify threshold functions like {\tt kingToBadSquares}.
+Machine learning would be used to optimize weights and thresholds (e.g.
+for whether a king's position is bad) while the engine would be used to explore future paths.
+For example, a deep traversal of many future paths by a chess engine may reveal that an attack is possible ({\tt attackPossible}) or that the player can force a pawn promotion ({\tt pawnPromotion}).
 
 \subsection{Generalization to Other Games}
-One goal of this project was to develop a statistical machine learning 
-approach that was general enough to be easily adapted to other 
-applications like StarCraft or Checkers. Given the light dependence on 
-chess-specific concepts of the approach described in this paper, 
-specifically with the auto-encoder, we should be able to easily adapt 
-this to other turn-based games.
+One goal of this project was to develop a statistical machine learning approach that was general enough to be easily adapted to other applications like StarCraft or Checkers.
+Given the light dependence on chess-specific concepts of the approach described in this paper, specifically with the auto-encoder, we should be able to easily adapt this to other turn-based games.
 
 \section{Conclusion}
-We built two statistical learning systems to classify high-level 
-positional concepts for chess, and described various issues involved in 
-implementing these systems. Both classifiers are feature driven and 
-classified by a support vector machine. The first system uses manually 
-generated features for a chessboard configuration and the move that led 
-to it, while the second system uses a sparse auto-encoder to 
-automatically discover high-level structures (and hence features) for a 
-given chessboard configuration in an unsupervised fashion. We found that 
-the auto-encoder-based classifier outperformed the manual feature 
-vector-based classifier, which outperformed the random baseline 
-classifier: on the same segmented data set, the random baseline 
-classifier had an F1 score of $\mathbf{7.60}$, the best manual feature 
-vector-based classifier had an F1 score of $\mathbf{14.05}$, and the 
-best auto-encoder-based classifier had an F1 score of $\mathbf{24.84}$.
+We built two statistical learning systems to classify high-level positional concepts for chess, and described various issues involved in implementing these systems.
+Both classifiers are feature driven and classified by a support vector machine.
+The first system uses manually generated features for a chessboard configuration and the move that led to it, while the second system uses a sparse auto-encoder to automatically discover high-level structures (and hence features) for a given chessboard configuration in an unsupervised fashion.
+We found that the auto-encoder-based classifier outperformed the manual feature vector-based classifier, which outperformed the random baseline classifier: on the same segmented data set, the random baseline classifier had an F1 score of $\mathbf{7.60}$, the best manual feature vector-based classifier had an F1 score of $\mathbf{14.05}$, and the best auto-encoder-based classifier had an F1 score of $\mathbf{24.84}$.
 
 \section{Acknowledgements}
-We would like to thank Dan Klein for teaching us statistical machine 
-learning techniques, Darren Kuo for helping with initial planning and 
-game de-duplication, and Soda Hall for accommodating us for many days 
-and nights.
+We would like to thank Dan Klein for teaching us statistical machine learning techniques, Darren Kuo for helping with initial planning and game de-duplication, and Soda Hall for accommodating us for many days and nights.
 

### 2011-05-18 paper/paper.tex
Submitted paper.
Paper debugged

@@
 However, we identified some shortcomings of the AMBL system.
-First, it relies on experts to provide arguments as to why a certain position exemplifies (or doesn’t) a particular positional feature.
+First, it relies on experts to provide arguments as to why a certain position exemplifies (or doesn't) a particular positional feature.
 While this is an improvement from defining a classification rule manually, it is infeasible to repeat this procedure multiple times.
@@
 \tt{backrankWeakness} & There exists a future checkmate by a rook or queen on the edges of the board. \\
-\tt{bishopPair} & There exists a decisive strategic advantage for the player with two bishops. \\
+\tt{bishopPair} & There exists a decisive advantage for the player with 2 bishops. \\
 \tt{clearance} & There exists a maneuver to break the pawn structure. \\ 
@@
 \tt{decoy} & There exists a maneuver to distract a crucial piece. \\
-\tt{defensePressure} & There exists a maneuver to break the opposing King’s defense. \\
+\tt{defensePressure} & There exists a maneuver to break the opposing King's defense. \\
 \tt{deflection} & There exists a maneuver to deflect an attack.  \\ 
 \tt{discoveredAttack} & There exists a maneuver to allow for a discovered attack (an attack that was blocked by a friendly piece.) \\
-\tt{doubleAttack} & There exists a move that simultaneously attacks at least two crucial squares. \\
+\tt{doubleAttack} & There exists a move that simultaneously attacks two squares. \\
 \tt{drawCombinationsStalemate} & There exists a combination that allows for decisive material advantage or immediate checkmate. \\
